### 2013-11-12 writeup/paper.tex
added paper template

\documentclass[10pt,twocolumn]{article}

\usepackage[margin=1in]{geometry}

\title{TODO: Intrusion Detection through Provenance}
\author{
    Kenny Yu\\
    Harvard University\\
    kennyyu@college.harvard.edu
  \and
    R. J. Aquino\\
    Harvard University\\
    rjaquino@college.harvard.edu
}
\date{November 12, 2013}

\begin{document}

\maketitle

\begin{abstract}
TODO
\end{abstract}

\section{Introduction}
TODO

\section{Provenance Graphs}
TODO

\section{Results}
TODO

\section{Discussion}
TODO

\section{Related Work}
TODO

\section{References}
TODO

\end{document}

### 2013-11-13 writeup/paper.tex
added references to paper

@@
 \documentclass[10pt,twocolumn]{article}
+
+\usepackage{algorithm}
+\usepackage{moreverb}   
+\usepackage{longtable}
+\usepackage{fancyhdr}
+\usepackage{algorithmic}             
+\usepackage{algorithm}
+\usepackage{array}     
+\usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{fullpage}
+\usepackage{amsmath, amssymb, amsthm}
+\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
+\usepackage{mathabx}
+\usepackage{float}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABSTRACT
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
@@
 
-\section{Provenance Graphs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% DESIGN
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Design}
 TODO
 
-\section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% EVALUATION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Evaluation}
 TODO
 
-\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% CONCLUSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusion}
 TODO
 
-\section{Related Work}
-TODO
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% BIBLIOGRAPHY
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{References}
-TODO
+\begin{thebibliography}{99}
+
+\bibitem{somayaji-recent}
+\textsc{Inoue, H. and Somayaji, A.} Lookahead Pairs and Full Sequences: A Tale of Two Anomaly Detection Methods. In {\em 2nd Annual Symposium on Information Assurance} (June 2007).
+
+\bibitem{clustering}
+\textsc{Macko, P., Margo, D., Seltzer, M.} Local Clustering in Provenance Graphs (Extended Version).
+In {\em Proceedings of the 22nd ACM international conference on Conference on information \& knowledge management} (August 2013).
+
+\bibitem{orbiter}
+\textsc{Macko, P.
+and Seltzer, M.} Provenance Map Orbiter: Interactive Exploration of Large Provenance Graphs.
+In {\em TaPP'11 Proceedings of the 2nd conference on Theory and practice of provenance} (June 2011).
+
+\bibitem{fileattributes}
+\textsc{Margo, D., and Smogor, R.} Using Provenance to Extract Semantic File Attributes. In {\em TaPP'10 Proceedings of the 2nd conference on Theory and practice of provenance} (February 2010).
+
+\bibitem{passv2}
+\textsc{Muniswamy-Reddy, K., Braun, U., Holland, D.
+A., Macko, P., Maclean, D., Margo, D., Seltzer, M., and Smogor, R.} Layering in Provenance Systems.
+In {\em Proceedings of the 2009 USENIX Annual Technical Conference} (June 2009).
+
+\bibitem{pass}
+\textsc{Muniswamy-Reddy, K., Holland, D. A., Braun, U., and Seltzer, M.} Provenance-Aware Storage Systems. In {\em Proceedings of the 2006 USENIX Annual Technical Conference} (June 2006).
+
+\bibitem{somayaji}
+\textsc{Somayaji, A. and Forrest, S.} Automated Response Using System-Call Delays. In {\em Proceedings of the 2000 USENIX Annual Technical Conference} (August 2000).
+
+\end{thebibliography}
 

### 2013-11-13 passtools-20100524/fsck.lasagna/fsck.lasagna.sh
remove pass tools

@@ -1,38 +1,0 @@
-#! /bin/sh
-MKDIR=/bin/mkdir
-RMDIR=/bin/rmdir
-MOUNT=/bin/mount
-UMOUNT=/bin/umount
-
-tempdir=/mnt/__passtmp
-
-FAILED=
-
-fsck.ext3 "$@"
-RESULT=$?
-if [ $RESULT -ne 0 ]; then
-    echo "$0: fsck.ext3 failed" 1>&2
-    exit $RESULT
-fi
-
-if [ ! -d $tempdir ]; then
-	$MKDIR $tempdir
-fi
-
-echo "$MOUNT -t ext3 $@ $tempdir"
-$MOUNT -t ext3 "$@" $tempdir
-if ! (
-    cd $tempdir/.lasagna_stuff || exit 1
-    $PREFIX/libexec/passtools/waldo -r || exit 1
-); then
-    FAILED=$?
-fi
-$UMOUNT $tempdir
-$RMDIR $tempdir
-
-if [ "x$FAILED" != x ]; then
-    echo "$0: recovery failed." 1>&2
-    exit $FAILED
-fi
-
-exit 0

### 2013-11-13 passtools-20100524/include/debug.h

@@ -1,40 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef DEBUG_H
-#define DEBUG_H
-
-// The following is for printing debugging messages
-#ifdef DEBUG
-#define dprintf(...) printf(__VA_ARGS__)
-#else
-#define dprintf(...)
-#endif
-
-#endif /* DEBUG_H */

### 2013-11-13 passtools-20100524/include/pass/lasagna.h

@@ -1,30 +1,0 @@
-#ifndef _LASAGNA_H
-#define _LASAGNA_H
-
-#define LASAGNA_DIR               ".lasagna_stuff/"
-
-#define LASAGNA_METADB_FILENAME   (LASAGNA_DIR "lasagna.metadb")
-#define LASAGNA_CLEAN_FILENAME    (LASAGNA_DIR "clean")
-#define LASAGNA_NEXTNUMS_FILENAME (LASAGNA_DIR "lasagna.state")
-
-/* provenance object state: either frozen or thawed */
-enum prov_pnode_state {
-    PROV_STATE_THAWED,
-    PROV_STATE_FROZEN
-};
-
-struct lasagna_metadb_entry {
-    unsigned long ino;
-    pnode_t       pnode;
-    version_t     version;
-    icapi_flags_t icapi_flags; /* unused */
-    enum prov_pnode_state state;
-};
-
-struct lasagna_nextnums {
-    pnode_t   pnode;
-    lsn_t     lsn;
-    uint32_t  lognum;
-} __attribute__((__packed__));
-
-#endif /* _LASAGNA_H */

### 2013-11-13 passtools-20100524/include/pass/provabi.h

@@ -1,222 +1,0 @@
-#ifndef _PROVABI_H
-#define _PROVABI_H
-
-#ifdef __KERNEL__
-#include <linux/types.h>
-#else
-#include <sys/types.h>
-#include <stdint.h>
-#define __user
-#endif
-
-#define PROV_ABI_VERSION 2
-
-/*
- * Base types.
- */
-
-#ifdef __KERNEL__
-typedef __u64	__pnode_t;          /* provenance node */
-typedef __u32   icapi_flags_t;      /* icapi flags */
-typedef __u32	version_t;	    /* provenance version */
-typedef __u64   provtxn_t;          /* provenance transaction id */
-#else
-typedef uint64_t __pnode_t;         /* provenance node */
-typedef uint32_t icapi_flags_t;     /* icapi flags */
-typedef uint32_t version_t;         /* provenance version */
-typedef uint64_t provtxn_t;          /* provenance transaction id */
-#endif
-
-/*
- * Constants.
- */
-
-/* Provenance record value types for dv_type et al. */
-#define PROV_TYPE_NIL		0	/* nothing */
-#define PROV_TYPE_STRING	1	/* string */
-#define PROV_TYPE_MULTISTRING	2	/* zero or more strings */
-#define PROV_TYPE_INT		3	/* int32_t */
-#define PROV_TYPE_REAL		4	/* double */
-#define PROV_TYPE_TIMESTAMP	5	/* struct prov_timestamp */
-#define PROV_TYPE_INODE		6	/* uint32_t */
-#define PROV_TYPE_PNODE		7	/* pnode_t */
-#define PROV_TYPE_PNODEVERSION	8	/* struct prov_pnodeversion */
-#define PROV_TYPE_OBJECT	9	/* context-dependent object type */
-#define PROV_TYPE_OBJECTVERSION	10	/* same, plus version_t */
-/* XXX these three should go away */
-#define PROV_TYPE_FREEZE        11      /* bump the object up to next version */
-#define PROV_TYPE_SUB           12      /* belongs to a sub transction */
-#define PROV_TYPE_ENDSUB        13      /* end sub transction */
-
-/* Flags for DPAPI dp_flags */
-#define PROV_IS_ANCESTRY  1	/* Ancestry record (otherwise, identity) */
-
-/* Conversions for DPAPI da_conversion */
-typedef enum {
-   PROV_CONVERT_NONE,			/* nothing */
[…]

### 2013-11-13 passtools-20100524/include/pql.h

@@ -1,207 +1,0 @@
-/*
- * Copyright 2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef PQL_H
-#define PQL_H
-
-#include <stddef.h>
-#include <stdint.h>
-#include <stdbool.h>
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-struct pqlcontext;
-struct pqlquery;
-struct pqlvalue;
-
-typedef uint64_t pqloid_t;
[…]

### 2013-11-13 passtools-20100524/include/primarray.h

@@ -1,99 +1,0 @@
-//
-// 	File:        primarray.h
-// 	Written:     30-Jul-95
-// 	Updated:     29-Oct-99
-//	Description: array template for primitive types
-//
-// Copyright (c) 1995, 1999 David A. Holland.
-// All rights reserved.
-//
-// Redistribution and use in source and binary forms, with or without
-// modification, are permitted provided that the following conditions
-// are met:
-// 1. Redistributions of source code must retain the above copyright
-//    notice, this list of conditions and the following disclaimer.
-// 2. Redistributions in binary form must reproduce the above copyright
-//    notice, this list of conditions and the following disclaimer in the
-//    documentation and/or other materials provided with the distribution.
-// 3. Neither the name of the Author nor the names of any contributors
-//    may be used to endorse or promote products derived from this software
-//    without specific prior written permission.
-//
-// THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
-// ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
-// IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
-// ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
-// FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
-// DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
-// OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
-// HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
-// LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
-// OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
-// SUCH DAMAGE.
-//
-//   $Id: primarray.h,v 1.1 2007/03/16 00:42:17 dholland Exp $
-//
-
-#ifndef PRIMARRAY_H
-#define PRIMARRAY_H
-
-#ifndef assert
-#include <assert.h>
-#endif
-
[…]

### 2013-11-13 passtools-20100524/include/ptrarray.h

@@ -1,98 +1,0 @@
-//
-// 	File:        ptrarray.h
-// 	Written:     16-Jul-95
-// 	Updated:     29-Oct-99
-//	Description: Array of pointers
-//
-// Copyright (c) 1995, 1999 David A. Holland.
-// All rights reserved.
-//
-// Redistribution and use in source and binary forms, with or without
-// modification, are permitted provided that the following conditions
-// are met:
-// 1. Redistributions of source code must retain the above copyright
-//    notice, this list of conditions and the following disclaimer.
-// 2. Redistributions in binary form must reproduce the above copyright
-//    notice, this list of conditions and the following disclaimer in the
-//    documentation and/or other materials provided with the distribution.
-// 3. Neither the name of the Author nor the names of any contributors
-//    may be used to endorse or promote products derived from this software
-//    without specific prior written permission.
-//
-// THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
-// ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
-// IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
-// ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
-// FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
-// DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
-// OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
-// HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
-// LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
-// OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
-// SUCH DAMAGE.
-//
-//   $Id: ptrarray.h,v 1.1 2007/03/16 00:42:17 dholland Exp $
-//
-
-#ifndef PTRARRAY_H
-#define PTRARRAY_H
-
-#ifndef assert
[…]

### 2013-11-13 passtools-20100524/include/twig_file.h

@@ -1,67 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef TWIG_FILE_H
-#define TWIG_FILE_H
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-/**
- * Provides an interface to twig format files.
- *
- * Supports the standard open/close/read/write.
- */
-#include <stdint.h>
-
-/**
- * There are two valid open modes:
[…]

### 2013-11-13 passtools-20100524/include/wdb.h

@@ -1,121 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef _WDB_H
-#define _WDB_H
-
-#include <db.h>
-#include "schema.h"
-
-//
-// Constants
-//
-
-#define TRUE                 1
-#define FALSE                0
-
-//
-// WDB Open flags
-//
-
-// open modes
[…]

### 2013-11-13 passtools-20100524/libpass-java/libpass-java.c

@@ -1,238 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#include <stdlib.h>
-#include <stdio.h>
-#include <string.h>
-#include <sys/types.h>
-#include <sys/ioctl.h>
-#include <unistd.h>
[…]

### 2013-11-13 passtools-20100524/libpass-java/org/harvard/pass/DPAPI.java

@@ -1,110 +1,0 @@
-package org.harvard.pass;
-
-import java.io.File;
-
-/**
- * The Java interface to PASS DPAPI
- *
- * @author Peter Macko
- */
-public class DPAPI
-{
-	static {
-		System.loadLibrary("pass-java");
-		init();
-	}
-	
-	/**
-	 * Initialize the PASS DPAPI
-	 */
-	private static native int _init();
-	
-	/**
-	 * Initialize the PASS DPAPI
-	 */
-	private static void init()
-	{
-	    int r = _init();
-	    if (r < 0) throw new RuntimeException("Cannot initialize DPAPI, error code: " + r);
-	}
-	
-	
-	/**
-	 * Open a file and return the handle
-	 *
-	 * @param name the filename
-	 * @param forWriting whether the file should also be opened for writing
-	 * @return the file handle (may need to be closed)
-	 */
-	public static native int openFile(String name, boolean forWriting);
-	
-	/**
-	 * Close a file handle
-	 *
-	 * @param fd the file handle
-	 */
-	public static native void closeHandle(int fd);
-	
-	/**
-	 * Create a phony P-node
-	 * 
-	 * @param adj the adjacent node
-	 * @return the descriptor of a created node
-	 */
-	private static native int _createNode(int adj);
-	
-	/**
-	 * Create a phony P-node
-	 * 
-	 * @param adj the adjacent node
-	 * @return the descriptor of a created node
-	 */
-	public static int createNode(int adj)
-	{
-	    int r = _createNode(adj);
-	    if (r < 0) throw new RuntimeException("Cannot create a P-Node, error code: " + r);
-	    return r;
-	}
-	
-	/**
-	 * Create a phony P-node
-	 * 
-	 * @return the descriptor of a created node
-	 */
-	public static int createNode() { return createNode(-1); }
-	
-	/**
-	 * Freeze a P-node
-	 * 
-	 * @param fd the p-node handle or the file descriptor
-	 */
-	public static native void freeze(int fd);
-	
-	/**
-	 * Add an ancestry cross-reference
-	 * 
-	 * @param fd the p-node to add the reference to
-	 * @param key the key of the reference
-	 * @param xref the cross-reference p-node
-	 */
-	public static native void addXRef(int fd, String key, int xref);
-	
-	/**
-	 * Add an ancestry cross-reference
-	 * 
-	 * @param fd the p-node to add the reference to
-	 * @param key the key of the reference
-	 * @param xref the cross-reference p-node
[…]

### 2013-11-13 passtools-20100524/libpass-python/libpass-python.c

@@ -1,140 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#include <stdlib.h>
-#include <stdio.h>
-#include <sys/types.h>
-#include <sys/ioctl.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <errno.h>
-
-#include <pass/dpapi.h>
[…]

### 2013-11-13 passtools-20100524/libpass/libpass.c

@@ -1,31 +1,0 @@
-#include <sys/types.h>
-#include <sys/ioctl.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <errno.h>
-#include <err.h>
-
-#include <pass/provabi.h>
-#include "libpass.h"
-
-int __libpass_hook_fd = -1;
-int __libpass_initialized = 0;
-
-int __libpass_initialize(void) {
-   int ver;
-
-   __libpass_hook_fd = open(_PATH_DEV_PROVENANCE, O_RDWR);
-   if (__libpass_hook_fd < 0) {
-      return -1;
-   }
-   if (ioctl(__libpass_hook_fd, PASSIOCGETABI, &ver) < 0) {
-      return -1;
-   }
-   if (ver != PROV_ABI_VERSION) {
-      warnx("Wrong provenance ABI version");
-      errno = ENXIO;
-      return -1;
-   }
-   __libpass_initialized = 1;
-   return 0;
-}

### 2013-11-13 passtools-20100524/libpasstest/dpapi.c

@@ -1,268 +1,0 @@
-#include <stdarg.h>
-#include <stdio.h>
-#include <string.h>
-#include <unistd.h>
-#include <err.h>
-#include <errno.h>
-#include <fcntl.h>
-#include <paths.h>
-#include <assert.h>
-
-#include <pass/dpapi.h>
-#include "libpasstest.h"
-
-#ifndef DPAPI_SELF // XXX this should get added already
-#define DPAPI_SELF (-1)
-#endif
-
-static int dpapi_say(const char *fmt, ...) {
-   char buf[4096];
-   va_list ap;
-   ssize_t ret;
-
-   va_start(ap, fmt);
-   vsnprintf(buf, sizeof(buf), fmt, ap);
-   va_end(ap);
-   ret = write(__libpasstest_outputfd, buf, strlen(buf));
-   if (ret < 0 || (size_t)ret != strlen(buf)) {
-      return -1;
-   }
-   return 0;
-}
-
-static const char *dpapi_name(int fd, int bufnum) {
-   static struct {
-      char b[32];
-   } bufs[2];
-
-   if (fd == DPAPI_SELF) {
-      return "myself";
-   }
-
-   snprintf(bufs[bufnum].b, sizeof(bufs[bufnum].b), "fd%d", fd);
-   return bufs[bufnum].b;
-}
-
-int dpapi_init(void) {
-   if (LIBPASSTEST_CHECKINIT()) {
-      return -1;
-   }
-   if (dpapi_say("format provtrace v3\n")) {
-      return -1;
-   }
-   if (dpapi_say("create %d\n", dpapi_name(DPAPI_SELF, 0))) {
-      return -1;
-   }
-   return 0;
-}
-
-int dpapi_freeze(int fd) {
-   if (LIBPASSTEST_CHECKINIT()) {
-      return -1;
-   }
-   return dpapi_say("freeze %s\n", dpapi_name(fd, 0));
-}
-
-int dpapi_mkphony(int reference_fd) {
-   if (LIBPASSTEST_CHECKINIT()) {
-      return -1;
-   }
-
-   int fd = open(_PATH_DEVNULL, O_RDWR);
-   if (fd < 0) {
-      return -1;
-   }
-   if (dpapi_say("phony %s %s\n",
-		 dpapi_name(fd, 0), dpapi_name(reference_fd, 1))) {
-      close(fd);
-      return -1;
-   }
-
-   return fd;
-}
-
-ssize_t paread(int fd, void *data, size_t datalen, 
[…]

### 2013-11-13 passtools-20100524/libpasstest/libpasstest.h

@@ -1,8 +1,0 @@
-extern int __libpasstest_outputfd;
-extern int __libpasstest_initialized;
-
-#define LIBPASSTEST_CHECKINIT() \
-	__libpasstest_initialized ? 0 : __libpasstest_initialize()
-
-int __libpasstest_initialize(void);
-

### 2013-11-13 passtools-20100524/libpql/array.c

@@ -1,109 +1,0 @@
-/*
- * Copyright (c) 2009 David A. Holland.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the Author nor the names of any contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <string.h>
-
-#define ARRAYINLINE
-#include "array.h"
-
-struct array *array_create(struct pqlcontext *pql) {
-   struct array *a;
-   a = domalloc(pql, sizeof(*a));
-   if (a != NULL) {
-      array_init(a);
-   }
-   return a;
-}
-
[…]

### 2013-11-13 passtools-20100524/libpql/array.h

@@ -1,210 +1,0 @@
-/*
- * Copyright (c) 2009 David A. Holland.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the Author nor the names of any contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef ARRAY_H
-#define ARRAY_H
-
-#include "inlinedefs.h"
-#include "memdefs.h"
-
-#define ARRAYS_CHECKED
-
-#ifdef ARRAYS_CHECKED
-#include "utils.h"
-#define arrayassert PQLASSERT
-#else
-#define arrayassert(x) ((void)(x))
-#endif
-
-////////////////////////////////////////////////////////////
[…]

### 2013-11-13 passtools-20100524/libpql/baseopt.c

@@ -1,3412 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Simple optimizations.
- *
- * This module should contain only simple transforms that require
- * little or no analysis, context tracking, or non-local knowledge,
- * and do only one pass over the tree.
- *
- * It should also not contain any transforms that are subtle in either
[…]

### 2013-11-13 passtools-20100524/libpql/bindnil.c

@@ -1,366 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Arrange to have column-variable bindings that are skipped over
- * (e.g. because of being inside optional path sections) bound to nil.
- */
-
-#include "pttree.h"
-#include "passes.h"
-
-struct bindnil {
[…]

### 2013-11-13 passtools-20100524/libpql/columns.h

@@ -1,161 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef COLUMNS_H
-#define COLUMNS_H
-
-#include <stdbool.h>
-
-#include "array.h"
-struct pqlcontext;
-
-
-struct columnmanager;   /* Opaque. */
-struct colname;         /* Opaque. */
-struct colset;          /* Opaque. */
-struct coltree;         /* Opaque. */
-
-DECLARRAY(colname);
[…]

### 2013-11-13 passtools-20100524/libpql/datatype.c

@@ -1,993 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h> // for snprintf
-#include <string.h>
-
-#include "array.h"
-#include "pqlcontext.h"
-
-#define DATATYPE_INLINE
-#include "datatype.h"
-
-
-enum datatyperepresentations {
-   DT_BASE,			/* one of the base types in the manager */
-   DT_SET,			/* set(MEMBER) */
[…]

### 2013-11-13 passtools-20100524/libpql/datatype.h

@@ -1,190 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef DATATYPE_H
-#define DATATYPE_H
-
-#include <stdbool.h>
-#include "array.h"
-
-struct pqlcontext;
-
-
-struct datatypemanager; /* Opaque. */
-struct datatype; /* Opaque. */
-
-#ifndef DATATYPE_INLINE
-#define DATATYPE_INLINE INLINE
-#endif
[…]

### 2013-11-13 passtools-20100524/libpql/dequantify.c

@@ -1,233 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Convert quantifier expressions to map expressions.
- */
-
-#include "pttree.h"
-#include "passes.h"
-
-/*
- * Context. We don't need anything but the pqlcontext, but for now at least
[…]

### 2013-11-13 passtools-20100524/libpql/functions.c

@@ -1,154 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <string.h>
-
-#include "utils.h"
-#include "functions.h"
-
-/*
- * XXX this currently needs to be synchronized with ptlex.c
- */
-enum functions function_getbyname(const char *name, size_t namelen) {
-   if (namelen == 5 && !memcmp("ctime", name, namelen)) {
-      return F_CTIME;
[…]

### 2013-11-13 passtools-20100524/libpql/functions.h

@@ -1,111 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef FUNCTIONS_H
-#define FUNCTIONS_H
-
-#include <stddef.h>
-#include <stdbool.h>
-
[…]

### 2013-11-13 passtools-20100524/libpql/layout.c

@@ -1,1017 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdbool.h>
-#include <string.h>
-
-#include "utils.h"
-#include "layout.h"
-
-DEFARRAY(layout, /*noinline*/);
-
-static struct layout *mklayout(struct pqlcontext *pql, enum layouttypes type) {
-   struct layout *l;
-
-   (void)pql;
-   l = domalloc(pql, sizeof(*l));
-   l->type = type;
-   return l;
-}
-
[…]

### 2013-11-13 passtools-20100524/libpql/layout.h

@@ -1,116 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef LAYOUT_H
-#define LAYOUT_H
-
-
-#include "array.h"
-struct pqlcontext;
-
-DECLARRAY(layout);
-
-
-/*
- * Formatted string tree, for printing dumps.
- */
-
-enum layouttypes {
-   L_NEWLINE,
-   L_TEXT,
-   L_SEQUENCE,
-   L_LEFTALIGN,
[…]

### 2013-11-13 passtools-20100524/libpql/memdefs.c

@@ -1,364 +1,0 @@
-/*
- * Copyright 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdint.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <limits.h>
-#include <err.h>
-
-#include "pqlcontext.h"
-#include "utils.h"
-#include "memdefs.h"
-
-/*
- * Uncomment this to enable logging all malloc and free operations.
[…]

### 2013-11-13 passtools-20100524/libpql/memdefs.h

@@ -1,56 +1,0 @@
-/*
- * Copyright 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stddef.h>
-
-struct pqlcontext;
-
-
-// return address extraction
-
-#ifdef __GNUC__
-#define GETCALLER() __builtin_return_address(0)
-#endif
-
-#ifndef GETCALLER
-#define GETCALLER() (NULL)
-#endif
-
-// allocation wrappers
[…]

### 2013-11-13 passtools-20100524/libpql/movepaths.c

@@ -1,619 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include "pttree.h"
-#include "passes.h"
-#include "pqlcontext.h" // for complain()
-
-/*
- * Move paths to the from clause.
- *
- * This code does not attempt to deal with sets vs. non-sets or
[…]

### 2013-11-13 passtools-20100524/libpql/norenames.c

@@ -1,300 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Apply and remove all rename nodes.
- *
- * Because all our column names are distinguished (even if the user
- * recklessly gives some of them the same string names), there's no
- * danger of any nonsense about variable capture or conflating things
- * that are supposed to be different.
[…]

### 2013-11-13 passtools-20100524/libpql/passes.h

@@ -1,53 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stddef.h>
-
-struct pqlcontext;
-struct ptexpr;
-struct tcexpr;
-
-struct ptexpr *parse(struct pqlcontext *pql, const char *buf, size_t buflen);
-struct ptexpr *resolvevars(struct pqlcontext *pql, struct ptexpr *pe);
[…]

### 2013-11-13 passtools-20100524/libpql/pqlcontext.h

@@ -1,92 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdbool.h>
-
-#include "array.h"
-#include "utils.h"
-
-enum dumpstages {
-   DUMP_PARSER = 0,
-   DUMP_RESOLVE = 1,
-   DUMP_NORMALIZE = 2,
-   DUMP_UNIFY = 3,
[…]

### 2013-11-13 passtools-20100524/libpql/pqlvalue.c

@@ -1,1877 +1,0 @@
-/*
- * Copyright 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h> // for snprintf
-#include <string.h>
-
-#include "pql.h"
-#include "array.h"
-#include "layout.h"
-#include "datatype.h"
-#include "functions.h"
-
-#define VALUE_INLINE
-#include "pqlvalue.h"
-#include "valuerep.h"
-
-/*
[…]

### 2013-11-13 passtools-20100524/libpql/pqlvalue.h

@@ -1,155 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * value functions not published in pql.h
- */
-
-#ifndef PQLVALUE_H
-#define PQLVALUE_H
-
-#include "array.h"
-#include "pql.h"
-
-struct pqlcontext;
-struct pqlvalue;
-
-#ifndef VALUE_INLINE
-#define VALUE_INLINE INLINE
[…]

### 2013-11-13 passtools-20100524/libpql/ptparse.c

@@ -1,2303 +1,0 @@
-/*
- * Copyright 2008, 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <string.h>
-#include <stdlib.h>  // for strtod
-#include <limits.h>  // for INT_MAX
-#include <errno.h>   // for strtod
-
-#include "pqlvalue.h"
-#include "pttree.h"
-#include "ptprivate.h"
-#include "pqlcontext.h"
-#include "pql.h"
-
-
-/* type of input tokens */
[…]

### 2013-11-13 passtools-20100524/libpql/pttree.c

@@ -1,1210 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <string.h>
-
-#include "pqlvalue.h"
-#include "layout.h"
-#include "passes.h"	// where ptmemory_create currently lives
-#include "pqlcontext.h"
-
-#define PTTREE_INLINE
-#include "pttree.h"
-
-////////////////////////////////////////////////////////////
-
-/*
[…]

### 2013-11-13 passtools-20100524/libpql/tcalc.h

@@ -1,485 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef TCALC_H
-#define TCALC_H
-
-#include <stdbool.h>
-
-#include "array.h"
-#include "functions.h"
-
-struct pqlcontext;
-
-
-/*
- * (Nested) tuple calculus.
- */
-
-#ifndef TCALC_INLINE
[…]

### 2013-11-13 passtools-20100524/libpql/tuplify.c

@@ -1,1171 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Convert parse tree to tuple calculus.
- */
-
-#include <string.h>
-
-#include "utils.h"
-#include "pql.h"
-#include "columns.h"
-#include "pqlvalue.h"
-#include "pttree.h"
-#include "tcalc.h"
-#include "passes.h"
-
[…]

### 2013-11-13 passtools-20100524/libpql/typecheck.c

@@ -1,2401 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Check the types (and column names) for tuple calculus.
- */
-
-#include <stdarg.h>
-#include <stdio.h>
-
-#include "datatype.h"
-#include "columns.h"
-#include "pqlvalue.h"
-#include "tcalc.h"
-#include "passes.h"
-#include "pqlcontext.h"
-
-struct tc {
-   struct pqlcontext *pql;
-   bool failed;
-};
-
[…]

### 2013-11-13 passtools-20100524/libpql/typeinf.c

@@ -1,1355 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Figure out types (and column names) for tuple calculus.
- */
-
-#include "datatype.h"
-#include "columns.h"
-#include "pqlvalue.h"
-#include "tcalc.h"
-#include "passes.h"
-
-struct typeinf {
-   struct pqlcontext *pql;
-};
-
-////////////////////////////////////////////////////////////
-// context management
-
[…]

### 2013-11-13 passtools-20100524/libpql/unify.c

@@ -1,907 +1,0 @@
-/*
- * Copyright 2008, 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-/*
- * Path unification.
- * Mostly copied from the old engine.
- */
-
-#include <string.h>
-
-#include "utils.h"
-#include "pttree.h"
-#include "passes.h"
-
-/*
- * Operation context.
- */
-struct unify {
-   struct pqlcontext *pql;			/* Global context */
[…]

### 2013-11-13 passtools-20100524/libpql/utils.c

@@ -1,87 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-
-#include "pql.h"
-#include "memdefs.h"
-#include "utils.h"
-
-static unsigned evil = 0;
-static pqlassertionhandler handler;
-
-void pql_set_assertion_handler(pqlassertionhandler newhandler) {
[…]

### 2013-11-13 passtools-20100524/libpql/utils.h

@@ -1,51 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stddef.h>
-
-struct pqlcontext;
-
-
-#ifdef __GNUC__
-#define DEAD __attribute__((__noreturn__))
-#define PF(a,b) __attribute__((__format__(__printf__, a, b)))
-#else
-#define DEAD
-#define PF(a,b)
-#endif
-
-
-#define PQLASSERT(x) \
[…]

### 2013-11-13 passtools-20100524/libpql/valuerep.c

@@ -1,375 +1,0 @@
-/*
- * Copyright 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include "valuerep.h"
-
-DEFARRAY(tuplevalue, /*no inline*/);
-DEFARRAY(tupleslice, /*no inline*/);
-
-
-static struct tuplevalue *tuplevalue_create(struct pqlcontext *pql) {
-   struct tuplevalue *ret;
-
-   ret = domalloc(pql, sizeof(*ret));
-
[…]

### 2013-11-13 passtools-20100524/libpql/valuerep.h

@@ -1,613 +1,0 @@
-/*
- * Copyright 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef VALUEREP_H
-#define VALUEREP_H
-
-/*
- * Value representations.
- *
- * This file is only supposed to be included by valuerep*.c and pqlvalue.c.
- */
-
-#include <string.h>
-
-#include "pqlvalue.h"
-
[…]

### 2013-11-13 passtools-20100524/libpqlutil/compile.c

@@ -1,87 +1,0 @@
-/*
- * Copyright 2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/types.h>
-#include <sys/mman.h>
-#include <sys/stat.h>
-#include <string.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <err.h>
-
-#ifndef MAP_COPY
-#define MAP_COPY MAP_PRIVATE
-#endif
-
-#include "pql.h"
-#include "pqlutil.h"
-
[…]

### 2013-11-13 passtools-20100524/libpqlutil/interact.c

@@ -1,212 +1,0 @@
-/*
- * Copyright 2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <getopt.h>
-#include <assert.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
[…]

### 2013-11-13 passtools-20100524/libpqlutil/pqlpickle.c

@@ -1,942 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/types.h>
-#include <sys/utsname.h>
-#include <netinet/in.h>
-#include <stdlib.h>
-#include <string.h>
-#include <limits.h>
-#include <math.h>
[…]

### 2013-11-13 passtools-20100524/libpqlutil/printerrors.c

@@ -1,53 +1,0 @@
-/*
- * Copyright 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
-static const char *errorname;
-
-void pql_setprinterrorname(const char *name) {
-   errorname = name;
-}
-
-void pql_printerrors(struct pqlcontext *pql) {
-   unsigned n, i;
-   const char *msg;
-
[…]

### 2013-11-13 passtools-20100524/libpqlutil/printtrace.c

@@ -1,48 +1,0 @@
-/*
- * Copyright 2008, 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
-void pql_printtrace(struct pqlcontext *pql) {
-   unsigned num, i;
-   const char *line;
-
-   fprintf(stderr, "******** eval trace ********\n\n");
-
-   num = pqlcontext_getnumtracelines(pql);
[…]

### 2013-11-13 passtools-20100524/libpqlutil/ptrarray.h

@@ -1,110 +1,0 @@
-//
-// 	File:        ptrarray.h
-// 	Written:     16-Jul-95
-// 	Updated:     29-Oct-99
-//	Description: Array of pointers
-//
-// Copyright (c) 1995, 1999 David A. Holland.
-// All rights reserved.
-//
-// Redistribution and use in source and binary forms, with or without
-// modification, are permitted provided that the following conditions
-// are met:
-// 1. Redistributions of source code must retain the above copyright
-//    notice, this list of conditions and the following disclaimer.
-// 2. Redistributions in binary form must reproduce the above copyright
-//    notice, this list of conditions and the following disclaimer in the
-//    documentation and/or other materials provided with the distribution.
-// 3. Neither the name of the Author nor the names of any contributors
-//    may be used to endorse or promote products derived from this software
-//    without specific prior written permission.
-//
-// THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
-// ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
-// IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
-// ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
-// FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
-// DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
-// OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
-// HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
-// LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
-// OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
-// SUCH DAMAGE.
-//
-//   $Id: ptrarray.h,v 1.1.1.1 2008/11/06 08:55:33 dholland Exp $
-//
-
-#ifndef PTRARRAY_H
-#define PTRARRAY_H
-
[…]

### 2013-11-13 passtools-20100524/libpqlutil/tdb.c

@@ -1,273 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include <assert.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
-////////////////////////////////////////////////////////////
-// malloc
-
-static void *tdb_malloc(size_t size) {
-   void *ret;
-
-   ret = malloc(size);
-   if (ret == NULL) {
-      fprintf(stderr, "Out of memory\n");
-      exit(1);
-   }
-   return ret;
-}
-
-////////////////////////////////////////////////////////////
-// array
-
-/*
- * Probably this should use libpql's typesafe arrays, but sharing them
- * is a pain and we don't really need the full generality. Also I
- * don't want to either pollute the linker namespace or require
- * setting up all the abi.txt crap libpql needs.
- */
-
-struct tdbarray {
-   unsigned num, max;
-   void **v;
-};
-
-static inline void tdbarray_init(struct tdbarray *a) {
-   a->num = a->max = 0;
-   a->v = NULL;
-}
-
-static inline void tdbarray_cleanup(struct tdbarray *a) {
-   assert(a->num == 0);
-   free(a->v);
-}
-
-static inline unsigned tdbarray_num(const struct tdbarray *a) {
-   return a->num;
-}
-
-static inline void *tdbarray_get(const struct tdbarray *a, unsigned index) {
-   assert(index < a->num);
-   return a->v[index];
-}
-
-static inline void tdbarray_set(const struct tdbarray *a, unsigned index,
-				void *val) {
-   assert(index < a->num);
-   a->v[index] = val;
-}
-
-static void tdbarray_setsize(struct tdbarray *a, unsigned num) {
-   unsigned newmax;
-   void **newptr;
-
-   if (num > a->max) {
-      newmax = a->max;
-      while (num > newmax) {
-	 newmax = newmax ? newmax*2 : 4;
-      }
-      newptr = realloc(a->v, newmax*sizeof(*a->v));
-      if (newptr == NULL) {
-	 fprintf(stderr, "Out of memory\n");
-	 exit(1);
-      }
-      a->v = newptr;
-      a->max = newmax;
-   }
-   a->num = num;
-}
-
-static inline void tdbarray_add(struct tdbarray *a, void *val,
-				unsigned *index_ret) {
[…]

### 2013-11-13 passtools-20100524/libtwig/twig.c

@@ -1,236 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <stdio.h>
-#include <string.h>
-#include <errno.h>
-#include <assert.h>
-#include "twig.h"
-
-static void twig_print_rec_header(const struct twig_rec_header *header);
-static void twig_print_rec_begin(const struct twig_rec_begin *begin);
-static void twig_print_rec_end(const struct twig_rec_end *end);
[…]

### 2013-11-13 passtools-20100524/libwdb/provdb.c

@@ -1,372 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <string.h>
-#include <assert.h>
-#include <db.h>
-
-#include "schema.h"
-
-#ifdef __GNUC__
-#define UNUSED __attribute__((__unused__))
-#else
-#define UNUSED
-#endif
-
-#ifndef MIN
[…]

### 2013-11-13 passtools-20100524/libwdb/wdb.c

@@ -1,831 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <stdarg.h>
-#include <errno.h>
-#include <string.h>
-#include <assert.h>
-#include <unistd.h>
-#include <sys/stat.h>
-#include <sys/types.h>
-
-#include "debug.h"
-
-#include "wdb.h"
-#include "dbs.h"
-
-//
-// File and name constants
[…]

### 2013-11-13 passtools-20100524/mount.lasagna/mount.lasagna.sh

@@ -1,59 +1,0 @@
-#! /bin/sh
-waldo_location=$PREFIX/libexec/passtools/waldo
-sage_location=$PREFIX/bin/sage
-
-if [ -d $1 ]; then
-    # underlying fs is already mounted and we're being asked to mount
-    # just the stack.
-    echo "$0: assuming $1 is an already-mounted underlying volume" 1>&2
-    echo "$0: and not starting daemons" 1>&2
-    DEV="$1"
-    pid=''
-else
-    DEV="$2"
-    mount -t ext3 $3 $4 $1 $2
-    result=$?
-    if [ "$result" -ne "0" ]; then
-	echo "mount.lasagna: ext3 mount failed with $result"
-	exit $result
-    fi
-
-    cd $DEV/.lasagna_stuff
-    if [ "`pwd`" != "$2/.lasagna_stuff" ]; then
-	echo "mount.lasagna: could not cd into $DEV/.lasagna_stuff"
-	exit -1
-    fi
-
-    $waldo_location &
-    pid=$!
-    filename=`echo waldo$DEV | sed 's/\//_/g'`
-    fullpath=/var/run/$filename
-    echo $pid > $fullpath
-
-    $sage_location -p db -s &
-    spid=$!
-    filename=`echo sage$DEV | sed 's/\//_/g'`
-    fullpath=/var/run/$filename.pid
-    echo $spid > $fullpath
-fi
-
-if ! cat /proc/filesystems | grep -w lasagna >/dev/null 2>&1; then
-    echo "mount.lasagna: loading lasagna module explicitly just in case" 1>&2
-    modprobe lasagna || true
-fi
-
-OPTIONS=`echo "$4" | sed 's/noauto,//;s/,noauto//;s/^noauto$//'`
-$ROOTPREFIX/sbin/mount_lasagna_helper $DEV $2 $3 "$OPTIONS"
-result=$?
-if [ "$result" -ne "0" ]; then
-	echo "mount.lasagna: lasagna mount failed with $result"
-	if [ "x$pid" != x ]; then
-	    kill -USR1 $pid
-	    kill $spid
-	    cd /
-	    sleep 1
-	    umount $2
-	fi
-	exit $result
-fi
-

### 2013-11-13 passtools-20100524/mount_lasagna_helper/mount_lasagna_helper.c

@@ -1,69 +1,0 @@
-#include <errno.h>
-#include <mntent.h>
-#include <stdio.h>
-#include <string.h>
-#include <sys/mount.h>
-
-int
-main(int argc, char **argv)
-{
-	FILE *fh;
-	struct mntent *mnt_exist;
-	struct mntent m;
-
-	if (argc != 5) {
-		fprintf(stderr, "Usage: %s dev mountpoint . args\n",
-			argv[0]);
-		return EINVAL;
-	}
-
-	fh = setmntent("/etc/mtab","r");
-	if (fh == NULL) {
-		perror("mount.lasagna");
-		return EACCES;
-	}
-
-	while ((mnt_exist = getmntent(fh)) != NULL) {
-		if (!strcmp(mnt_exist->mnt_dir, argv[2]) && 
-			!strcmp(mnt_exist->mnt_type, "lasagna")) {
-			endmntent(fh);
-			return EBUSY;
-		}
-	}
-	endmntent(fh);
-
-	if (strlen(argv[4]) == 2) {
-		argv[4][0] = '\0';
-	} else {
-		memmove(argv[4], argv[4]+3, strlen(argv[4])-2);
-	}
-
-#if 0
-	int i;
-	for (i = 0; i <= argc; ++i) {
-		printf("mount.lasagna: --%s--\n", argv[i]);
-	}
-#endif
-
-	if (mount(argv[1], argv[2], "lasagna", 0, argv[4])) {
-		int serrno = errno;
-		perror("mount.lasagna");
-		return serrno;
-	} else {
-		fh = setmntent("/etc/mtab","a");
-		m.mnt_fsname = argv[1];
-		m.mnt_dir = argv[2];
-		m.mnt_type = "lasagna";
-		m.mnt_opts = MNTOPT_DEFAULTS;
-		m.mnt_freq = 0;
-		m.mnt_passno = 0;
-		if (addmntent(fh,&m) != 0) {
-			perror("mount.lasagna");
-			endmntent(fh);
-			umount(argv[1]);
-			return EACCES;
-		}
-		endmntent(fh);
-		return 0;
-	}
-}

### 2013-11-13 passtools-20100524/osage/ast.cc

@@ -1,300 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <string.h>
-#include <assert.h>
-#include "dump.h"
-#include "ast.h"
-
-var *var_create(void) {
-   static int nextvar;
-
-   var *v = new var;
-   v->id = nextvar++;
-   v->val = NULL;
-   return v;
-}
-
[…]

### 2013-11-13 passtools-20100524/osage/ast.h

@@ -1,107 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include "operators.h"
-#include "value.h"
-
-enum exprtypes {
-   ET_FOR,
-   ET_LET,
-   ET_LAMBDA,
-   ET_COND,
-   ET_OP,
-   ET_TUPLE,
-   ET_REF,
-   ET_VAL,
-};
-
-struct var {
[…]

### 2013-11-13 passtools-20100524/osage/builtins.cc

@@ -1,177 +1,0 @@
-/*
- * Copyright 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <string.h>
-#include <time.h>
-#include "utils.h"
-#include "dump.h"
-#include "ast.h"
-#include "value.h"
-#include "builtins.h"
-#include "main.h"
-
-struct builtin {
-   const char *name;
-   value *(*func)(value *args);
-};
-
-////////////////////////////////////////////////////////////
-
-static int alldigits(const char *s) {
[…]

### 2013-11-13 passtools-20100524/osage/builtins.h

@@ -1,38 +1,0 @@
-/*
- * Copyright 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-
-struct builtin;
-
-const builtin *builtin_lookup(const char *name);
-value *builtin_exec(const builtin *b, value *args);
-
-int builtin_cmp(const builtin *a, const builtin *b);
-const char *builtin_name(const builtin *b);
-void builtin_dump(const builtin *b);

### 2013-11-13 passtools-20100524/osage/dbops.cc

@@ -1,785 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include "primarray.h"
-#include "ast.h"
-#include "dbops.h"
-#include "main.h" // for eval
-
-//#define COMPILED_IN_TEST_DATA
-
[…]

### 2013-11-13 passtools-20100524/osage/dbops.h

@@ -1,49 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-value *db_evalallprov(var *ref, expr *where, expr *in);
-
-// get from PROV table
-value *db_get_attr(uint64_t pnode, const char *attribute);
-value *db_get_allattr(uint64_t pnode);
-
-// get from NAME index
-value *db_get_name(const char *name);
-
-// get from I2P index
-value *db_get_i2p(long val);
-
-// get from ARGV index
[…]

### 2013-11-13 passtools-20100524/osage/dump.cc

@@ -1,110 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdarg.h>
-#include <string.h>
-#include <assert.h>
-#include "primarray.h"
-#include "dump.h"
-
-struct dumpstate {
-   bool atbol;
-   int indent;
-};
-
-static primarray<dumpstate> prevdumps;
-static dumpstate curdump;
[…]

### 2013-11-13 passtools-20100524/osage/main.h

@@ -1,44 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-struct expr;
-struct value;
-
-extern int g_dodumps;
-extern int g_dotrace;
-
-expr *compile_file(const char *path);
-expr *compile_string(const char *string);
-
-expr *baseopt(expr *e);
-expr *indexify(expr *e);
-value *eval(expr *e);
-value *eval_path(expr *e);
-
-void output_result(value *v);

### 2013-11-13 passtools-20100524/osage/operators.h

@@ -1,99 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef OPERATORS_H
-#define OPERATORS_H
-
-// operators
-enum ops {
-   // reserved
-   OP_NOP,
-
[…]

### 2013-11-13 passtools-20100524/osage/output.cc

@@ -1,287 +1,0 @@
-/*
- * Copyright 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <ctype.h> // XXX shouldn't use
-#include "ast.h"
-#include "value.h"
-#include "builtins.h"
-#include "dbops.h"
-#include "main.h"
-
-static int g_long_pnodes = 1;
-
-////////////////////////////////////////////////////////////
-
-static expr *make_eval_all_code(void) {
-   // for x = ** in x, without optimizing it
-   var *x = var_create();
[…]

### 2013-11-13 passtools-20100524/osage/parse.cc

@@ -1,2238 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdarg.h>
-#include <stdlib.h>
-#include "ptnode.h"
-
-/*
- * AnaGram, A System for Syntax Directed Programming
[…]

### 2013-11-13 passtools-20100524/osage/parse.h

@@ -1,77 +1,0 @@
-#ifndef PARSE_H_1182556825
-#define PARSE_H_1182556825
-
-typedef union {
-  long alignment;
-  char ag_vt_2[sizeof(int)];
-  char ag_vt_4[sizeof(ptnode *)];
-  char ag_vt_5[sizeof(char *)];
-} parse_vs_type;
-
-typedef enum {
-  parse_query_token = 6, parse_ws_token, parse_expr_token = 9,
-  parse_eof_token = 12, parse_bind_expr_token, parse_lambda_expr_token,
-  parse_filter_expr_token, parse_for_head_token, parse_let_head_token,
-  parse_lambda_head_token, parse_binding_name_token = 21,
-  parse_ident_token = 23, parse_tuple_expr_token = 25,
-  parse_real_tuple_expr_token, parse_path_expr_token,
-  parse_qualified_path_element_token = 29, parse_path_element_token = 33,
-  parse_value_expr_token = 35, parse_logical_expr_token = 37,
-  parse_equality_expr_token, parse_comparison_expr_token = 41,
-  parse_cons_token = 44, parse_term_token = 51,
-  parse_range_start_token = 53, parse_number_token, parse_factor_token = 56,
-  parse_suffix_expr_token = 60, parse_prefix_expr_token = 64,
-  parse_base_expr_token = 68, parse_string_token,
-  parse_string_body_token = 79, parse_string_char_token,
-  parse_hex_digit_token = 84, parse_letter_token, parse_digit_token = 87,
-  parse_ws_char_token = 90
-} parse_token_type;
-
-typedef struct parse_pcb_struct{
-  parse_token_type token_number, reduction_token, error_frame_token;
-  int input_code;
-  int input_value;
-  int line, column;
-  int ssx, sn, error_frame_ssx;
-  int drt, dssx, dsn;
-  int ss[128];
-  parse_vs_type vs[128];
-  int ag_ap;
-  const char *error_message;
-  char read_flag;
-  char exit_flag;
-  int bts[128], btsx;
-  int (* const  *gt_procs)(void);
-  int (* const *r_procs)(void);
-  int (* const *s_procs)(void);
-  int lab[9], rx, fx;
-  const unsigned char *key_sp;
-  int save_index, key_state;
[…]

### 2013-11-13 passtools-20100524/osage/path.cc

@@ -1,711 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h> // XXX temporary
-#include <string.h>
-#include "utils.h"
-#include "dump.h"
-#include "ast.h"
-#include "dbops.h"
-#include "main.h"
-
-/*
- * Eval a path expression.
- *
- * (1) Collect the whole path.
- * (2) Pick which end to start from.
- * (3) Compile it into regexp states.
- * (4) Eval at one end to get starting hypotheses.
[…]

### 2013-11-13 passtools-20100524/osage/value.cc

@@ -1,490 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <string.h>
-#include <assert.h>
-#include "utils.h"
-#include "dump.h"
-#include "value.h"
-#include "ast.h" // for expr_clone
-#include "builtins.h"
-
-static value *mkvalue(valuetypes type) {
-   value *v = new value;
-   v->type = type;
[…]

### 2013-11-13 passtools-20100524/osage/value.h

@@ -1,123 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef VALUE_H
-#define VALUE_H
-
-#include <stdint.h>
-#include "ptrarray.h"
-
-struct expr;
-struct builtin;
-
-
-enum whichindex {
-   INDEX_ID,
-   INDEX_I2P,
-   INDEX_NAME,
-   INDEX_ARGV,
-};
-
-enum valuetypes {
[…]

### 2013-11-13 passtools-20100524/sage/backend.cc

@@ -1,1109 +1,0 @@
-/*
- * Copyright 2008, 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <string.h>
-#include <stdlib.h>
-#include <errno.h>
-#include <assert.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
-#include "wdb.h"
-
-#include "primarray.h"
-#include "backend.h"
-
-/*
- * Database numbers
- */
-#define DBNUM_TEMP  0
-#define DBNUM_MAGIC 1
[…]

### 2013-11-13 passtools-20100524/sage/backend.h

@@ -1,34 +1,0 @@
-/*
- * Copyright 2008, 2009
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-int backend_init(const char *dbpath);
-int backend_shutdown(void);
-
-extern const struct pqlbackend_ops myops;
-

### 2013-11-13 passtools-20100524/sage/interact.cc

@@ -1,204 +1,0 @@
-/*
- * Copyright 2008, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <assert.h>
-
-#include "interact.h"
-#include "user.h"
-#include "query.h"
-
[…]

### 2013-11-13 passtools-20100524/sage/interact.h

@@ -1,41 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-struct result;
-
-void interact(bool forceprompt);
-void print_result(const struct result *res);
-
-#ifdef __cplusplus
-}
-#endif

### 2013-11-13 passtools-20100524/sage/local.h

@@ -1,39 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-struct result;
-
-void local_init(const char *dbpath);
-void local_shutdown(void);
-
-void local_dodumps(bool val);
-void local_dotrace(bool val);
-
-void local_submit_file(const char *file, struct result *res);
-void local_submit_string(const char *cmd, struct result *res);

### 2013-11-13 passtools-20100524/sage/main.cc

@@ -1,108 +1,0 @@
-/*
- * Copyright 2008, 2009, 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <getopt.h>
-
-#include "interact.h"
-#include "user.h"
-#include "server.h"
-#include "query.h"
-
-
-static void usage(const char *av0) {
-   fprintf(stderr, "Usage:\n");
[…]

### 2013-11-13 passtools-20100524/sage/pathnames.h

@@ -1,36 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <paths.h>
-
-#ifndef _PATH_VARRUN
-#define _PATH_VARRUN "/var/run/"
-#endif
-
-#define PATH_LINUX_PROC_MOUNTS "/proc/mounts"

### 2013-11-13 passtools-20100524/sage/query.h

@@ -1,37 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-void query_init(const char *dbpath, const char *socketpath);
-void query_shutdown(void);
-
-void query_dodumps(bool);
-void query_dotrace(bool);
-
-void query_submit_file(const char *file, struct result *res);
-void query_submit_string(const char *cmd, struct result *res);

### 2013-11-13 passtools-20100524/sage/remote.cc

@@ -1,382 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <sys/un.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdarg.h>
-#include <string.h>
-#include <unistd.h>
-#include <assert.h>
-#include <errno.h>
-#include <err.h>
-
-#include "pql.h"
-#include "pqlutil.h"
-
-#include "utils.h"
-#include "socketpath.h"
-#include "result.h"
[…]

### 2013-11-13 passtools-20100524/sage/remote.h

@@ -1,39 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-struct result;
-
-int remote_init(const char *socketpath);
-int remote_shutdown(void);
-
-void remote_dodumps(bool val);
-void remote_dotrace(bool val);
-
-void remote_submit_file(const char *file, struct result *res);
-void remote_submit_string(const char *string, struct result *res);

### 2013-11-13 passtools-20100524/sage/socketpath.cc

@@ -1,138 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <sys/un.h>
-#include <sys/statvfs.h>
-#include <stdio.h>
-#include <string.h>
-#include <limits.h>
-#include <err.h>
-
-#include "pathnames.h"
-#include "socketpath.h"
-
-/*
[…]

### 2013-11-13 passtools-20100524/sage/socketpath.h

@@ -1,34 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#define PROTOCOL_VERSION        0
-#define MAXIMUM_LINE_LENGTH     4096
-#define MAXIMUM_LINE_COUNT      1000
-
-void getsocketaddr(const char *socketpath, struct sockaddr_un *sun);

### 2013-11-13 passtools-20100524/sage/user.h

@@ -1,33 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-struct result;
-
-int user_submit_file(const char *file);
-int user_submit_string(const char *string);

### 2013-11-13 passtools-20100524/sage/utils.h

@@ -1,30 +1,0 @@
-/*
- * Copyright 2010
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-char *xstrdup(const char *);

### 2013-11-13 passtools-20100524/umount.lasagna/umount.lasagna.sh

@@ -1,54 +1,0 @@
-#! /bin/sh
-VOL="$1"
-VOLSTRING=`echo "$VOL" | sed 's/\//_/g'`
-
-# usage:
-#   $1: name
-#   $2: pidfile
-#   $3: signal
-killit() {
-    local name="$1"
-    local pidfile="$2"
-    local signal="$3"
-    local pid
-
-    #echo "$0: $name: pidfile $pidfile"
-
-    if [ -e "$pidfile" ]; then
-	pid=`cat "$pidfile"`
-	#echo "$0: $name: pid $pid"
-	if kill -0 "$pid" >/dev/null 2>&1; then
-	    #echo "$0: killing $name..."
-	    kill "$signal" "$pid"
-	    # wait a moment
-	    echo | cat | cat >/dev/null
-	    while kill -0 "$pid" >/dev/null 2>&1; do
-		echo "$0: waiting for $name to exit"
-		sleep 1
-	    done
-	    echo "$0: $name: stopped"
-	else
-	    echo "$0: $name: pid $pid: not running" 1>&2
-	    return 0
-	fi
-	rm -f "$pidfile"
-    fi
-    return 0
-}
-
-killit "waldo" /var/run/waldo"$VOLSTRING" -USR1
-killit "sage" /var/run/sage"$VOLSTRING".pid -TERM
-
-$ROOTPREFIX/sbin/umount_lasagna_helper $1
-result=$?
-if [ "$result" -ne "0" ]; then
-	echo "umount.lasagna: lasagna umount failed with $result"
-	exit $result
-fi
-umount $1
-result=$?
-if [ "$result" -ne "0" ]; then
-	echo "umount.lasagna: ext3 umount failed with $result"
-	exit $result
-fi
-

### 2013-11-13 passtools-20100524/umount_lasagna_helper/umount_lasagna_helper.c

@@ -1,46 +1,0 @@
-#include <errno.h>
-#include <mntent.h>
-#include <stdio.h>
-#include <string.h>
-#include <sys/mount.h>
-
-int
-main(int argc, char **argv)
-{
-	int retval;
-	FILE *fh, *fh_new;
-	struct mntent *mnt_exist;
-
-	if ((retval = umount(argv[1])) != 0) {
-		perror("umount.lasagna");
-		return retval;
-	}
-
-	fh = setmntent("/etc/mtab","r");
-	if (fh == NULL) {
-		perror("umount.lasagna");
-		return EACCES;
-	}
-	fh_new = setmntent("/etc/mtab.pass.new", "w");
-	if (fh_new == NULL) {
-		endmntent(fh);
-		perror("umount.lasagna");
-		return EACCES;
-	}
-
-	while ((mnt_exist = getmntent(fh)) != NULL) {
-		if (strcmp(mnt_exist->mnt_dir, argv[1]) || 
-			strcmp(mnt_exist->mnt_type, "lasagna")) {
-			
-			addmntent(fh_new, mnt_exist);
-		}
-	}
-	endmntent(fh_new);
-	endmntent(fh);
-	if ((retval = rename("/etc/mtab.pass.new", "/etc/mtab")) != 0) {
-		perror("umount.lasagna");
-		return retval;
-	}
-	
-	return 0;
-}

### 2013-11-13 passtools-20100524/waldo/cleanpages.cc

@@ -1,105 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#include <stdlib.h>
-#include <string.h>
-#include <map>
-
-#include "cleanpages.h"
-
-using namespace std;
-
-/** Our idea of a "page" */
-struct cleanpage {
-   ino_t    inode;
-   uint64_t off;
-};
-
-/** Compare class for cleanpage struct (used by map) */
-struct classcomp {
-   bool operator() (const cleanpage& lhs, const cleanpage& rhs) const
-   {
[…]

### 2013-11-13 passtools-20100524/waldo/cleanpages.h

@@ -1,46 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#ifndef _CLEANPAGES_H
-#define _CLEANPAGES_H
-
-#include <stdint.h>
-#include <sys/types.h>
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-   int cleanpages_lookup(ino_t inode, uint64_t off);
[…]

### 2013-11-13 passtools-20100524/waldo/i2n.c

@@ -1,218 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#include <stdlib.h>
-#include <string.h>
-#include <ftw.h>
-#include <assert.h>
-#include <db.h>
-
-#include "wdb.h"
-#include "i2n.h"
-
-#define NOPENFD 20
-
-extern int inode_cmp(DB *db, const DBT *dbt1, const DBT *dbt2);
-
-static struct waldo_db *i2ndb = NULL;
-
-static struct waldo_db tmpdb =
[…]

### 2013-11-13 passtools-20100524/waldo/log.h

@@ -1,51 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#ifndef LOG_H
-#define LOG_H
-
-#include <stdint.h>
-
-enum log_state { LOG_STATE_KERNEL, LOG_STATE_ACTIVE, LOG_STATE_BACKUP };
-
-int log_startup(const char *logpath);
-int log_shutdown(void);
-int log_find_files(const char *logpath);
-int log_last_lognum(uint64_t *num);
-int log_get_number(const char *name, uint64_t *number);
[…]

### 2013-11-13 passtools-20100524/waldo/process.c

@@ -1,1126 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <signal.h>
-#include <errno.h>
-#include <assert.h>
-
-#include <sys/types.h>
-#include <sys/stat.h>
-#include <unistd.h>
-
-#include "debug.h"
-
-#include "schema.h"
-#include "twig_file.h"
-#include "wdb.h"
[…]

### 2013-11-13 passtools-20100524/waldo/process.h

@@ -1,46 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#ifndef _PROCESS_H
-#define _PROCESS_H
-
-#include <db.h>
-#include "twig.h"
-#include "log.h"
-
-int waldo_process_fullpath(const char *fullpath, DB_TXN *txn);
-int waldo_process_file(const char *dir, const char *filename, DB_TXN *txn);
[…]

### 2013-11-13 passtools-20100524/waldo/ptr_array.h

@@ -1,18 +1,0 @@
-#ifndef PTR_ARRAY_H
-#define PTR_ARRAY_H
-
-#include <stdint.h>
-
-int ptr_array_add(void ***array,
-                  uint32_t *count,
-                  const void *ptr);
-
-int ptr_array_add_nodup(void ***array,
-                        uint32_t *count,
-                        const void *ptr);
-
-int ptr_array_contains(const void **array,
-                       const uint32_t count,
-                       const void *ptr);
-
-#endif /* PTR_ARRAY_H */

### 2013-11-13 passtools-20100524/waldo/recover.h

@@ -1,41 +1,0 @@
-/*
- * Copyright 2006-2008
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-#ifndef RECOVER_H
-#define RECOVER_H
-
-#include <db.h>
-#include <stdint.h>
-
-#define MD5_BYTES 16
-
-int waldo_recover(const char *fsroot, const char *logpath, DB_TXN *txn);
-
-int compute_md5(const uint8_t *buf, const int len, uint8_t *md5);
-
-#endif /* RECOVER_H */

### 2013-11-13 passtools-20100524/waldo/waldo.c

@@ -1,234 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <getopt.h>
-#include <unistd.h>
-#include <sys/stat.h>
-#include <sys/types.h>
-#include <sys/time.h>
-#include <assert.h>
-#include <errno.h>
-#include <limits.h>
-#include <signal.h>
[…]

### 2013-11-13 passtools-20100524/wdb_s3dump/wdb_s3dump.c

@@ -1,648 +1,0 @@
-/*
- * Copyright 2006, 2007
- *	The President and Fellows of Harvard College.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE UNIVERSITY AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE UNIVERSITY OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <stdlib.h>
-#include <string.h>
-#include <stdio.h>
-#include <ctype.h>
-#include <getopt.h>
-#include <assert.h>
-
-#include <db.h>
-
-#include "twig.h"
-#include "schema.h"
-#include "wdb.h"
-
[…]

### 2013-11-13 writeup/paper.tex
part of intro written

@@
 
-\title{TODO: Intrusion Detection through Provenance}
+\title{{\bf TODO: Intrusion Detection through Provenance}}
 \author{
@@
     Harvard University\\
-    kennyyu@college.harvard.edu
+    \href{mailto:kennyyu@college.harvard.edu}{\texttt{kennyyu@college.harvard.edu}}
   \and
@@
     Harvard University\\
-    rjaquino@college.harvard.edu
+    \href{mailto:rjaquino@college.harvard.edu}{\texttt{rjaquino@college.harvard.edu}}
 }
-\date{November 12, 2013}
+\date{CS261, Fall 2013}
 
@@
 % ABSTRACT
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 % INTRODUCTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
 \section{Introduction}
-TODO
+Provenance is metadata that tracks the history of all changes to files.
+In provenance-aware storage systems like PASS \cite{pass} and PASSv2 \cite{passv2}, the file system automatically tracks all dependencies whenever a file is created, modified, or deleted.
+These dependencies include the command that was executed to modify the file, the environment of the command, and all input files.
+The generated provenance forms a directed acyclic graph of typed nodes (e.g.
+files, processes, pipes) with properties (e.g.
+name, execution time, pid) and typed edges (e.g.
+forked, input, versioning).
+
+Applications that require provenance collection typically place great emphasis on data integrity.
+Given this priority, a natural goal of provenance systems is to detect intrusions on the system.
+Somayaji et.
+al.
+have developed techniques for detecting intrusions based on system call sequences, and counteract these by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
+These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
+
+Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies, it seems plausible to detect intrusions based on provenance data collected by PASS.
+
+====
+
+Several existing systems 
+intrusion detection systems using system calls
+
+other systems to collect provenance to detect intrusions
+
+our approach: analyze properties of dags, histograms?
 
@@
 % DESIGN
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 % EVALUATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 % CONCLUSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 % BIBLIOGRAPHY
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 

### 2013-11-13 writeup/paper.tex
finished introduction

@@
 
+To make sense of the large amounts of data, Macko et.
+al.
+have developed Orbiter, a tool to visualize provenance graphs with semantic zoom \cite{orbiter}.
+Margo et.
+al.
+have developed techniques to analyze large provenance graphs and to extract useful information from these graphs, including semantic file attributes \cite{fileattributes}.
+Using simple provenance statistics on nodes (e.g.
+neighbors, file count, process count, edge count) and features from \texttt{stat}, they determine with high accuracy the type of files (e.g.
+source files, configuration files).
+Furthermore, they have identified key properties of provenance graphs and developed new centrality metrics to perform local clustering of graphs, separating the huge provenance graphs into smaller semantic tasks or workloads \cite{clustering}.
+
 Applications that require provenance collection typically place great emphasis on data integrity.
@@
 Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies, it seems plausible to detect intrusions based on provenance data collected by PASS.
+Tariq et.
+al.
+generalize Somayaji's sequences of system calls to sequences of provenance system events \cite{correlated-anomalies} to develop an intrusion detection system (IDS) in a distributed system.
+Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of provenance events--and then use these bit vectors to correlate anomalies across multiple hosts.
+King et.
+al.
+have developed Backtrack \cite{backtrack}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
+Given an intrusion, Backtrack builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtrack with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
+However, their system does not find a way of detecting an intrusion.
+Lei et.
+al.
+build similar process-file dependency trees, which they call access trees, and define a compatibility function to compare access trees with the goal of predicting future file accesses based on the current access tree pattern \cite{fileprefetch}.
 
-====
+Despite the existing work to detect intrusions and anomalies, Cao et.
+al.
+argue that any intrusion detection system cannot always distinguish good behaviors from bad behaviors because users behave too randomly to have an accurate model of the user's behavior \cite{fuzzy}.
+They propose their fuzzy anomaly detection approach to extend the training models in existing intrusion detection systems to account for the inherent fuzzy nature of intrusion detection.
+When training an IDS, instead of assigning hard 0s or 1s for features, their approach provides a way to reliably map these features to weights in the interval $[0,1]$.
 
-Several existing systems 
-intrusion detection systems using system calls
+In this paper, we present a technique to analyze existing provenance data to detect if an intrusion occurred.
 
-other systems to collect provenance to detect intrusions
-
-our approach: analyze properties of dags, histograms?
+{\bf TODO: TALK ABOUT OUR TECHNIQUES.} We build on TODO's work and extend ... blah blah blah ... combine techniques together ... blah blah. The contributions of this paper are:
+\begin{enumerate}
+\item TODO: the techniques we use to analyze provenance data
+\item TODO: types of intrusions that this system can detect, and the accuracy in detecting intrusions
+\item TODO: future work for using provenance data to detect intrusions
+\end{enumerate}
 
@@
 
+\bibitem{fuzzy}
+\textsc{Cao, D., Qiu, M., Chen, Z., Hu, F., Zhu, Y., and Wang, B.} Intelligent Fuzzy Anomaly Detection of Malicious Software.
+In {\em Internal Journal of Advanced Intelligence}, vol.
+4, no.
+1, pp 69-86 (December 2012).
+
 \bibitem{somayaji-recent}
 \textsc{Inoue, H. and Somayaji, A.} Lookahead Pairs and Full Sequences: A Tale of Two Anomaly Detection Methods. In {\em 2nd Annual Symposium on Information Assurance} (June 2007).
+
+\bibitem{backtrack}
+\textsc{King, S. T. and Chen, P. M.} Backtracking Intrusions. In {\em SOSP'03 Proceedings of the nineteenth ACM symposium on Operating systems principles} (December 2003).
+
+\bibitem{multihost}
+\textsc{King, S.
+T., Mao Z.
+M., Lucchetti, D.
+G., and Chen, P.
+M.} Enriching intrusion alerts through multi-host causality.
+In {\em Proceedings of the 2005 Network and Distributed System Security Symposium} (February 2005).
+
+\bibitem{fileprefetch}
+\textsc{Lei, H. and Duchamp, D.} An Analytical Approach to File Prefetching. In {\em Proceedings of the USENIX 1997 Annual Technical Conference} (January 1997).
 
@@
 
+\bibitem{correlated-anomalies}
+\textsc{Tariq, D., Baig, B., Gehani, A., Mahmood, S., Tahir, R., Aqil, A., and Zaffar, F.} Identifying the provenance of correlated anomalies.
+In {\em SAC'11 Proceedings of the 2011 ACM Symposium on Applied Computing} (March 2011).
+
 \end{thebibliography}

### 2013-11-13 writeup/paper.tex
add design and implementation ideas and todos in other sections

@@
 
-\section{Design}
-TODO
+\section{Design and Implementation}
+TODO: ideas:
+\begin{enumerate}
+\item somayaji: sequence of system calls
+\item semantic file attributes: determine types of files. use simple statistics on files.
+\item histograms: use simple statistics and build histogram of file attributes on a process name/file
+\item local clustering: use centrality metrics
+\item fuzzy rules
+\item hash k-tuples of provenance events and compare bit vectors
+\item backwards and forwards causality graphs
+\item access compatibility $\frac{1}{2}(C_d + C_e)$
+\item intersect good DAG and bad DAG and look at intersection of DAG
+\end{enumerate}
 
@@
 \section{Evaluation}
-TODO
+
+\subsection{Evaluation Methodology}
+TODO:
+types of exploits we ran.
+how we ran it.
+how we collected the data.
+the software we used.
+
+\subsection{Results}
+TODO:
+true positives.
+false positives.
+graphs?
+
+\subsection{Discussion}
+TODO: 
+why the numbers are the way they are.
+what types of intrusions can we detect.
+which provenance data is most useful
 
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 TODO
+not realtime. can it be made to detect intrusions in realtime?
 

### 2013-11-21 writeup/paper.tex
small tweaks
Exploits Methodology

@@
 \usepackage{algorithm}
-\usepackage{moreverb}   
-\usepackage{longtable}
+%\usepackage{moreverb}   
+%\usepackage{longtable}
 \usepackage{fancyhdr}
 \usepackage{algorithmic}             
-\usepackage{algorithm}
-\usepackage{array}     
+%\usepackage{algorithm}
+%\usepackage{array}     
 \usepackage{hyperref}
@@
 \usepackage{amsmath, amssymb, amsthm}
-\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
-\usepackage{mathabx}
+%\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
+%\usepackage{mathabx}
 \usepackage{float}
@@
 
-\title{{\bf TODO: Intrusion Detection through Provenance}}
+\title{{\bf Intrusion Detection through Provenance-Based Histograms}}
 \author{
@@
 These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
+Somayaji's work relies on building a {\em normal} profile of a process, and then detecting when the process deviates from normal.
 
@@
 \subsection{Evaluation Methodology}
-TODO:
-types of exploits we ran.
-how we ran it.
-how we collected the data.
-the software we used.
+In order to test our techniques, we needed to run “exploits” on our PASS-enabled machine.
+We looked to prior work to get a sense of the different types of techniques commonly used to exploit security vulnerabilities.
+Commonly, remote network techniques [CITE DUKE PAPER] are used to broach insecure systems.
+Metasploit [CITE] is a software package designed to facilitate running exploits on insecure system.
+Searching the Metasploit database, as well as the online Exploit Database (exploit-db.com) [CITE], we found that most successful techniques involve exploiting vulnerabilities in software that a target system is running, as opposed to vulnerabilities in the target system itself.
+Put another way, we sought to exploit software packages, like mcrypt and gcc, not the Linux kernel itself.
+
+We found that many of the popular exploits involved buffer overflows and arbitrary code execution.
+Programming bugs in a piece of software allowed a malicious user, sending well-crafted data, to run arbitrary code on the target system.
+In particular, the mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
+Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like [CITE DUKE].
+To emulate this exploit, we designed our own program, which would decrypt (via ROT13) an input file.
+When sent a particular “exploit file” which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
+
+Another type of exploit is more reminiscent of a system administration failure.
+If a malicious user replaces a program in /usr/bin, it could be used to log user actions.
+To emulate this sort of attack, we replaced /usr/bin/gcc with our own version of gcc, which, in addition to compiling the user’s code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
+
+To generate the provenance data necessary to run the above analyses, we needed to run both “good” and “exploited” versions of our programs.
+For the mcrypt example, we ran our program on over 200 “good” inputs, and stored the provenance data.
+We then ran our program on two “bad” inputs, first to call “ls” as a Proof of Concept, and again to call “/bin/sh” to demonstrate that more severe attacks can be performed.
+We again exported the provenance generated by these steps, for comparison with our “good” usage.
+This is similar to recording all uses of a program during a server’s lifetime.
+
+For our system administration attack, we compiled over 200 C files using the unexploited version of GCC, and exported the provenance from these runs.
+We then replaced GCC with our exploited version, and compiled a few files, and exported the provenance again.
+In this way, we again acquired provenance for “normal” usage, before acquiring provenance for our attempted exploit.
 

### 2013-11-22 writeup/paper.tex
More to paper

@@
 
-{\bf TODO: TALK ABOUT OUR TECHNIQUES.} We build on TODO's work and extend ... blah blah blah ... combine techniques together ... blah blah. The contributions of this paper are:
+In particular, we generate statistical models of normal behavior, against which we can test a node of interest.
+These statiscal models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
+The contributions of this paper are:
 \begin{enumerate}
-\item TODO: the techniques we use to analyze provenance data
-\item TODO: types of intrusions that this system can detect, and the accuracy in detecting intrusions
-\item TODO: future work for using provenance data to detect intrusions
+\item A histogram-based technique for using provenance data to detect intrustions
+\item An analysis of the different intrusions that can be identified, and the types of provenance which best contribute to these detections
+\item A cogent plan for future work for using provenance data to detect intrusions
 \end{enumerate}
@@
 \section{Design and Implementation}
-TODO: ideas:
+Here we present our intrusion detection technique: a general and extendable statistical method for analyzing and comparing provenance data.
+At a high level, we process a provenance data graph, generating a model of "normal" behavior for each process/file in the graph.
+Using this model, we can determine the likelihood that a given "test" node is normal or whether it deviates from normal behavior and is thus potentially an exploited or malicious program.
+\subsection{Models}
+Our approach requires statistical models to evaluate the normalcy of a test node.
+The statistical model we chose to use was a kernel density estimation (KDE), which estimates the probability density function of a value with unknown distribution.
+To create the KDE, we needed to provide counts of a particular statistic.
+In this way, for each statistic that we chose to collect, we would be able to evaluate a test node's normalcy according to that statistic.
+For example, we could then ask "Does node $X_n$ have a statistically believable number of output processes compared to the other $X$ nodes?"
+\subsection{Provenance Statistics}
+As described above, in order to generate a KDE model, we need some statistic to measure.
+As we expected intrusions to differ in terms of the statistics that would stand out, we needed to generate a large number of statistics, and a model for each.
+With regards to provenance nodes directly, we compiled the following statistics:
 \begin{enumerate}
-\item somayaji: sequence of system calls
-\item semantic file attributes: determine types of files. use simple statistics on files.
-\item histograms: use simple statistics and build histogram of file attributes on a process name/file
-\item local clustering: use centrality metrics
-\item fuzzy rules
-\item hash k-tuples of provenance events and compare bit vectors
-\item backwards and forwards causality graphs
-\item access compatibility $\frac{1}{2}(C_d + C_e)$
-\item intersect good DAG and bad DAG and look at intersection of DAG
+\item The number of input files to the node.
+\item The number of input processes to the node.
+\item The number of input pipes to the node.
+\item The number of output files to the node.
+\item The number of output processes to the node.
+\item The number of ouput pipes to the node.
+\item file\_in\_version ??? XXX
+\item file\_out\_version ??? XXX
 \end{enumerate}
+For ease of computation and comparison, we treated the different node types (e.g.
+file, process, pipe) separately so that we actually had different counts for "input files to a file node" and "input files to a process node", for instance.
+\subsection{Centrality Metrics}
+In addition to the data that can be collected on the nodes directly, we could also compute clustering statistics by considering each node in the context of the larger graph.
+Using the methods espoused by Macko et al.
+\cite{clustering}, we could compute different centrality and clustering metrics for a node, which can then be tabulated into a histogram like the input/output counts above.
+KENNY TODO: EXPAND THIS
+\subsection{Implementation}
+The implementation of the above techniques consists of multiple Python files, spanning about 850 lines of (commented) code.
+The raw BerkeleyDB database generated by PASS is provided as input, and textual/graphical results are outputted.
+KENNY TODO: EXPAND THIS
 
@@
 
-For our system administration attack, we compiled over 200 C files using the unexploited version of GCC, and exported the provenance from these runs.
+For our system administration attack, we compiled the PASS toolset using GCC, and exported the provenance from these runs.
 We then replaced GCC with our exploited version, and compiled a few files, and exported the provenance again.
@@
 \subsection{Results}
+\subsection{Provenance Data}
+After running the normal and exploited programs described above, we visualized the provenance graphs using Orbiter \cite{orbiter}.
+After searching for the nodes named after our programs, we were able to find both the clean and exploited nodes.
+The clean nodes for "hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked "/usr/bin/ls" process.
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 1).
+Similarly, the normal GCC and exploited GCC displayed very different provenance graphs.
+\begin{figure*}
+  \caption{Descendents of an unexploited mcrypt-style program.
+Simply an output file.
+Below, descendents of an exploited mcrypt-style program.
+In addition to an (offscreen) output file, there is also an outgoing /bin/sh process, with its own children.}
+  \centering
+    \includegraphics[width=\textwidth]{img/mcrypt.png}
+\end{figure*}
+
+\subsection{Statistical Data}
+\subsection{KDE Predictions}
 TODO:

### 2013-11-22 writeup/paper.tex
Added Table

@@
 \subsection{Statistical Data}
+Table 1 summarizes some of the preliminary results we have generated.
+Here we have included, for each statistic and test, the average over the "normal" nodes, the observed value for the "test" node, and the KDE probability output for that value.
+We have chosen four metrics that provided a statistically significant result.
+The other statistics failed to be as powerful (see Table 2 XXX).
+% from http://hhh123.wordpress.com/2008/08/13/latex-table-across-two-columns/
+\begin{center}
+\begin{table*}[ht]
+{\small
+  \begin{tabular}{| l | l | l | l |}
+    \hline
+    metric & hello & mcrypt & gcc \\ \hline
+     \# forked processes & 0.0 | 1 | 0.108 & 0.0 | 1 | 0.107 & 0.9 | 9 | 0.408 \\ \hline
+     \# input files & 0.5 | 0 | 0.453 & 0.5 | 0 | 0.453 & 0.5 | 1 | 0.107 \\ \hline
+     \# ouptut files & 0.5 | 1 | 0.453 & 0.5 | 1 | 0.453 & 0.0 | 3 | 0.000 \\ \hline
+    Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\
+    \hline
+  \end{tabular}
+}
+\hfill{}
+\caption{average over good | test node value | kde(test node is normal)
+}
+\label{tb:tablename}
+\end{table*}
+\end{center}
 \subsection{KDE Predictions}

### 2013-11-22 writeup/paper.tex
Future work/conclusion

@@
 
-\section{Conclusion and Future Work}
-TODO
-not realtime. can it be made to detect intrusions in realtime?
+\section{Future Work}
+\subsection{N-depth Statistics}
+Currently, the statistics and counts we generate apply only to the test node itself.
+We found that certain programs did have the counts we would expect.
+GCC, for instance, forks processes for the linker/assembler that handles generating output, so GCC does not directly have output files.
+In order to account for these sorts of actions, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
+We could experiment with aggregation techniques to determine how best to incorporate this information.
+Perhaps it is best to simply sum all neighbor statistics together, or perhaps this information calls for a separate histogram and model.
+By utilizing information about neighboring nodes, we would hope to see an increase in accuracy and the ability to detect less blatant intrusions.
+\subsection{Histograms by Name}
+Currently, our system compute statistics based on node type, but not node name.
+For instance, When we compute the outgoing (i.e.
+"forked") processes, we simply count them.
+Thus, if a program normally forks something benign, like say "cp", but an exploited version forks "rm", the raw counts may not reveal this (a call to "/bin/sh", however, should impact the centrality measures).
+One direction to attempt to rectify this problem lies in generating more specific histograms.
+Instead of having a model for "number of processes forked", we could generate one model per type of process forked, e.g.
+"number of 'cp' processes forked".
+This more specific data might enable us to catch intrusions in larger programs that already modify a lot of files and fork a lot of processes.
+\subsection{ENV and ARGV}
+Currently we ignore the environment and argument information stored in the provenance graph.
+By generating histograms based on this information, we could catch subtler intrustions/exploits that rely on modified environment information.
+Additionally, if an exploit relies on particular argument or argument patterns, this information might reflected in the ARGV information.
+\subsection{Parameters}
+Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for construction the KDE.
+We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values.
+Additionally, we demonstrated that odd behavior has a "low" KDE, but we did not define what "low" means.
+For this system to be deployed more widely, we will need to define or compute what thresholds signify a potential intrusion or exploit.
+\subsection{Online Intrusion Detection}
+Currently, this system runs after-the-fact.
+We can identify intrusions that occurred on a machine by analyzing the data after it has been lifted from the target machine.
+Because PASS collect data in real time, however, our approach can be modified to implement online intrusion detection.
+To acheive this goal, we would need to only use metrics that can be updated without recomputing the entire statistic.
+In particular, some of the centrality measures do not meet this criterion.
+We would also need to add additional daemons (or modify the existing PASS daemons) to run the appropriate statistic gathering and model generating code.
+\section{Conclusion}
+We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomolies in program behavior, which often signal a security intrusion or exploit.
+Using provenance-based histograms, we can contruct models of the regular usage patterns for a particular program.
+When the intrusion or exploit results in behavior deviating from these models, its provenance allows us to flag it as a potential intrusion.
+Replaced binaries that acted significantly differently were especially easy to detect.
+Histograms on simple provenance statistics are good for identifying simple anomalies, but more complex statistics (see future work) may enable us to detect subtler intrusions.
 

### 2013-11-22 writeup/paper.tex
Example hist figure

@@
 KENNY TODO: EXPAND THIS
+\subsection{Examples}
+In Figure 1, we see a few examples of a generated histogram and KDE model.
+The green bars represent the histogram for the particular statistic, and the blue curve is the KDE.
+Finally, the red line is the value for a "test" node.
+In this case, the test node appears to be a possible exploit, because it intersects the KDE at low values.
+\begin{figure*}
+  \caption{Example histograms of a centrality measure and process statistic, with the assoicated KDE}
+  \centering
+    \includegraphics[width=\textwidth]{img/hist.png}
+\end{figure*}
 
@@
 The clean nodes for "hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked "/usr/bin/ls" process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 1).
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 2).
 Similarly, the normal GCC and exploited GCC displayed very different provenance graphs.
@@
 
-\subsection{Statistical Data}
+\subsection{Statistical Data and KDE Predictions}
 Table 1 summarizes some of the preliminary results we have generated.
@@
 \end{center}
-\subsection{KDE Predictions}
-TODO:
-true positives.
-false positives.
-graphs?
 

### 2013-11-22 writeup/paper.tex
Updated results/discussion

@@
 \subsection{Discussion}
-TODO: 
-why the numbers are the way they are.
-what types of intrusions can we detect.
-which provenance data is most useful
+\subsection{Successes}
+We were able to accurately identify our three exploited programs.
+Because each was a slightly different exploit, the data on which they deviated differed, but each did have metrics which flagged them as non-normal.
+\subsection{Useful Data for Intrusion Detection}
+We set out to identify both buffer overflows (i.e.
+arbitrary code execution) and modified binaries.
+The buffer overflows were best detected by the number of outgoing processes and the centrality measures.
+This makes sense, as arbitrary code execution will generate additional processes.
+The centrality measures are also very sensitive to the addition of nodes, and additional processes add both a new process node, as well as new nodes for the data touched by the new process.
+The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that the exploit will be caught by our model.
 
+If the exploit does directly not generate additional files, however, statistics like "number of output files" will fail to be statistically significant.
+Indeed, the two programs (hello and mcrypt) that did not generate additional files were not flagged by that measure.
+On the other hand, the GCC that was modified to generate log files was quickly caught by this measure, as it generated many more files than a normal GCC process.
+
+\subsection{Confusing Data Patterns}
+At first, some of the data we collected seemed wrong.
+Why, for instance, does GCC seem to normally output zero files?
+Upon further inspection, we found that GCC forks processes for the linker/assembler that handles generating output, so GCC does not directly have output files.
+Similarly, when running a python script, the code file itself is not directly executed, it is read in by a python process, which then goes off and collects input, generates output, and forks additional processes.
+Thus, when running a python file, the file itself only has one node in the provenance graph, with many python process nodes accessing it.
+To run our analyses, we needed to manually identify this connection, and compute statistics for the python process, instead of the single code file node.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Currently, the statistics and counts we generate apply only to the test node itself.
-We found that certain programs did have the counts we would expect.
-GCC, for instance, forks processes for the linker/assembler that handles generating output, so GCC does not directly have output files.
+As mentioned above, some processes fork helper processes that take in input or generate output, which would be useful information to collect.
 In order to account for these sorts of actions, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
@@
 By utilizing information about neighboring nodes, we would hope to see an increase in accuracy and the ability to detect less blatant intrusions.
+
+Additionally, by using this approach, we could avoid the manual process identification mentioned above.
+On python files, for instance, N-depth statistics would compute statistics for the python process, which is always a direct neighbor of the code file we initially cared about.
 \subsection{Histograms by Name}

### 2013-11-24 writeup/paper.tex
Abstract and Related Works

@@
 \begin{abstract}
-TODO
+Data-driven intrusion detection is a common system security technique.
+Here we present a method for generating a statistical model based approach to intrusion detection, using provenance data generated by a provenance-aware modified file system.
+Our analyses reveal that provenance data can be directly used to detect intrusions, and an in-depth analysis reveals which data is most useful in distinguishing between safe and malicious programs and program executions.
+Our histogram based approach requires no prior intuition about the types of exploits potentially being run on a system, and identifies suspicious behavior based on deviations from prior normal system behavior.
 \end{abstract}
@@
 al.
-have developed Backtrack \cite{backtrack}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
-Given an intrusion, Backtrack builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtrack with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
+have developed Backtracker \cite{backtracker}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
+Given an intrusion, Backtracker builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtracker with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
 However, their system does not find a way of detecting an intrusion.
@@
 
+\section{Background and Related Works}
+\subsection{PASS}
+Provenance data is meta-data detailing the creation, transformation, and flow of data.
+PASS \cite{pass, passv2} is a provenance-aware storage system, designed to generate and track this provenance data seamlessly.
+The PASS system automatically creates and updates provenance data fore every file that was created or modified the volume.
+The PASS provenance data is very rich data.
+For example, the provenance for a data file that sent in a zip file would reveal both the name of the program that unzipped the file, as well as the web browser through which the zip file was downloaded!
+It is our belief that this level of specificity will lend itself to detecting patterns in program usage.
+
+The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
+\subsection{Intrusion Detection}
+Intrusion detection in live systems is a common problem in the systems community.
+There are two primary components to an intrusion detection system (IDS): (1) the data used to fuel the system's analysis, and (2) the process in which this data is used.
+
+Somayaji et al.
+\cite{somayaji} aimed to detect intrusions by keeping track of all systems calls that occur during a process's execution time.
+After collecting this data over a long period of "normal" time, if a proces exhibits system call patterns that deviate from the "normal" patterns for the process, their system flags it as potentialy dangerous, and slows its execution.
+In this way, a program that continually deviates from normal behavior will be descheduled entirely.
+
+Tariq et al.
+\cite{correlated-anomalies} extend the work of Somayaji and others, designing a system to run an IDS on a set of distributed nodes.
+They use the IDS of other researchers, distilling a programs execution into a set of values (a "k-tuple") used to identify information of interest regarding the program's exeuction.
+Our system is designed to run offline, so we run our algorithms on the full set of provenance data.
+While this approach is slower, it allows us to take in the full picture of what occurred during a process's execution, without having to preselect the information we think will be useful.
+This approach is especially helpful because different information leads to the detection of different types of intrusions.
+
+Backtracker \cite{backtracker} is a system most similar to ours in design.
+It, like PASS, collects data while the system is running, and like our system analyzes it offline for potential intrusions.
+Whereas PASS collects data directly at the operating system level (using a modified file system), Backtracker runs its target OS inside of a virtual machine, and uses a modified virtual machine monitor to collect data.
+The data they collect is very similar to the data PASS generates, and their exploits and analysis techniques directly inspired our work.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We looked to prior work to get a sense of the different types of techniques commonly used to exploit security vulnerabilities.
-Commonly, remote network techniques [CITE DUKE PAPER] are used to broach insecure systems.
+Commonly, remote network techniques (as seen in Backtracker \cite{backtracker}) are used to broach insecure systems.
 Metasploit [CITE] is a software package designed to facilitate running exploits on insecure system.
@@
 In particular, the mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
-Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like [CITE DUKE].
+Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like the authors of Backtracker \cite{backtracker} found.
 To emulate this exploit, we designed our own program, which would decrypt (via ROT13) an input file.
@@
 
-\bibitem{backtrack}
+\bibitem{backtracker}
 \textsc{King, S. T. and Chen, P. M.} Backtracking Intrusions. In {\em SOSP'03 Proceedings of the nineteenth ACM symposium on Operating systems principles} (December 2003).

### 2013-11-24 writeup/paper.tex
Expanded methodology. 2 more figures

@@
 \item The number of ouput pipes to the node.
-\item file\_in\_version ??? XXX
-\item file\_out\_version ??? XXX
+\item file\_in\_version The "version" number of an incoming file.
+\item file\_out\_version The "version" number of an outgoing file.
 \end{enumerate}
@@
 Commonly, remote network techniques (as seen in Backtracker \cite{backtracker}) are used to broach insecure systems.
-Metasploit [CITE] is a software package designed to facilitate running exploits on insecure system.
-Searching the Metasploit database, as well as the online Exploit Database (exploit-db.com) [CITE], we found that most successful techniques involve exploiting vulnerabilities in software that a target system is running, as opposed to vulnerabilities in the target system itself.
+Metasploit \cite{metasploit} is a software package designed to facilitate running exploits on insecure system.
+Searching the Metasploit database, as well as the online Exploit Database \cite{exploitdb}, we found that most successful techniques involve exploiting vulnerabilities in software that a target system is running, as opposed to vulnerabilities in the target system itself.
 Put another way, we sought to exploit software packages, like mcrypt and gcc, not the Linux kernel itself.
@@
 
+\subsection{Selecting Metrics}
+Because we had no prior intuition into what the provenance data would actually look like, we approached selecting metrics from an objective, data-based standpoint.
+We wanted to use metrics that had variation over our workload, so as not to minimize the introduction of workload dependent data.
+Towards this end, we graphed all of the metrics over the "normal" provenance data, to see where there were trends and variation.
+Our results for GCC are in Figures 2 and 3.
+We can see that proc\_out\_proc (outgoing processes) and the ancestor and opsahl centrality metrics have more variation, and will be less likely to introduce biases.
+As we include more data (as discussed in Future Works), these sorts of graphs will become increasingly important in selecting good metrics for evaluation.
+\begin{figure}
+  \caption{Histograms over all "simple" provenance statistics for GCC}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/gcc_stats.png}
+\end{figure}
+\begin{figure}
+  \caption{Histograms over all centrality metrics for GCC}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/gcc_cent_stats.png}
+\end{figure}
+
 \subsection{Results}
@@
 The clean nodes for "hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked "/usr/bin/ls" process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 2).
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 4).
 Similarly, the normal GCC and exploited GCC displayed very different provenance graphs.
@@
 We have chosen four metrics that provided a statistically significant result.
-The other statistics failed to be as powerful (see Table 2 XXX).
+% The other statistics failed to be as powerful (see Table 2 XXX).
 % from http://hhh123.wordpress.com/2008/08/13/latex-table-across-two-columns/
@@
 
+\bibitem{metasploit}
+\textsc{Rapid 7 Inc.} Metasploit Framework. \tt{http://www.metasploit.com}.
+
+\bibitem{exploitdb}
+\textsc{Offensive Security, Inc.} The Exploit Database. \tt{http://www.exploit-db.com}.
 \end{thebibliography}

### 2013-11-24 writeup/paper.tex
about to start evaluation

@@
 Data-driven intrusion detection is a common system security technique.
-Here we present a method for generating a statistical model based approach to intrusion detection, using provenance data generated by a provenance-aware modified file system.
-Our analyses reveal that provenance data can be directly used to detect intrusions, and an in-depth analysis reveals which data is most useful in distinguishing between safe and malicious programs and program executions.
-Our histogram based approach requires no prior intuition about the types of exploits potentially being run on a system, and identifies suspicious behavior based on deviations from prior normal system behavior.
+We present a histogram-based approach to intrusion detection, using provenance data generated by a provenance-aware storage system.
+By looking at simple provenance statistics and centrality metrics on provenance graphs, our approach identifies suspicious behavior based on deviations from prior normal system behavior.
+Our histogram-based approach requires no prior intuition about the types of exploits potentially being run on a system, and our analysis of two buffer overflow attacks and a modified version of \texttt{gcc} to log actions reveal which data is most useful in distinguishing between safe and malicious programs and program executions.
 \end{abstract}
@@
 forked, input, versioning).
+The edges show dependency, not data flow, and as a result, edges are directed from outputs to inputs.
 
@@
 al.
-have developed techniques for detecting intrusions based on system call sequences, and counteract these by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
+have developed techniques for detecting intrusions based on system call sequences, and their system counteracts these intrusions by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
 These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
@@
 
-Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies, it seems plausible to detect intrusions based on provenance data collected by PASS.
 Tariq et.
 al.
-generalize Somayaji's sequences of system calls to sequences of provenance system events \cite{correlated-anomalies} to develop an intrusion detection system (IDS) in a distributed system.
-Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of provenance events--and then use these bit vectors to correlate anomalies across multiple hosts.
+generalize Somayaji's sequences of system calls to sequences of system events \cite{correlated-anomalies} to develop an intrusion detection system (IDS) in a distributed system.
+Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of systemevents--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
 King et.
@@
 Given an intrusion, Backtracker builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtracker with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
-However, their system does not find a way of detecting an intrusion.
+However, their system does not have a way of detecting an intrusion.
 Lei et.
@@
 
-In this paper, we present a technique to analyze existing provenance data to detect if an intrusion occurred.
+Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies and existing work in intrusion detection that uses these kinds of directed graphs of files and process data, it seems plausible to detect intrusions based on provenance data collected by PASS.
+In this paper, we present a histogram-based technique for analyzing existing PASS provenance data to detect if an intrusion occurred.
 
 In particular, we generate statistical models of normal behavior, against which we can test a node of interest.
-These statiscal models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
+These statistical models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
+The lower the kernel density estimate, the less ``normal" the node.
 The contributions of this paper are:
 \begin{enumerate}
-\item A histogram-based technique for using provenance data to detect intrustions
+\item A histogram-based technique for using provenance data to detect intrusions
 \item An analysis of the different intrusions that can be identified, and the types of provenance which best contribute to these detections
-\item A cogent plan for future work for using provenance data to detect intrusions
+\item Future extensions to our work to create better intrusion detection models and make more use of provenance data
 \end{enumerate}
@@
 \subsection{PASS}
-Provenance data is meta-data detailing the creation, transformation, and flow of data.
-PASS \cite{pass, passv2} is a provenance-aware storage system, designed to generate and track this provenance data seamlessly.
-The PASS system automatically creates and updates provenance data fore every file that was created or modified the volume.
-The PASS provenance data is very rich data.
+Provenance data is met-data detailing the creation, transformation, and flow of data.
+PASS \cite{pass, passv2} is a provenance-aware storage system, designed to generate and track this provenance data seamlessly, and Table \ref{passobjects} indicates the types of objects the PASS tracks and attributes with those objects.
+PASS allocates pnode (``PASS" node) number and version numbers whenever PASS encounters a new object or a modification to an existing object.
+The combination of (pnode number, version number) uniquely identifies a node in our provenance graph.
+
+The PASS system automatically creates and updates provenance data for every file that was created or modified the volume.
 For example, the provenance for a data file that sent in a zip file would reveal both the name of the program that unzipped the file, as well as the web browser through which the zip file was downloaded!
@@
 
-The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
+There are two versions of PASS. We conducted our experiments only with PASSv2; thus in this paper ``PASS" to refer to PASSv2.
+
+TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
+
+\begin{table}[htdp]
+\begin{center}
+\begin{tabular}{|c|c|}
+\hline
+{\bf Node Types} & {\bf Description} \\
+\hline
+PROC & process \\
+FILE & file \\
+PIPE & pipe \\
+NPFILE & non-PASS file \\
+OTHER & other types of objects \\
+\hline
+{\bf Node Attributes} & {\bf Description} \\
+\hline
+INODE & inode of file \\
+PATH & path of file \\
+FREEZETIME & timestamp of record \\
+CREAT & creation of file \\
+UNLINK & unlinking file \\
+EXECTIME & exec time of process\\
+NAME & name of process \\
+PID & pid of of process \\
+ARGV & argument vector to process \\
+ENV & environment of a process \\
+\hline
+{\bf Edge Types} & {\bf Description} \\
+\hline
+FORKPARENT & indicates parent PROC\\
+&  of existing PROC node \\
+INPUT & indicates input files\\
+& or processes \\
+VERSION & indicates a modification  \\
+& to an existing node (e.g. \\
+& modifying a file) \\
+\hline
+\end{tabular}
+\end{center}
+\label{passobjects}
+\caption{Types of data collected by PASS}
+\end{table}%
+
 \subsection{Intrusion Detection}
 Intrusion detection in live systems is a common problem in the systems community.
-There are two primary components to an intrusion detection system (IDS): (1) the data used to fuel the system's analysis, and (2) the process in which this data is used.
+There are two primary components to an intrusion detection system (IDS): (1) the data used to fuel the system's analysis, and (2) how the data is used.
 
 Somayaji et al.
-\cite{somayaji} aimed to detect intrusions by keeping track of all systems calls that occur during a process's execution time.
-After collecting this data over a long period of "normal" time, if a proces exhibits system call patterns that deviate from the "normal" patterns for the process, their system flags it as potentialy dangerous, and slows its execution.
+\cite{somayaji} aim to detect intrusions by keeping track of all systems calls that occur during a process's execution time.
+After collecting this data over a long period of "normal" time, if a proces exhibits system call patterns that deviate from the ``normal" patterns for the process, their system flags it as potentially dangerous, and slows its execution.
 In this way, a program that continually deviates from normal behavior will be descheduled entirely.
@@
 Tariq et al.
-\cite{correlated-anomalies} extend the work of Somayaji and others, designing a system to run an IDS on a set of distributed nodes.
-They use the IDS of other researchers, distilling a programs execution into a set of values (a "k-tuple") used to identify information of interest regarding the program's exeuction.
+\cite{correlated-anomalies} extend the work of Somayaji, designing a system to run an IDS on a set of distributed nodes.
+They use the IDS of other researchers, distilling a programs execution into a set of $k$-tuples of system events used to identify normal behavior during the program's execution.
 Our system is designed to run offline, so we run our algorithms on the full set of provenance data.
-While this approach is slower, it allows us to take in the full picture of what occurred during a process's execution, without having to preselect the information we think will be useful.
+While this approach is slower, it allows us to take in the full picture of the events during a process's execution, without having to preselect the information we think will be useful.
 This approach is especially helpful because different information leads to the detection of different types of intrusions.
@@
 Backtracker \cite{backtracker} is a system most similar to ours in design.
-It, like PASS, collects data while the system is running, and like our system analyzes it offline for potential intrusions.
-Whereas PASS collects data directly at the operating system level (using a modified file system), Backtracker runs its target OS inside of a virtual machine, and uses a modified virtual machine monitor to collect data.
+It, like PASS, builds a directed graph of process-file dependencies while the system is running, and, like our system, analyzes it offline for potential intrusions.
 The data they collect is very similar to the data PASS generates, and their exploits and analysis techniques directly inspired our work.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here we present our intrusion detection technique: a general and extendable statistical method for analyzing and comparing provenance data.
-At a high level, we process a provenance data graph, generating a model of "normal" behavior for each process/file in the graph.
-Using this model, we can determine the likelihood that a given "test" node is normal or whether it deviates from normal behavior and is thus potentially an exploited or malicious program.
+At a high level, we create a provenance data graph and generate a model of ``normal" behavior for each process/file in the graph.
+Using this model, we can determine the likelihood that a given ``test" deviates from normal behavior and is thus potentially an exploited or malicious program.
 \subsection{Models}
@@
 To create the KDE, we needed to provide counts of a particular statistic.
-In this way, for each statistic that we chose to collect, we would be able to evaluate a test node's normalcy according to that statistic.
-For example, we could then ask "Does node $X_n$ have a statistically believable number of output processes compared to the other $X$ nodes?"
+For each statistic:
+\begin{enumerate}
+\item We computed the counts for that statistic over all the nodes in the provenance graph.
+\item We aggregated the counts by the NAME field of the node (e.g.
+the path to a file for a FILE node, or path to an executable for PROC node).
+For objects with no NAME field (e.g.
+PIPE objects), we used the tuple of (pnode number, VERSION number) as the node's name.
+\item For each unique name in the graph, we use the counts on that name to generate a histogram of ``normal" behavior.
+We convert these histograms into kernel density objects to smoothen our histogram and generalize to all continuous values for the statistic.
+\end{enumerate}
+
+This model allows us to evaluate a test node's normalcy according to that statistic.
+For example, we could then ask ``Does node $X'$ have a statistically believable number of output processes compared to the other $X$ nodes?"
+
 \subsection{Provenance Statistics}
-As described above, in order to generate a KDE model, we need some statistic to measure.
-As we expected intrusions to differ in terms of the statistics that would stand out, we needed to generate a large number of statistics, and a model for each.
-With regards to provenance nodes directly, we compiled the following statistics:
+We expected intrusions to differ in terms of the statistics that would stand out, and as a result, we needed to generate a diverse suite of statistics and a model for each.
+For our simple provenance graph statistics, we compiled the following statistics:
 \begin{enumerate}
@@
 \item The number of output processes to the node.
-\item The number of ouput pipes to the node.
-\item file\_in\_version The "version" number of an incoming file.
-\item file\_out\_version The "version" number of an outgoing file.
+\item The number of output pipes to the node.
+\item The number of VERSION edges into this node.
+\item The number of VERSION edges out of this node.
 \end{enumerate}
 For ease of computation and comparison, we treated the different node types (e.g.
-file, process, pipe) separately so that we actually had different counts for "input files to a file node" and "input files to a process node", for instance.
+file, process, pipe) separately so that we can treat, for instance, "input files to a file node" and "input files to a process node" as separate counts.
+
 \subsection{Centrality Metrics}
-In addition to the data that can be collected on the nodes directly, we could also compute clustering statistics by considering each node in the context of the larger graph.
-Using the methods espoused by Macko et al.
-\cite{clustering}, we could compute different centrality and clustering metrics for a node, which can then be tabulated into a histogram like the input/output counts above.
-KENNY TODO: EXPAND THIS
+In addition to the data that can be collected on the nodes directly, we computed centrality metrics on the nodes using the centrality metrics proposed by Macko et.
+al.
+\cite{clustering}.
+In their work, they noted 3 properties of provenance graphs:
+\begin{enumerate}
+\item The ubiquity (importance) of a node is a function of a node's descendants, not its ancestors.
+(Note: Because edges in provenance graphs are in the direction of data dependence instead of data flow, ``descendants" in this case means data flow descendants and NOT descendants by traversing provenance graph edges).
+\item There is no agreed-upon granularity in which provenance should be captured.
+\item Provenance has a temporal component.
+\end{enumerate}
+
+Using these observations, they propose several centrality metrics on provenance graphs:
+\begin{itemize}
+\item \textbf{In-degree Centrality}.
+Because a node's importance is based on its output descendants, then the number of times a node was used as input to another node indicates a simple measure of importance.
+\item \textbf{Age}. They note that large "jumps" in timestamps or age might indicate breaks between tasks, but this interpretation is complicated when scripts execute tasks in short succession.
+\item \textbf{Ancestor Centrality}. This measures the total number of descendants of a node, normalized by the total number of nodes in a graph.
+\item \textbf{Opsahl's Closeness Centrality}. This accounts for the number of edges between a node and its descendants. To compute this, we calculate
+$$CC'(v) = \sum_{x \in V -  \{v\}} (d'_{xv})^{-1}$$
+where $d'_{xv}$ is the distance from node $x$ to $v$ by only following outgoing edges from $x$. Unreachable nodes from $x$ will have $d'_{xv} = \infty$, and so $(d'_{xv})^{-1} = 0$.
+\item \textbf{Provenance Eigenvector Centrality}. The centrality of the nodes is given by the left dominant eigenvector of the matrix:
+$$M_{ij} =
+\begin{cases}
+1 & \text{if there is an edge } i \to j \\
+1/|V| & \text{if there is no out edge from } i \\
+0 & \text{otherwise}
+\end{cases}
+$$
+\end{itemize}
+
+We each metric, we computed the centrality of all the nodes and aggregated the counts by name to generate our histogram and kernel density estimations like the simple statistics we used above.
+
 \subsection{Implementation}
-The implementation of the above techniques consists of multiple Python files, spanning about 850 lines of (commented) code.
-The raw BerkeleyDB database generated by PASS is provided as input, and textual/graphical results are outputted.
-KENNY TODO: EXPAND THIS
+We used Orbiter \cite{orbiter} to visualize and examine the provenance graphs.
+We wrote Python scripts to (1) parse the raw BerkeleyDB database files storing the PASS data (2) build the provenance graphs (3) collect the counts for each statistic (which include running the centrality calculations), and (4) optionally outputting the histograms and kernel density estimations for a node of interest in a second provenance graph.
+
 \subsection{Examples}
@@
 The green bars represent the histogram for the particular statistic, and the blue curve is the KDE.
-Finally, the red line is the value for a "test" node.
+Finally, the red line is the value for a ``test" node.
 In this case, the test node appears to be a possible exploit, because it intersects the KDE at low values.
 \begin{figure*}
-  \caption{Example histograms of a centrality measure and process statistic, with the assoicated KDE}
+  \caption{Example histograms of a centrality measure and process statistic for the \texttt{gcc} exploit, with the associated KDE.
+The green bars represent the raw histogram counts, the blue curve represents the kernel density function, and the red vertical line represents the estimation of a node of interest.}
   \centering

### 2013-11-24 writeup/paper.tex
about to merge intro and related works

@@
 %\usepackage{mathabx}
+
 \usepackage{float}
@@
 
-TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
+% TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
 
@@
 \subsection{Examples}
-In Figure 1, we see a few examples of a generated histogram and KDE model.
+In Figure \ref{kde-example}, we see a few examples of a generated histogram and KDE model.
 The green bars represent the histogram for the particular statistic, and the blue curve is the KDE.
@@
 \begin{figure*}
+  \label{kde-example}
   \caption{Example histograms of a centrality measure and process statistic for the \texttt{gcc} exploit, with the associated KDE.
@@
 \subsection{Evaluation Methodology}
-In order to test our techniques, we needed to run “exploits” on our PASS-enabled machine.
+In order to test our techniques, we needed to run exploits on our PASS-enabled machine.
 We looked to prior work to get a sense of the different types of techniques commonly used to exploit security vulnerabilities.
-Commonly, remote network techniques (as seen in Backtracker \cite{backtracker}) are used to broach insecure systems.
+Commonly, remote network techniques (as seen in Backtracker \cite{backtracker}) are used to breach insecure systems.
 Metasploit \cite{metasploit} is a software package designed to facilitate running exploits on insecure system.
@@
 
+To generate the provenance data necessary to run the analyses, we needed to run both good and exploited versions of our programs.
+
+\subsubsection{hello}
 We found that many of the popular exploits involved buffer overflows and arbitrary code execution.
 Programming bugs in a piece of software allowed a malicious user, sending well-crafted data, to run arbitrary code on the target system.
-In particular, the mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
+We created a simple \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a process.
+
+We ran \texttt{hello} normally 40 times on various inputs and collected the provenance times.
+We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
+See Figure \ref{hello-orbiter} for provenance graphs displaying normal and exploited behavior of \texttt{hello}.
+
+\subsubsection{mcrypt}
+The mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
 Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like the authors of Backtracker \cite{backtracker} found.
-To emulate this exploit, we designed our own program, which would decrypt (via ROT13) an input file.
-When sent a particular “exploit file” which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
+To emulate this exploit, we designed a Python program, which would decrypt (via ROT13) an input file.
+When sent a particular exploit fileâ which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
 
+We ran our program on over 200 good inputs, and stored the provenance data.
+We then ran our program on two bad inputs, first to call \texttt{ls} as a Proof of Concept, and again to call \texttt{/bin/sh} to demonstrate that more severe attacks can be performed.
+We again exported the provenance generated by these steps, for comparison with our good usage.
+This is similar to recording all uses of a program during a server's lifetime.
+See Figure \ref{mcrypt-orbiter} displaying normal and exploited behavior for our version of mcrypt.
+
+\subsubsection{gcc}
 Another type of exploit is more reminiscent of a system administration failure.
-If a malicious user replaces a program in /usr/bin, it could be used to log user actions.
-To emulate this sort of attack, we replaced /usr/bin/gcc with our own version of gcc, which, in addition to compiling the user’s code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
-
-To generate the provenance data necessary to run the above analyses, we needed to run both “good” and “exploited” versions of our programs.
-For the mcrypt example, we ran our program on over 200 “good” inputs, and stored the provenance data.
-We then ran our program on two “bad” inputs, first to call “ls” as a Proof of Concept, and again to call “/bin/sh” to demonstrate that more severe attacks can be performed.
-We again exported the provenance generated by these steps, for comparison with our “good” usage.
-This is similar to recording all uses of a program during a server’s lifetime.
+If a malicious user replaces a program in \texttt{/usr/bin}, it could be used to log user actions.
+To emulate this sort of attack, we replaced \texttt{/usr/bin/gcc} with our own version of gcc, which, in addition to compiling the user's code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
 
 For our system administration attack, we compiled the PASS toolset using GCC, and exported the provenance from these runs.
-We then replaced GCC with our exploited version, and compiled a few files, and exported the provenance again.
-In this way, we again acquired provenance for “normal” usage, before acquiring provenance for our attempted exploit.
+We then replaced GCC with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
+In this way, we collected provenance for normal usage and from our exploited usage.
+\begin{figure*}
+  \label{hello-orbiter}
+  \caption{Descendants of a \texttt{hello} program.
+The top image shows multiple invocations of a \texttt{hello} behaving normally when invoked with \texttt{cat foo.txt | ./hello > c.txt}.
+The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.}
+  \centering
+    \includegraphics[width=\textwidth]{img/hello.png}
+\end{figure*}
+\begin{figure*}
+  \label{mcrypt-orbiter}
+  \caption{Descendants of an unexploited mcrypt-style program.
+Simply an output file.
+Below, descandents of an exploited mcrypt-style program.
+In addition to an (offscreen) output file, there is also an outgoing \texttt{/bin/sh} process, with its own children.}
+  \centering
+    \includegraphics[width=\textwidth]{img/mcrypt.png}
+\end{figure*}
 
 \subsection{Selecting Metrics}
-Because we had no prior intuition into what the provenance data would actually look like, we approached selecting metrics from an objective, data-based standpoint.
-We wanted to use metrics that had variation over our workload, so as not to minimize the introduction of workload dependent data.
-Towards this end, we graphed all of the metrics over the "normal" provenance data, to see where there were trends and variation.
-Our results for GCC are in Figures 2 and 3.
-We can see that proc\_out\_proc (outgoing processes) and the ancestor and opsahl centrality metrics have more variation, and will be less likely to introduce biases.
-As we include more data (as discussed in Future Works), these sorts of graphs will become increasingly important in selecting good metrics for evaluation.
-\begin{figure}
+Because we had no prior intuition into what the histograms on provenance data would actually look like, we approached selecting statistics from an objective, data-based standpoint.
+We wanted to use statistics that had variation over our workload to minimize the dependency of our statistics on the workload data.
+Towards this end, we graphed all of the statistics over the ``normal" provenance data, to see where there were trends and variation.
+Our results for GCC are in Figures \ref{simple-gcc} and \ref{centrality-gcc}.
+We can see that proc\_out\_proc (outgoing processes), the ancestor centrality, and Opsahl Closness Centrality statistics have more variation, whereas the other statistics were unimodal or bimodal (0 or 1).
+
+The unimodal and bimodal statistics would not generate useful kernel density estimates, as they would predict roughly 0.5 for either a 0 or 1 count, giving us no information about whether the node in question is behaving normally.
+On the other hand, the statistics with more variation in data would actually provide more useful information when determining normalcy in new test nodes.
+As we include more data (as discussed in Future Work), these sorts of graphs will become increasingly important in selecting good statistics for evaluation.
+\begin{figure*}
+  \label{simple-gcc}
   \caption{Histograms over all "simple" provenance statistics for GCC}
   \centering
-    \includegraphics[width=0.5\textwidth]{img/gcc_stats.png}
-\end{figure}
-\begin{figure}
+    \includegraphics[width=0.7\textwidth]{img/gcc_stats.png}
+\end{figure*}
+\begin{figure*}
+  \label{centrality-gcc}
   \caption{Histograms over all centrality metrics for GCC}
   \centering
-    \includegraphics[width=0.5\textwidth]{img/gcc_cent_stats.png}
-\end{figure}
+    \includegraphics[width=0.7\textwidth]{img/gcc_cent_stats.png}
+\end{figure*}
 
@@
 After searching for the nodes named after our programs, we were able to find both the clean and exploited nodes.
-The clean nodes for "hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked "/usr/bin/ls" process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendents (see Figure 4).
-Similarly, the normal GCC and exploited GCC displayed very different provenance graphs.
-\begin{figure*}
-  \caption{Descendents of an unexploited mcrypt-style program.
-Simply an output file.
-Below, descendents of an exploited mcrypt-style program.
-In addition to an (offscreen) output file, there is also an outgoing /bin/sh process, with its own children.}
-  \centering
-    \includegraphics[width=\textwidth]{img/mcrypt.png}
-\end{figure*}
+The clean nodes for ``hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked \texttt{/usr/bin/ls} process.
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendants (see Figure \ref{hello-orbiter} and Figure \ref{mcrypt-orbiter}).
+Similarly, the normal GCC and exploited GCC displayed very different provenance graphs, but they were too big to fit legibly on a page.
 
 \subsection{Statistical Data and KDE Predictions}
-Table 1 summarizes some of the preliminary results we have generated.
+Table \ref{results} summarizes some of the preliminary results we have generated.
 Here we have included, for each statistic and test, the average over the "normal" nodes, the observed value for the "test" node, and the KDE probability output for that value.
@@
 % from http://hhh123.wordpress.com/2008/08/13/latex-table-across-two-columns/
+% TODO: include difference/ratio from normal?
+\begin{table*}[ht]
+\label{results}
+{\small
   \begin{center}
-\begin{table*}[ht]
-{\small
   \begin{tabular}{| l | l | l | l |}
@@
     metric & hello & mcrypt & gcc \\ \hline
-     \# forked processes & 0.0 | 1 | 0.108 & 0.0 | 1 | 0.107 & 0.9 | 9 | 0.408 \\ \hline
+     \# forked processes & 0.0 | 1 | 0.108 & 0.0 | 1 | 0.107 & 0.9 | 0 | 0.408 \\ \hline
      \# input files & 0.5 | 0 | 0.453 & 0.5 | 0 | 0.453 & 0.5 | 1 | 0.107 \\ \hline
@@
   \end{tabular}
+  \end{center}
 }
 \hfill{}
-\caption{average over good | test node value | kde(test node is normal)
+\caption{average over good | test node value | kernel density estimate of test node being ``normal"
 }
@@
 \end{table*}
-\end{center}
 
@@
 \subsection{Successes}
-We were able to accurately identify our three exploited programs.
-Because each was a slightly different exploit, the data on which they deviated differed, but each did have metrics which flagged them as non-normal.
+We were able to accurately identify our three exploited programs. Each was a slightly different exploit, but each exploit deviated enough on some metrics which flagged them as non-normal.
+
 \subsection{Useful Data for Intrusion Detection}
@@
 This makes sense, as arbitrary code execution will generate additional processes.
-The centrality measures are also very sensitive to the addition of nodes, and additional processes add both a new process node, as well as new nodes for the data touched by the new process.
+Of the centrality metrics we implemented, the Opsahl Centrality metric did the best in distinguishing an exploited node from normally behaving nodes.
+Because the Opsahl Centrality metric accounts for the number of edges between a node and all of its descendants, this centrality metric is very sensitive to the addition of nodes.
+Exploits that have additional outputs (e.g.
+new processes and output files through a remote shell), thus generate a higher Opsahl centrality metric for the offending node.
 The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that the exploit will be caught by our model.
@@
 Indeed, the two programs (hello and mcrypt) that did not generate additional files were not flagged by that measure.
-On the other hand, the GCC that was modified to generate log files was quickly caught by this measure, as it generated many more files than a normal GCC process.
+On the other hand, the GCC that was modified to generate log files was caught by this measure, as it generated many more files than a normal GCC process.
 
@@
 To run our analyses, we needed to manually identify this connection, and compute statistics for the python process, instead of the single code file node.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We could experiment with aggregation techniques to determine how best to incorporate this information.
-Perhaps it is best to simply sum all neighbor statistics together, or perhaps this information calls for a separate histogram and model.
-By utilizing information about neighboring nodes, we would hope to see an increase in accuracy and the ability to detect less blatant intrusions.
+Perhaps it is best to simply sum all neighbor statistics together, or perhaps this information calls for new histogram and model.
+By utilizing information about neighboring nodes, we hope to see an increase in accuracy and the ability to detect less blatant intrusions.
 
@@
 On python files, for instance, N-depth statistics would compute statistics for the python process, which is always a direct neighbor of the code file we initially cared about.
+
+\subsection{Other Centrality Metrics}
+Progressive Pagerank
+
 \subsection{Histograms by Name}
@@
 "forked") processes, we simply count them.
-Thus, if a program normally forks something benign, like say "cp", but an exploited version forks "rm", the raw counts may not reveal this (a call to "/bin/sh", however, should impact the centrality measures).
+Thus, if a program normally forks something benign, like say "cp", but an exploited version forks "rm", the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures).
 One direction to attempt to rectify this problem lies in generating more specific histograms.
@@
 By generating histograms based on this information, we could catch subtler intrustions/exploits that rely on modified environment information.
-Additionally, if an exploit relies on particular argument or argument patterns, this information might reflected in the ARGV information.
+Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
 \subsection{Parameters}
@@
 Because PASS collect data in real time, however, our approach can be modified to implement online intrusion detection.
-To acheive this goal, we would need to only use metrics that can be updated without recomputing the entire statistic.
+To achieve this goal, we would need to only use metrics that can be updated without recomputing the entire statistic.
 In particular, some of the centrality measures do not meet this criterion.
@@
 \section{Conclusion}
-We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomolies in program behavior, which often signal a security intrusion or exploit.
-Using provenance-based histograms, we can contruct models of the regular usage patterns for a particular program.
+We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomalies in program behavior, which often signal a security intrusion or exploit.
+Using provenance-based histograms, we can construct models of the regular usage patterns for a particular program.
 When the intrusion or exploit results in behavior deviating from these models, its provenance allows us to flag it as a potential intrusion.
@@
 
+\bibitem{exploitdb}
+\textsc{Offensive Security, Inc.} The Exploit Database. {\tt http://www.exploit-db.com}.
+
+\bibitem{metasploit}
+\textsc{Rapid 7 Inc.} Metasploit Framework. {\tt http://www.metasploit.com}.
+
 \bibitem{somayaji}
@@
 
-\bibitem{metasploit}
-\textsc{Rapid 7 Inc.} Metasploit Framework. \tt{http://www.metasploit.com}.
-
-\bibitem{exploitdb}
-\textsc{Offensive Security, Inc.} The Exploit Database. \tt{http://www.exploit-db.com}.
 \end{thebibliography}

### 2013-11-24 writeup/paper.tex
merged intro and related work; add local vs global to future work

@@
 
-\section{Introduction}
+\section{Introduction and Related Work}
+
+\subsection{Provenance and PASS}
+
 Provenance is metadata that tracks the history of all changes to files.
@@
 The edges show dependency, not data flow, and as a result, edges are directed from outputs to inputs.
-
-To make sense of the large amounts of data, Macko et.
-al.
-have developed Orbiter, a tool to visualize provenance graphs with semantic zoom \cite{orbiter}.
-Margo et.
-al.
-have developed techniques to analyze large provenance graphs and to extract useful information from these graphs, including semantic file attributes \cite{fileattributes}.
-Using simple provenance statistics on nodes (e.g.
-neighbors, file count, process count, edge count) and features from \texttt{stat}, they determine with high accuracy the type of files (e.g.
-source files, configuration files).
-Furthermore, they have identified key properties of provenance graphs and developed new centrality metrics to perform local clustering of graphs, separating the huge provenance graphs into smaller semantic tasks or workloads \cite{clustering}.
-
-Applications that require provenance collection typically place great emphasis on data integrity.
-Given this priority, a natural goal of provenance systems is to detect intrusions on the system.
-Somayaji et.
-al.
-have developed techniques for detecting intrusions based on system call sequences, and their system counteracts these intrusions by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
-These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
-Somayaji's work relies on building a {\em normal} profile of a process, and then detecting when the process deviates from normal.
-
-Tariq et.
-al.
-generalize Somayaji's sequences of system calls to sequences of system events \cite{correlated-anomalies} to develop an intrusion detection system (IDS) in a distributed system.
-Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of systemevents--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
-King et.
-al.
-have developed Backtracker \cite{backtracker}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
-Given an intrusion, Backtracker builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtracker with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
-However, their system does not have a way of detecting an intrusion.
-Lei et.
-al.
-build similar process-file dependency trees, which they call access trees, and define a compatibility function to compare access trees with the goal of predicting future file accesses based on the current access tree pattern \cite{fileprefetch}.
-
-Despite the existing work to detect intrusions and anomalies, Cao et.
-al.
-argue that any intrusion detection system cannot always distinguish good behaviors from bad behaviors because users behave too randomly to have an accurate model of the user's behavior \cite{fuzzy}.
-They propose their fuzzy anomaly detection approach to extend the training models in existing intrusion detection systems to account for the inherent fuzzy nature of intrusion detection.
-When training an IDS, instead of assigning hard 0s or 1s for features, their approach provides a way to reliably map these features to weights in the interval $[0,1]$.
-
-Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies and existing work in intrusion detection that uses these kinds of directed graphs of files and process data, it seems plausible to detect intrusions based on provenance data collected by PASS.
-In this paper, we present a histogram-based technique for analyzing existing PASS provenance data to detect if an intrusion occurred.
-
-In particular, we generate statistical models of normal behavior, against which we can test a node of interest.
-These statistical models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
-The lower the kernel density estimate, the less ``normal" the node.
-The contributions of this paper are:
-\begin{enumerate}
-\item A histogram-based technique for using provenance data to detect intrusions
-\item An analysis of the different intrusions that can be identified, and the types of provenance which best contribute to these detections
-\item Future extensions to our work to create better intrusion detection models and make more use of provenance data
-\end{enumerate}
-
-\section{Background and Related Works}
-\subsection{PASS}
-Provenance data is met-data detailing the creation, transformation, and flow of data.
-PASS \cite{pass, passv2} is a provenance-aware storage system, designed to generate and track this provenance data seamlessly, and Table \ref{passobjects} indicates the types of objects the PASS tracks and attributes with those objects.
+Table \ref{passobjects} indicates the types of objects the PASS tracks and attributes with those objects.
 PASS allocates pnode (``PASS" node) number and version numbers whenever PASS encounters a new object or a modification to an existing object.
 The combination of (pnode number, version number) uniquely identifies a node in our provenance graph.
+There are two versions of PASS.
+We conducted our experiments only with PASSv2; thus in this paper ``PASS" to refer to PASSv2.
 
-The PASS system automatically creates and updates provenance data for every file that was created or modified the volume.
-For example, the provenance for a data file that sent in a zip file would reveal both the name of the program that unzipped the file, as well as the web browser through which the zip file was downloaded!
-It is our belief that this level of specificity will lend itself to detecting patterns in program usage.
-
-There are two versions of PASS. We conducted our experiments only with PASSv2; thus in this paper ``PASS" to refer to PASSv2.
-
-% TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
-
-\begin{table}[htdp]
+\begin{table}[H]
 \begin{center}
@@
 
+The PASS system automatically creates and updates provenance data for every file that was created or modified the volume.
+For example, the provenance for a data file that sent in a zip file would reveal both the name of the program that unzipped the file, as well as the web browser through which the zip file was downloaded!
+It is our belief that this level of specificity will lend itself to detecting patterns in program usage.
+
+% TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
+
+To make sense of the large amounts of data, Macko et.
+al.
+have developed Orbiter, a tool to visualize provenance graphs with semantic zoom \cite{orbiter}.
+Margo et.
+al.
+have developed techniques to analyze large provenance graphs and to extract useful information from these graphs, including semantic file attributes \cite{fileattributes}.
+Using simple provenance statistics on nodes (e.g.
+neighbors, file count, process count, edge count) and features from \texttt{stat}, they determine with high accuracy the type of files (e.g.
+source files, configuration files).
+Furthermore, they have identified key properties of provenance graphs and developed new centrality metrics to perform local clustering of graphs, separating the huge provenance graphs into smaller semantic tasks or workloads \cite{clustering}.
+
 \subsection{Intrusion Detection}
+
+Applications that require provenance collection typically place great emphasis on data integrity.
+Given this priority, a natural goal of provenance systems is to detect intrusions on the system.
 Intrusion detection in live systems is a common problem in the systems community.
@@
 
-Somayaji et al.
-\cite{somayaji} aim to detect intrusions by keeping track of all systems calls that occur during a process's execution time.
-After collecting this data over a long period of "normal" time, if a proces exhibits system call patterns that deviate from the ``normal" patterns for the process, their system flags it as potentially dangerous, and slows its execution.
+Somayaji et.
+al.
+have developed techniques for detecting intrusions based on system call sequences, and their system counteracts these intrusions by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
+Their system collects this data over a long period of "normal" time.
+If a process exhibits system call patterns that deviate from the ``normal" patterns for the process, their system flags it as potentially dangerous, and slows its execution.
 In this way, a program that continually deviates from normal behavior will be descheduled entirely.
+These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
+Somayaji's work relies on building a {\em normal} profile of a process, and then detecting when the process deviates from normal.
 
-Tariq et al.
-\cite{correlated-anomalies} extend the work of Somayaji, designing a system to run an IDS on a set of distributed nodes.
-They use the IDS of other researchers, distilling a programs execution into a set of $k$-tuples of system events used to identify normal behavior during the program's execution.
+Tariq et.
+al.
+generalize Somayaji's sequences of system calls to sequences of system events \cite{correlated-anomalies} to develop an IDS in a distributed system.
+Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of systemevents--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
+King et.
+al.
+have developed Backtracker \cite{backtracker}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
+Given an intrusion, Backtracker builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtracker with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
+However, their system does not have a way of detecting an intrusion.
+Lei et.
+al.
+build similar process-file dependency trees, which they call access trees, and define a compatibility function to compare access trees with the goal of predicting future file accesses based on the current access tree pattern \cite{fileprefetch}.
+
+Despite the existing work to detect intrusions and anomalies, Cao et.
+al.
+argue that any intrusion detection system cannot always distinguish good behaviors from bad behaviors because users behave too randomly to have an accurate model of the user's behavior \cite{fuzzy}.
+They propose their fuzzy anomaly detection approach to extend the training models in existing intrusion detection systems to account for the inherent fuzzy nature of intrusion detection.
+When training an IDS, instead of assigning hard 0s or 1s for features, their approach provides a way to reliably map these features to weights in the interval $[0,1]$.
+
+\subsection{Our Goals and Contributions}
+
+Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies and existing work in intrusion detection that uses these kinds of directed graphs of files and process data, it seems plausible to detect intrusions based on provenance data collected by PASS.
+In particular, we hope to answer the questions:
+\begin{itemize}
+\item Given PASS provenance data, what kinds of intrusions can be detected?
+\item What provenance data is most useful for detecting these kinds of intrusions.
+\end{itemize}
+
 Our system is designed to run offline, so we run our algorithms on the full set of provenance data.
-While this approach is slower, it allows us to take in the full picture of the events during a process's execution, without having to preselect the information we think will be useful.
+While this approach is slower than running an online system like Somayaji and Tariq, it allows us to take in the full picture of the events during a process's execution, without having to preselect the information we think will be useful.
 This approach is especially helpful because different information leads to the detection of different types of intrusions.
+The data collected by Backtracker's dependency graph of process and files is very similar to the data PASS generates, and their exploits and analysis techniques directly inspired our work.
 
-Backtracker \cite{backtracker} is a system most similar to ours in design.
-It, like PASS, builds a directed graph of process-file dependencies while the system is running, and, like our system, analyzes it offline for potential intrusions.
-The data they collect is very similar to the data PASS generates, and their exploits and analysis techniques directly inspired our work.
+In this paper, we present a histogram-based technique for analyzing existing PASS provenance data to detect if an intrusion occurred.
+To detect intrusions, we generate statistical models of normal behavior, against which we can test a node of interest.
+These statistical models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
+The lower the kernel density estimate, the less ``normal" the node.
+
+We hypothesize that intrusions are deviations from normal behavior, and these deviations are likely to manifest themselves in different number of inputs,  outputs, and centrality counts.
+We chose to use histograms because histograms are easy and fast to make, and if this approach works well, can be turned into an online method for detecting intrusions as the system is running.
+
+The contributions of this paper are:
+\begin{enumerate}
+\item A histogram-based technique for using provenance data to detect intrusions
+\item An analysis of the different intrusions that can be identified, and the types of provenance which best contribute to these detections
+\item Future extensions to our work to create better intrusion detection models and make more use of provenance data
+\end{enumerate}
 
@@
 Simply an output file.
-Below, descandents of an exploited mcrypt-style program.
+Below, descendants of an exploited mcrypt-style program.
 In addition to an (offscreen) output file, there is also an outgoing \texttt{/bin/sh} process, with its own children.}
@@
 Our results for GCC are in Figures \ref{simple-gcc} and \ref{centrality-gcc}.
-We can see that proc\_out\_proc (outgoing processes), the ancestor centrality, and Opsahl Closness Centrality statistics have more variation, whereas the other statistics were unimodal or bimodal (0 or 1).
+We can see that proc\_out\_proc (outgoing processes), the ancestor centrality, and Opsahl Closeness Centrality statistics have more variation, whereas the other statistics were unimodal or bimodal (0 or 1).
 
@@
 
-\subsection{Other Centrality Metrics}
-Progressive Pagerank
-
 \subsection{Histograms by Name}
@@
 "forked") processes, we simply count them.
-Thus, if a program normally forks something benign, like say "cp", but an exploited version forks "rm", the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures).
+Thus, if a program normally forks something benign, like say \texttt{cp}, but an exploited version forks \texttt{rm}, the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures).
 One direction to attempt to rectify this problem lies in generating more specific histograms.
 Instead of having a model for "number of processes forked", we could generate one model per type of process forked, e.g.
-"number of 'cp' processes forked".
+"number of \texttt{cp} processes forked".
 This more specific data might enable us to catch intrusions in larger programs that already modify a lot of files and fork a lot of processes.
@@
 Currently we ignore the environment and argument information stored in the provenance graph.
-By generating histograms based on this information, we could catch subtler intrustions/exploits that rely on modified environment information.
+By generating histograms based on this information, we could catch subtler intrusions/exploits that rely on modified environment information.
 Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
@@
 We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values.
-Additionally, we demonstrated that odd behavior has a "low" KDE, but we did not define what "low" means.
+Additionally, we demonstrated that odd behavior has a ``low" KDE, but we did not define what ``low" means.
 For this system to be deployed more widely, we will need to define or compute what thresholds signify a potential intrusion or exploit.
@@
 We would also need to add additional daemons (or modify the existing PASS daemons) to run the appropriate statistic gathering and model generating code.
+Another extension is to apply progressive versions of our centrality metrics, including a progressive form of PageRank which computes the importance of a node online.
+\subsection{Local vs. Global}
+In our experiments, we looked at provenance graphs in their entirety.
+However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph.
+To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
 \section{Conclusion}

### 2013-11-24 writeup/paper.tex
typoe in construction

@@
 \subsection{Parameters}
-Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for construction the KDE.
+Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for constructing the KDE.
 We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values.

### 2013-11-24 writeup/paper.tex
Fixed figure numbers

@@
 \subsection{Examples}
-In Figure \ref{kde-example}, we see a few examples of a generated histogram and KDE model.
+In Figure 1, we see a few examples of a generated histogram and KDE model.
 The green bars represent the histogram for the particular statistic, and the blue curve is the KDE.
@@
 We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
-See Figure \ref{hello-orbiter} for provenance graphs displaying normal and exploited behavior of \texttt{hello}.
+See Figure 2 for provenance graphs displaying normal and exploited behavior of \texttt{hello}.
 
@@
 This is similar to recording all uses of a program during a server's lifetime.
-See Figure \ref{mcrypt-orbiter} displaying normal and exploited behavior for our version of mcrypt.
+See Figure 3 displaying normal and exploited behavior for our version of mcrypt.
 
@@
 Towards this end, we graphed all of the statistics over the ``normal" provenance data, to see where there were trends and variation.
-Our results for GCC are in Figures \ref{simple-gcc} and \ref{centrality-gcc}.
+Our results for GCC are in Figures 4 and 5.
 We can see that proc\_out\_proc (outgoing processes), the ancestor centrality, and Opsahl Closeness Centrality statistics have more variation, whereas the other statistics were unimodal or bimodal (0 or 1).
@@
 The clean nodes for ``hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked \texttt{/usr/bin/ls} process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendants (see Figure \ref{hello-orbiter} and Figure \ref{mcrypt-orbiter}).
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendants (see Figure 2 and Figure 3).
 Similarly, the normal GCC and exploited GCC displayed very different provenance graphs, but they were too big to fit legibly on a page.
@@
 \subsection{Statistical Data and KDE Predictions}
-Table \ref{results} summarizes some of the preliminary results we have generated.
+Table 1 summarizes some of the preliminary results we have generated.
 Here we have included, for each statistic and test, the average over the "normal" nodes, the observed value for the "test" node, and the KDE probability output for that value.

### 2013-11-25 writeup/paper.tex
False positive data

@@
 We wrote Python scripts to (1) parse the raw BerkeleyDB database files storing the PASS data (2) build the provenance graphs (3) collect the counts for each statistic (which include running the centrality calculations), and (4) optionally outputting the histograms and kernel density estimations for a node of interest in a second provenance graph.
+We should note here that some centrality calculations proved difficult to run efficiently on large provenance graphs, particularly the Eigenvector Centrality mentioned above.
+As such, it is not included in our below analyses.
 
@@
 
-\subsection{Results}
+\section{Results}
 \subsection{Provenance Data}
@@
 \subsection{Statistical Data and KDE Predictions}
-Table 1 summarizes some of the preliminary results we have generated.
+Table 2 summarizes some of the preliminary results we have generated.
 Here we have included, for each statistic and test, the average over the "normal" nodes, the observed value for the "test" node, and the KDE probability output for that value.
@@
      \# input files & 0.5 | 0 | 0.453 & 0.5 | 0 | 0.453 & 0.5 | 1 | 0.107 \\ \hline
-     \# ouptut files & 0.5 | 1 | 0.453 & 0.5 | 1 | 0.453 & 0.0 | 3 | 0.000 \\ \hline
-    Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\
+     \# output files & 0.5 | 1 | 0.453 & 0.5 | 1 | 0.453 & 0.0 | 3 | 0.000 \\ \hline
+    Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\ \hline
+    %Ancestor centrality & 0.0009 | 0.0011 | 0.799 & 0.0002 | 0.0014 | 0.798 & 0.0007 | 0.0002 | 0.798 \\
+
     \hline
@@
 \end{table*}
+\subsection{Discrimination/False Positives}
+We were initially concerned that while our methods would reveal exploited nodes, they would also flag ``normal'' nodes as potential exploits.
+To determine whether this was the case, we ran our analyses on the ``normal'' nodes.
+To determine the worst possible outcome, here we report the minimum KDE value that a ``good'' node produced for a given test.
+If this value is significantly lower than the value we saw for a ``bad'' node, we can be reasonably confident in the power of that test.
+This method should also be investigated as a means for automatically determining thresholds for ``bad'' nodes.
 
-\subsection{Discussion}
+In Table 3, we report, for each exploit, the minimum KDE value we saw for a ``normal'' node, along with the number of nodes examined.
+Comparing these values to the values shown in Table 2 confirms our belief that these metrics can distinguish between ``normal'' and abnormal nodes.
+One exception is that of outgoing processes for gcc, where the minimum node was far below even the exploit node.
+The node in question only had one outgoing process, where most nodes (as shown in Figure 1) had 0 or 2.
+As shown in the histogram, there were nodes with this odd value.
+The node in question had normal values for the other metrics, however, so a combined view would not have necessarily flagged the node as abnormal.
+These combined metrics must be developed so as to minimize false positives and maximize true positives.
+\begin{table*}[ht]
+\label{results}
+{\small
+  \begin{center}
+  \begin{tabular}{| l | l | l | l |}
+    \hline
+    metric & hello (80 nodes) & mcrypt (424) & gcc (154) \\ \hline
+     \# forked processes & 0.453 & 0.798 & 0.166\\ \hline
+     \# input processes &  0.617 & 0.453 & 0.453\\ \hline
+
+     \# input files & 0.453 & 0.453 & 0.798\\ \hline
+     \# output files &  0.453 & 0.798 & 0.798\\ \hline
+
+     \# input pipes & 0.798 & 0.798 & 0.798\\ \hline
+     \# output pipes &  0.798 & 0.798 & 0.798\\ \hline
+
+    \hline
+  \end{tabular}
+  \end{center}
+}
+\hfill{}
+\caption{minimum kernel density estimate of test node being ``normal" seen on a ``normal'' node. With few exceptions, these values are high.
+}
+\label{tb:tablename}
+\end{table*}
+
+\section{Discussion}
 \subsection{Successes}

### 2013-11-25 writeup/paper.tex
Overlaid KDE in summary graphs

@@
   \label{simple-gcc}
-  \caption{Histograms over all "simple" provenance statistics for GCC}
+  \caption{Histograms over all "simple" provenance statistics for GCC, with KDE curve}
   \centering
-    \includegraphics[width=0.7\textwidth]{img/gcc_stats.png}
+    \includegraphics[width=0.7\textwidth]{img/gcc_stats2.png}
 \end{figure*}
@@
   \label{centrality-gcc}
-  \caption{Histograms over all centrality metrics for GCC}
+  \caption{Histograms over all centrality metrics for GCC, with KDE curve}
   \centering
-    \includegraphics[width=0.7\textwidth]{img/gcc_cent_stats.png}
+    \includegraphics[width=0.7\textwidth]{img/gcc_cent_stats2.png}
 \end{figure*}

### 2013-11-25 make_plot2.py

@@ -33,20 +33,13 @@ ls = {"proc_in_file" :[0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
 
 #ls['opsahl'] = [0.0, 0.0, 6.185714285714286, 5.426190476190476, 0.0, 0.0, 0.0, 3.583333333333333, 0.0, 0.0, 5.316666666666666, 6.185714285714286, 5.116666666666667, 6.510714285714286, 6.185714285714286, 0.0, 0.0, 0.0, 5.116666666666667, 0.0, 3.583333333333333, 0.0, 0.0, 5.116666666666667, 0.0, 0.0, 0.0, 6.185714285714287, 0.0, 0.0, 3.5833333333333335, 5.426190476190476, 3.5833333333333335, 0.0, 0.0, 0.0, 6.185714285714286, 6.185714285714286, 0.0, 3.5833333333333335, 5.4261904761904765, 0.0, 3.5833333333333335, 5.116666666666666, 6.185714285714286, 0.0, 0.0, 5.116666666666667, 0.0, 0.0, 0.0, 5.116666666666667, 6.185714285714286, 0.0, 0.0, 0.0, 0.0, 0.0, 3.5833333333333335, 5.116666666666667, 6.185714285714287, 4.166666666666666, 5.483333333333333, 0.0, 5.262301587301588, 3.583333333333333, 0.0, 0.0, 6.510714285714285, 6.185714285714286, 5.116666666666667, 0.0, 0.0, 0.0, 0.0, 6.286904761904762, 6.510714285714285, 6.185714285714286, 5.116666666666667, 0.0, 6.973809523809527, 0.0, 5.0511904761904765, 0.0, 6.185714285714286, 0.0, 5.116666666666666, 5.116666666666667, 5.116666666666666, 6.185714285714285, 0.0, 0.0, 6.788492063492065, 0.0, 0.0, 6.185714285714286, 0.0, 0.0, 5.933333333333334, 0.0, 6.510714285714286, 6.185714285714286, 5.116666666666666, 0.0, 6.510714285714286, 6.185714285714286, 0.0, 0.0, 4.533333333333333, 0.0, 0.0, 0.0, 9.054761904761905, 6.510714285714286, 6.045238095238096, 6.185714285714286, 0.0, 5.116666666666666, 0.0, 5.426190476190476, 0.0, 5.116666666666667, 0.0, 
[…]

### 2013-11-25 writeup/paper.tex
Added ancestor to chart

@@
 \item Given PASS provenance data, what kinds of intrusions can be detected?
-\item What provenance data is most useful for detecting these kinds of intrusions.
+\item What provenance data is most useful for detecting these kinds of intrusions?
 \end{itemize}
@@
 We have chosen four metrics that provided a statistically significant result.
+Additionally, we've included Ancestor centrality, which from Figure 5's distribution, should have been a good metric as well.
+However, as seen in Figure 5, the KDE curve does not really follow the distribution.
+This deviation is because the bandwidth parameter we used for the KDE is much larger than the differences between binds for the ancestor metric.
+Either scaling up our metric scores or choosing a smaller bandwidth could alleviate this result.
 % The other statistics failed to be as powerful (see Table 2 XXX).
@@
     Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\ \hline
-    %Ancestor centrality & 0.0009 | 0.0011 | 0.799 & 0.0002 | 0.0014 | 0.798 & 0.0007 | 0.0002 | 0.798 \\
+    Ancestor centrality & 0.0009 | 0.0011 | 0.799 & 0.0002 | 0.0014 | 0.798 & 0.0007 | 0.0002 | 0.798 \\
 

### 2013-11-25 writeup/paper.tex
Spellcheck

@@
 
-To make sense of the large amounts of data, Macko et.
-al.
+To make sense of the large amounts of data, Macko et al.
 have developed Orbiter, a tool to visualize provenance graphs with semantic zoom \cite{orbiter}.
-Margo et.
-al.
+Margo et al.
 have developed techniques to analyze large provenance graphs and to extract useful information from these graphs, including semantic file attributes \cite{fileattributes}.
@@
 
-Somayaji et.
-al.
+Somayaji et al.
 have developed techniques for detecting intrusions based on system call sequences, and their system counteracts these intrusions by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
@@
 
-Tariq et.
-al.
+Tariq et al.
 generalize Somayaji's sequences of system calls to sequences of system events \cite{correlated-anomalies} to develop an IDS in a distributed system.
-Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of systemevents--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
-King et.
-al.
+Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of system events--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
+King et al.
 have developed Backtracker \cite{backtracker}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
@@
 However, their system does not have a way of detecting an intrusion.
-Lei et.
-al.
+Lei et al.
 build similar process-file dependency trees, which they call access trees, and define a compatibility function to compare access trees with the goal of predicting future file accesses based on the current access tree pattern \cite{fileprefetch}.
 
-Despite the existing work to detect intrusions and anomalies, Cao et.
-al.
+Despite the existing work to detect intrusions and anomalies, Cao et al.
 argue that any intrusion detection system cannot always distinguish good behaviors from bad behaviors because users behave too randomly to have an accurate model of the user's behavior \cite{fuzzy}.
@@
 \subsection{Centrality Metrics}
-In addition to the data that can be collected on the nodes directly, we computed centrality metrics on the nodes using the centrality metrics proposed by Macko et.
-al.
+In addition to the data that can be collected on the nodes directly, we computed centrality metrics on the nodes using the centrality metrics proposed by Macko et al.
 \cite{clustering}.
@@
 To emulate this exploit, we designed a Python program, which would decrypt (via ROT13) an input file.
-When sent a particular exploit fileâ which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
+When sent a particular exploit file which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
 
@@
 The clean nodes for ``hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked \texttt{/usr/bin/ls} process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendants (see Figure 2 and Figure 3).
+Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendant normally, but when exploited has multiple extra processes/files as descendants (see Figure 2 and Figure 3).
 Similarly, the normal GCC and exploited GCC displayed very different provenance graphs, but they were too big to fit legibly on a page.

### 2013-12-09 writeup/paper-final.tex
rewrote abstract and title

--- 2013-11-25 writeup/paper.tex
@@
 
-\title{{\bf Intrusion Detection through Provenance-Based Histograms}}
+\title{{\bf Intrusion Detection using Provenance Graph Centrality Parzen-Windows}}
 \author{
@@
 \begin{abstract}
-Data-driven intrusion detection is a common system security technique.
-We present a histogram-based approach to intrusion detection, using provenance data generated by a provenance-aware storage system.
-By looking at simple provenance statistics and centrality metrics on provenance graphs, our approach identifies suspicious behavior based on deviations from prior normal system behavior.
-Our histogram-based approach requires no prior intuition about the types of exploits potentially being run on a system, and our analysis of two buffer overflow attacks and a modified version of \texttt{gcc} to log actions reveal which data is most useful in distinguishing between safe and malicious programs and program executions.
+Provenance is data that describes how a digital artifact came to be in its current state. We hypothesize
+that intrusions on a system leave behind anomalies in the lineage of digital artifacts.
+We present an intrusion detection approach to find these anomalies by analyzing centrality metrics
+on provenance graphs. We use a Parzen-Window approach (TODO CITE) on various provenance graph centrality metrics (TODO CITE)
+to determine probability density estimates of normal behavior, and we use these density estimates to 
+determine if an intrusion occurred. We used this approach to analyze {\em user-to-remote} (u2r) intrusions and 
+{\em remote-to-local} (r2l) intrusions (TODO: include r2l?) from the 1998 DARPA Intrusion Detection data sets (TODO CITE) and 
+achieved up to *TODO true positive rate for intrusions* accuracy in detecting
+intrusions with only *TODO false positive rate for intrusions* accuracy. We also present future work
+to extend our intrusion model to an online intrusion detection system.
+
 \end{abstract}
@@
 
-\section{Introduction and Related Work}
+\section{Introduction}
 
-\subsection{Provenance and PASS}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% RELATED WORK
+%
 
-Provenance is metadata that tracks the history of all changes to files.
-In provenance-aware storage systems like PASS \cite{pass} and PASSv2 \cite{passv2}, the file system automatically tracks all dependencies whenever a file is created, modified, or deleted.
-These dependencies include the command that was executed to modify the file, the environment of the command, and all input files.
-The generated provenance forms a directed acyclic graph of typed nodes (e.g.
-files, processes, pipes) with properties (e.g.
-name, execution time, pid) and typed edges (e.g.
-forked, input, versioning).
-The edges show dependency, not data flow, and as a result, edges are directed from outputs to inputs.
-Table \ref{passobjects} indicates the types of objects the PASS tracks and attributes with those objects.
-PASS allocates pnode (``PASS" node) number and version numbers whenever PASS encounters a new object or a modification to an existing object.
-The combination of (pnode number, version number) uniquely identifies a node in our provenance graph.
-There are two versions of PASS.
-We conducted our experiments only with PASSv2; thus in this paper ``PASS" to refer to PASSv2.
-
-\begin{table}[H]
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline
-{\bf Node Types} & {\bf Description} \\
-\hline
-PROC & process \\
-FILE & file \\
-PIPE & pipe \\
-NPFILE & non-PASS file \\
-OTHER & other types of objects \\
-\hline
-{\bf Node Attributes} & {\bf Description} \\
-\hline
-INODE & inode of file \\
-PATH & path of file \\
-FREEZETIME & timestamp of record \\
-CREAT & creation of file \\
-UNLINK & unlinking file \\
-EXECTIME & exec time of process\\
-NAME & name of process \\
-PID & pid of of process \\
-ARGV & argument vector to process \\
-ENV & environment of a process \\
-\hline
-{\bf Edge Types} & {\bf Description} \\
-\hline
-FORKPARENT & indicates parent PROC\\
-&  of existing PROC node \\
-INPUT & indicates input files\\
-& or processes \\
-VERSION & indicates a modification  \\
-& to an existing node (e.g. \\
-& modifying a file) \\
-\hline
-\end{tabular}
-\end{center}
-\label{passobjects}
-\caption{Types of data collected by PASS}
-\end{table}%
-
-The PASS system automatically creates and updates provenance data for every file that was created or modified the volume.
-For example, the provenance for a data file that sent in a zip file would reveal both the name of the program that unzipped the file, as well as the web browser through which the zip file was downloaded!
-It is our belief that this level of specificity will lend itself to detecting patterns in program usage.
-
-% TODO: The PASS system runs with minimal time overhead, and a reasonable space overhead (XXX cite real numbers).
-
-To make sense of the large amounts of data, Macko et.
-al.
-have developed Orbiter, a tool to visualize provenance graphs with semantic zoom \cite{orbiter}.
-Margo et.
-al.
-have developed techniques to analyze large provenance graphs and to extract useful information from these graphs, including semantic file attributes \cite{fileattributes}.
-Using simple provenance statistics on nodes (e.g.
-neighbors, file count, process count, edge count) and features from \texttt{stat}, they determine with high accuracy the type of files (e.g.
-source files, configuration files).
-Furthermore, they have identified key properties of provenance graphs and developed new centrality metrics to perform local clustering of graphs, separating the huge provenance graphs into smaller semantic tasks or workloads \cite{clustering}.
-
-\subsection{Intrusion Detection}
-
-Applications that require provenance collection typically place great emphasis on data integrity.
-Given this priority, a natural goal of provenance systems is to detect intrusions on the system.
-Intrusion detection in live systems is a common problem in the systems community.
-There are two primary components to an intrusion detection system (IDS): (1) the data used to fuel the system's analysis, and (2) how the data is used.
-
-Somayaji et.
-al.
-have developed techniques for detecting intrusions based on system call sequences, and their system counteracts these intrusions by exponentially delaying or aborting system calls, rendering the system useless for a malicious attacker \cite{somayaji}\cite{somayaji-recent}.
-Their system collects this data over a long period of "normal" time.
-If a process exhibits system call patterns that deviate from the ``normal" patterns for the process, their system flags it as potentially dangerous, and slows its execution.
-In this way, a program that continually deviates from normal behavior will be descheduled entirely.
-These intrusions include exploiting vulnerabilities in the SSH daemon and sendmail to obtain a shell with root privileges.
-Somayaji's work relies on building a {\em normal} profile of a process, and then detecting when the process deviates from normal.
-
-Tariq et.
-al.
-generalize Somayaji's sequences of system calls to sequences of system events \cite{correlated-anomalies} to develop an IDS in a distributed system.
-Their system uses bloom filters to store hashes of $k$-tuples--representing a sequence of systemevents--and then use these bit vectors to correlate anomalies across multiple hosts back to the associated provenance.
-King et.
-al.
-have developed Backtracker \cite{backtracker}, a modified Linux kernel that tracks dependencies between operating system objects (files, processes, file names) in a similar fashion to PASS.
-Given an intrusion, Backtracker builds a backwards casual graph to determine the entry point of an intrusion, and they have extended Backtracker with forward causal graphs to determine possibly tainted files, processes, and hosts from an intrusion in a distributed system \cite{multihost}.
-However, their system does not have a way of detecting an intrusion.
-Lei et.
-al.
-build similar process-file dependency trees, which they call access trees, and define a compatibility function to compare access trees with the goal of predicting future file accesses based on the current access tree pattern \cite{fileprefetch}.
-
-Despite the existing work to detect intrusions and anomalies, Cao et.
-al.
-argue that any intrusion detection system cannot always distinguish good behaviors from bad behaviors because users behave too randomly to have an accurate model of the user's behavior \cite{fuzzy}.
-They propose their fuzzy anomaly detection approach to extend the training models in existing intrusion detection systems to account for the inherent fuzzy nature of intrusion detection.
-When training an IDS, instead of assigning hard 0s or 1s for features, their approach provides a way to reliably map these features to weights in the interval $[0,1]$.
-
-\subsection{Our Goals and Contributions}
-
-Given the extensive amount of data PASS collects on file-file, file-process, and process-process dependencies and existing work in intrusion detection that uses these kinds of directed graphs of files and process data, it seems plausible to detect intrusions based on provenance data collected by PASS.
-In particular, we hope to answer the questions:
-\begin{itemize}
-\item Given PASS provenance data, what kinds of intrusions can be detected?
-\item What provenance data is most useful for detecting these kinds of intrusions.
-\end{itemize}
-
-Our system is designed to run offline, so we run our algorithms on the full set of provenance data.
-While this approach is slower than running an online system like Somayaji and Tariq, it allows us to take in the full picture of the events during a process's execution, without having to preselect the information we think will be useful.
-This approach is especially helpful because different information leads to the detection of different types of intrusions.
-The data collected by Backtracker's dependency graph of process and files is very similar to the data PASS generates, and their exploits and analysis techniques directly inspired our work.
-
-In this paper, we present a histogram-based technique for analyzing existing PASS provenance data to detect if an intrusion occurred.
-To detect intrusions, we generate statistical models of normal behavior, against which we can test a node of interest.
-These statistical models are histograms of different graph statistics, and we use these histograms to generate kernel density estimations to evaluate a test node.
-The lower the kernel density estimate, the less ``normal" the node.
-
-We hypothesize that intrusions are deviations from normal behavior, and these deviations are likely to manifest themselves in different number of inputs,  outputs, and centrality counts.
-We chose to use histograms because histograms are easy and fast to make, and if this approach works well, can be turned into an online method for detecting intrusions as the system is running.
-
-The contributions of this paper are:
-\begin{enumerate}
-\item A histogram-based technique for using provenance data to detect intrusions
-\item An analysis of the different intrusions that can be identified, and the types of provenance which best contribute to these detections
-\item Future extensions to our work to create better intrusion detection models and make more use of provenance data
-\end{enumerate}
+\section{Related Work}
 
@@
 \section{Design and Implementation}
-Here we present our intrusion detection technique: a general and extendable statistical method for analyzing and comparing provenance data.
-At a high level, we create a provenance data graph and generate a model of ``normal" behavior for each process/file in the graph.
-Using this model, we can determine the likelihood that a given ``test" deviates from normal behavior and is thus potentially an exploited or malicious program.
-\subsection{Models}
-Our approach requires statistical models to evaluate the normalcy of a test node.
-The statistical model we chose to use was a kernel density estimation (KDE), which estimates the probability density function of a value with unknown distribution.
-To create the KDE, we needed to provide counts of a particular statistic.
-For each statistic:
-\begin{enumerate}
-\item We computed the counts for that statistic over all the nodes in the provenance graph.
-\item We aggregated the counts by the NAME field of the node (e.g.
-the path to a file for a FILE node, or path to an executable for PROC node).
-For objects with no NAME field (e.g.
-PIPE objects), we used the tuple of (pnode number, VERSION number) as the node's name.
-\item For each unique name in the graph, we use the counts on that name to generate a histogram of ``normal" behavior.
-We convert these histograms into kernel density objects to smoothen our histogram and generalize to all continuous values for the statistic.
-\end{enumerate}
 
-This model allows us to evaluate a test node's normalcy according to that statistic.
-For example, we could then ask ``Does node $X'$ have a statistically believable number of output processes compared to the other $X$ nodes?"
-
-\subsection{Provenance Statistics}
-We expected intrusions to differ in terms of the statistics that would stand out, and as a result, we needed to generate a diverse suite of statistics and a model for each.
-For our simple provenance graph statistics, we compiled the following statistics:
-\begin{enumerate}
-\item The number of input files to the node.
-\item The number of input processes to the node.
-\item The number of input pipes to the node.
-\item The number of output files to the node.
-\item The number of output processes to the node.
-\item The number of output pipes to the node.
-\item The number of VERSION edges into this node.
-\item The number of VERSION edges out of this node.
-\end{enumerate}
-For ease of computation and comparison, we treated the different node types (e.g.
-file, process, pipe) separately so that we can treat, for instance, "input files to a file node" and "input files to a process node" as separate counts.
-
-\subsection{Centrality Metrics}
-In addition to the data that can be collected on the nodes directly, we computed centrality metrics on the nodes using the centrality metrics proposed by Macko et.
-al.
-\cite{clustering}.
-In their work, they noted 3 properties of provenance graphs:
-\begin{enumerate}
-\item The ubiquity (importance) of a node is a function of a node's descendants, not its ancestors.
-(Note: Because edges in provenance graphs are in the direction of data dependence instead of data flow, ``descendants" in this case means data flow descendants and NOT descendants by traversing provenance graph edges).
-\item There is no agreed-upon granularity in which provenance should be captured.
-\item Provenance has a temporal component.
-\end{enumerate}
-
-Using these observations, they propose several centrality metrics on provenance graphs:
-\begin{itemize}
-\item \textbf{In-degree Centrality}.
-Because a node's importance is based on its output descendants, then the number of times a node was used as input to another node indicates a simple measure of importance.
-\item \textbf{Age}. They note that large "jumps" in timestamps or age might indicate breaks between tasks, but this interpretation is complicated when scripts execute tasks in short succession.
-\item \textbf{Ancestor Centrality}. This measures the total number of descendants of a node, normalized by the total number of nodes in a graph.
-\item \textbf{Opsahl's Closeness Centrality}. This accounts for the number of edges between a node and its descendants. To compute this, we calculate
-$$CC'(v) = \sum_{x \in V -  \{v\}} (d'_{xv})^{-1}$$
-where $d'_{xv}$ is the distance from node $x$ to $v$ by only following outgoing edges from $x$. Unreachable nodes from $x$ will have $d'_{xv} = \infty$, and so $(d'_{xv})^{-1} = 0$.
-\item \textbf{Provenance Eigenvector Centrality}. The centrality of the nodes is given by the left dominant eigenvector of the matrix:
-$$M_{ij} =
-\begin{cases}
-1 & \text{if there is an edge } i \to j \\
-1/|V| & \text{if there is no out edge from } i \\
-0 & \text{otherwise}
-\end{cases}
-$$
-\end{itemize}
-
-We each metric, we computed the centrality of all the nodes and aggregated the counts by name to generate our histogram and kernel density estimations like the simple statistics we used above.
-
-\subsection{Implementation}
-We used Orbiter \cite{orbiter} to visualize and examine the provenance graphs.
-We wrote Python scripts to (1) parse the raw BerkeleyDB database files storing the PASS data (2) build the provenance graphs (3) collect the counts for each statistic (which include running the centrality calculations), and (4) optionally outputting the histograms and kernel density estimations for a node of interest in a second provenance graph.
-We should note here that some centrality calculations proved difficult to run efficiently on large provenance graphs, particularly the Eigenvector Centrality mentioned above.
-As such, it is not included in our below analyses.
-
-\subsection{Examples}
-In Figure 1, we see a few examples of a generated histogram and KDE model.
-The green bars represent the histogram for the particular statistic, and the blue curve is the KDE.
-Finally, the red line is the value for a ``test" node.
-In this case, the test node appears to be a possible exploit, because it intersects the KDE at low values.
-\begin{figure*}
-  \label{kde-example}
-  \caption{Example histograms of a centrality measure and process statistic for the \texttt{gcc} exploit, with the associated KDE.
-The green bars represent the raw histogram counts, the blue curve represents the kernel density function, and the red vertical line represents the estimation of a node of interest.}
-  \centering
-    \includegraphics[width=\textwidth]{img/hist.png}
-\end{figure*}
+\subsection{Selecting Metrics}
 
@@
 
-\subsection{Evaluation Methodology}
-In order to test our techniques, we needed to run exploits on our PASS-enabled machine.
-We looked to prior work to get a sense of the different types of techniques commonly used to exploit security vulnerabilities.
-Commonly, remote network techniques (as seen in Backtracker \cite{backtracker}) are used to breach insecure systems.
-Metasploit \cite{metasploit} is a software package designed to facilitate running exploits on insecure system.
-Searching the Metasploit database, as well as the online Exploit Database \cite{exploitdb}, we found that most successful techniques involve exploiting vulnerabilities in software that a target system is running, as opposed to vulnerabilities in the target system itself.
-Put another way, we sought to exploit software packages, like mcrypt and gcc, not the Linux kernel itself.
+\subsection{Experimental Setup}
 
-To generate the provenance data necessary to run the analyses, we needed to run both good and exploited versions of our programs.
+\subsection{Results}
 
-\subsubsection{hello}
-We found that many of the popular exploits involved buffer overflows and arbitrary code execution.
-Programming bugs in a piece of software allowed a malicious user, sending well-crafted data, to run arbitrary code on the target system.
-We created a simple \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a process.
-
-We ran \texttt{hello} normally 40 times on various inputs and collected the provenance times.
-We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
-See Figure 2 for provenance graphs displaying normal and exploited behavior of \texttt{hello}.
-
-\subsubsection{mcrypt}
-The mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
-Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like the authors of Backtracker \cite{backtracker} found.
-To emulate this exploit, we designed a Python program, which would decrypt (via ROT13) an input file.
-When sent a particular exploit fileâ which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
-
-We ran our program on over 200 good inputs, and stored the provenance data.
-We then ran our program on two bad inputs, first to call \texttt{ls} as a Proof of Concept, and again to call \texttt{/bin/sh} to demonstrate that more severe attacks can be performed.
-We again exported the provenance generated by these steps, for comparison with our good usage.
-This is similar to recording all uses of a program during a server's lifetime.
-See Figure 3 displaying normal and exploited behavior for our version of mcrypt.
-
-\subsubsection{gcc}
-Another type of exploit is more reminiscent of a system administration failure.
-If a malicious user replaces a program in \texttt{/usr/bin}, it could be used to log user actions.
-To emulate this sort of attack, we replaced \texttt{/usr/bin/gcc} with our own version of gcc, which, in addition to compiling the user's code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
-
-For our system administration attack, we compiled the PASS toolset using GCC, and exported the provenance from these runs.
-We then replaced GCC with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
-In this way, we collected provenance for normal usage and from our exploited usage.
-\begin{figure*}
-  \label{hello-orbiter}
-  \caption{Descendants of a \texttt{hello} program.
-The top image shows multiple invocations of a \texttt{hello} behaving normally when invoked with \texttt{cat foo.txt | ./hello > c.txt}.
-The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.}
-  \centering
-    \includegraphics[width=\textwidth]{img/hello.png}
-\end{figure*}
-\begin{figure*}
-  \label{mcrypt-orbiter}
-  \caption{Descendants of an unexploited mcrypt-style program.
-Simply an output file.
-Below, descendants of an exploited mcrypt-style program.
-In addition to an (offscreen) output file, there is also an outgoing \texttt{/bin/sh} process, with its own children.}
-  \centering
-    \includegraphics[width=\textwidth]{img/mcrypt.png}
-\end{figure*}
-
-\subsection{Selecting Metrics}
-Because we had no prior intuition into what the histograms on provenance data would actually look like, we approached selecting statistics from an objective, data-based standpoint.
-We wanted to use statistics that had variation over our workload to minimize the dependency of our statistics on the workload data.
-Towards this end, we graphed all of the statistics over the ``normal" provenance data, to see where there were trends and variation.
-Our results for GCC are in Figures 4 and 5.
-We can see that proc\_out\_proc (outgoing processes), the ancestor centrality, and Opsahl Closeness Centrality statistics have more variation, whereas the other statistics were unimodal or bimodal (0 or 1).
-
-The unimodal and bimodal statistics would not generate useful kernel density estimates, as they would predict roughly 0.5 for either a 0 or 1 count, giving us no information about whether the node in question is behaving normally.
-On the other hand, the statistics with more variation in data would actually provide more useful information when determining normalcy in new test nodes.
-As we include more data (as discussed in Future Work), these sorts of graphs will become increasingly important in selecting good statistics for evaluation.
-\begin{figure*}
-  \label{simple-gcc}
-  \caption{Histograms over all "simple" provenance statistics for GCC, with KDE curve}
-  \centering
-    \includegraphics[width=0.7\textwidth]{img/gcc_stats2.png}
-\end{figure*}
-\begin{figure*}
-  \label{centrality-gcc}
-  \caption{Histograms over all centrality metrics for GCC, with KDE curve}
-  \centering
-    \includegraphics[width=0.7\textwidth]{img/gcc_cent_stats2.png}
-\end{figure*}
-
-\section{Results}
-\subsection{Provenance Data}
-After running the normal and exploited programs described above, we visualized the provenance graphs using Orbiter \cite{orbiter}.
-After searching for the nodes named after our programs, we were able to find both the clean and exploited nodes.
-The clean nodes for ``hello" simply read in one file through a pipe and outputted one file, whereas the exploited version clearly has a forked \texttt{/usr/bin/ls} process.
-Similarly, the mcrypt simulation (implemented in Python) has only an output file as a descendent normally, but when exploited has multiple extra processes/files as descendants (see Figure 2 and Figure 3).
-Similarly, the normal GCC and exploited GCC displayed very different provenance graphs, but they were too big to fit legibly on a page.
-
-\subsection{Statistical Data and KDE Predictions}
-Table 2 summarizes some of the preliminary results we have generated.
-Here we have included, for each statistic and test, the average over the "normal" nodes, the observed value for the "test" node, and the KDE probability output for that value.
-We have chosen four metrics that provided a statistically significant result.
-% The other statistics failed to be as powerful (see Table 2 XXX).
-% from http://hhh123.wordpress.com/2008/08/13/latex-table-across-two-columns/
-% TODO: include difference/ratio from normal?
-\begin{table*}[ht]
-\label{results}
-{\small
-  \begin{center}
-  \begin{tabular}{| l | l | l | l |}
-    \hline
-    metric & hello & mcrypt & gcc \\ \hline
-     \# forked processes & 0.0 | 1 | 0.108 & 0.0 | 1 | 0.107 & 0.9 | 0 | 0.408 \\ \hline
-     \# input files & 0.5 | 0 | 0.453 & 0.5 | 0 | 0.453 & 0.5 | 1 | 0.107 \\ \hline
-     \# output files & 0.5 | 1 | 0.453 & 0.5 | 1 | 0.453 & 0.0 | 3 | 0.000 \\ \hline
-    Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\ \hline
-    %Ancestor centrality & 0.0009 | 0.0011 | 0.799 & 0.0002 | 0.0014 | 0.798 & 0.0007 | 0.0002 | 0.798 \\
-
-    \hline
-  \end{tabular}
-  \end{center}
-}
-\hfill{}
-\caption{average over good | test node value | kernel density estimate of test node being ``normal"
-}
-\label{tb:tablename}
-\end{table*}
-\subsection{Discrimination/False Positives}
-We were initially concerned that while our methods would reveal exploited nodes, they would also flag ``normal'' nodes as potential exploits.
-To determine whether this was the case, we ran our analyses on the ``normal'' nodes.
-To determine the worst possible outcome, here we report the minimum KDE value that a ``good'' node produced for a given test.
-If this value is significantly lower than the value we saw for a ``bad'' node, we can be reasonably confident in the power of that test.
-This method should also be investigated as a means for automatically determining thresholds for ``bad'' nodes.
-
-In Table 3, we report, for each exploit, the minimum KDE value we saw for a ``normal'' node, along with the number of nodes examined.
-Comparing these values to the values shown in Table 2 confirms our belief that these metrics can distinguish between ``normal'' and abnormal nodes.
-One exception is that of outgoing processes for gcc, where the minimum node was far below even the exploit node.
-The node in question only had one outgoing process, where most nodes (as shown in Figure 1) had 0 or 2.
-As shown in the histogram, there were nodes with this odd value.
-The node in question had normal values for the other metrics, however, so a combined view would not have necessarily flagged the node as abnormal.
-These combined metrics must be developed so as to minimize false positives and maximize true positives.
-\begin{table*}[ht]
-\label{results}
-{\small
-  \begin{center}
-  \begin{tabular}{| l | l | l | l |}
-    \hline
-    metric & hello (80 nodes) & mcrypt (424) & gcc (154) \\ \hline
-     \# forked processes & 0.453 & 0.798 & 0.166\\ \hline
-     \# input processes &  0.617 & 0.453 & 0.453\\ \hline
-
-     \# input files & 0.453 & 0.453 & 0.798\\ \hline
-     \# output files &  0.453 & 0.798 & 0.798\\ \hline
-
-     \# input pipes & 0.798 & 0.798 & 0.798\\ \hline
-     \# output pipes &  0.798 & 0.798 & 0.798\\ \hline
-
-    \hline
-  \end{tabular}
-  \end{center}
-}
-\hfill{}
-\caption{minimum kernel density estimate of test node being ``normal" seen on a ``normal'' node. With few exceptions, these values are high.
-}
-\label{tb:tablename}
-\end{table*}
-
-\section{Discussion}
-\subsection{Successes}
-We were able to accurately identify our three exploited programs. Each was a slightly different exploit, but each exploit deviated enough on some metrics which flagged them as non-normal.
-
-\subsection{Useful Data for Intrusion Detection}
-We set out to identify both buffer overflows (i.e.
-arbitrary code execution) and modified binaries.
-The buffer overflows were best detected by the number of outgoing processes and the centrality measures.
-This makes sense, as arbitrary code execution will generate additional processes.
-Of the centrality metrics we implemented, the Opsahl Centrality metric did the best in distinguishing an exploited node from normally behaving nodes.
-Because the Opsahl Centrality metric accounts for the number of edges between a node and all of its descendants, this centrality metric is very sensitive to the addition of nodes.
-Exploits that have additional outputs (e.g.
-new processes and output files through a remote shell), thus generate a higher Opsahl centrality metric for the offending node.
-The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that the exploit will be caught by our model.
-
-If the exploit does directly not generate additional files, however, statistics like "number of output files" will fail to be statistically significant.
-Indeed, the two programs (hello and mcrypt) that did not generate additional files were not flagged by that measure.
-On the other hand, the GCC that was modified to generate log files was caught by this measure, as it generated many more files than a normal GCC process.
-
-\subsection{Confusing Data Patterns}
-At first, some of the data we collected seemed wrong.
-Why, for instance, does GCC seem to normally output zero files?
-Upon further inspection, we found that GCC forks processes for the linker/assembler that handles generating output, so GCC does not directly have output files.
-Similarly, when running a python script, the code file itself is not directly executed, it is read in by a python process, which then goes off and collects input, generates output, and forks additional processes.
-Thus, when running a python file, the file itself only has one node in the provenance graph, with many python process nodes accessing it.
-To run our analyses, we needed to manually identify this connection, and compute statistics for the python process, instead of the single code file node.
+\subsection{Discussion}
 
@@
 \section{Future Work}
-\subsection{N-depth Statistics}
-Currently, the statistics and counts we generate apply only to the test node itself.
-As mentioned above, some processes fork helper processes that take in input or generate output, which would be useful information to collect.
-In order to account for these sorts of actions, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
-We could experiment with aggregation techniques to determine how best to incorporate this information.
-Perhaps it is best to simply sum all neighbor statistics together, or perhaps this information calls for new histogram and model.
-By utilizing information about neighboring nodes, we hope to see an increase in accuracy and the ability to detect less blatant intrusions.
-
-Additionally, by using this approach, we could avoid the manual process identification mentioned above.
-On python files, for instance, N-depth statistics would compute statistics for the python process, which is always a direct neighbor of the code file we initially cared about.
-
-\subsection{Histograms by Name}
-Currently, our system compute statistics based on node type, but not node name.
-For instance, When we compute the outgoing (i.e.
-"forked") processes, we simply count them.
-Thus, if a program normally forks something benign, like say \texttt{cp}, but an exploited version forks \texttt{rm}, the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures).
-One direction to attempt to rectify this problem lies in generating more specific histograms.
-Instead of having a model for "number of processes forked", we could generate one model per type of process forked, e.g.
-"number of \texttt{cp} processes forked".
-This more specific data might enable us to catch intrusions in larger programs that already modify a lot of files and fork a lot of processes.
-\subsection{ENV and ARGV}
-Currently we ignore the environment and argument information stored in the provenance graph.
-By generating histograms based on this information, we could catch subtler intrusions/exploits that rely on modified environment information.
-Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
-\subsection{Parameters}
-Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for constructing the KDE.
-We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values.
-Additionally, we demonstrated that odd behavior has a ``low" KDE, but we did not define what ``low" means.
-For this system to be deployed more widely, we will need to define or compute what thresholds signify a potential intrusion or exploit.
-\subsection{Online Intrusion Detection}
-Currently, this system runs after-the-fact.
-We can identify intrusions that occurred on a machine by analyzing the data after it has been lifted from the target machine.
-Because PASS collect data in real time, however, our approach can be modified to implement online intrusion detection.
-To achieve this goal, we would need to only use metrics that can be updated without recomputing the entire statistic.
-In particular, some of the centrality measures do not meet this criterion.
-We would also need to add additional daemons (or modify the existing PASS daemons) to run the appropriate statistic gathering and model generating code.
-Another extension is to apply progressive versions of our centrality metrics, including a progressive form of PageRank which computes the importance of a node online.
-\subsection{Local vs. Global}
-In our experiments, we looked at provenance graphs in their entirety.
-However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph.
-To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
-\section{Conclusion}
-We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomalies in program behavior, which often signal a security intrusion or exploit.
-Using provenance-based histograms, we can construct models of the regular usage patterns for a particular program.
-When the intrusion or exploit results in behavior deviating from these models, its provenance allows us to flag it as a potential intrusion.
-Replaced binaries that acted significantly differently were especially easy to detect.
-Histograms on simple provenance statistics are good for identifying simple anomalies, but more complex statistics (see future work) may enable us to detect subtler intrusions.
 

### 2013-12-09 writeup/paper-final.tex
rewrite introduction with story

@@
 
-\title{{\bf Intrusion Detection using Provenance Graph Centrality Parzen-Windows}}
+\title{{\bf Intrusion Detection using Parzen-Windows on Provenance Graph Statistics}}
 \author{
@@
 \section{Introduction}
+
+For as long as systems have existed, there have been malicious users that attempt to exploit vulnerabilities
+in systems to gain unintended privileged access. As a result, system designers and 
+administrators place a large effort in securing systems and preventing intrusions. However, intrusions
+inevitably occur because of bugs or flaws within the system, and as a result, system administrators want to
+have some automatic way of detecting these intrusions.
+
+Intrusions often leave behind digital artifacts, e.g., unusual output or log files, a process executed
+with an unusual arguments or environment, unusual system call activity by a process. In addition to
+unusual digital artifacts, we hypothesize that
+intrusions also intrusions leave behind abnormalities
+in the {\em provenance} of these digital artifacts, the lineage of how these digital artifacts were created.
+
+Provenance is data that describes how digital artifacts
+came to be in their current state. Provenance data is typically structured as a directed acyclic graph with
+typed nodes and typed edges.
+Nodes typically include processes and files, and edges are directed from nodes to their dependencies. For example,
+process nodes have edges directed towards input file nodes and to the parent process's node, and
+file nodes have edges directed towards previous versions of the file and to the nodes of the
+processes that modified the file.
+
+Existing intrusion detection systems (IDS) have used provenance data only to a limited extent.
+??? person analyzes
+basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
+easier manual intrusion detection by a human (CITE WORK). Their work, however, does not present a way of using provenance
+graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
+intrusion. For example, ??? Tariq use provenance graphs to build causality graphs of intrusions: once they have
+determined an intrusion has occurred, they follow the causality graph backwards determine the original source of the attack, and 
+they follow the causality graph forwards to determine all the objects tainted by the attack (CITE WORK). However, they also
+do not present a way of using provenance data to automatically detect intrusions.
+
+INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
+
+In this paper, we present an intrusion detection approach using Parzen-Windows density estimation. ???? have used
+Parzen-Windows for intrusion detection with a high degree of success in identifying various intrusion types (CITE WORK).
+They justify using Parzen-Windows for
+{\em novelty detection} when access to novel or abnormal patterns is difficult to obtain, as is the case with
+intrusions. We borrow their Parzen-Windows approach to build models of normal behavior using 
+various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
+on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
+
+The main contributions of this paper are the following:
+\begin{enumerate}
+\item Discuss which provenance graph statistics we chose to use, how we chose them, and why we chose them.
+\item Analyze the Parzen-Window technique on a real intrusion detection data set and evaluate its performance.
+\item Discuss the limitations of the approach and present future work to transform the technique into an online intrusion detection system.
+\end{enumerate}
 

### 2013-12-09 writeup/paper-final.tex
wrote related work

@@
 
-Existing intrusion detection systems (IDS) have used provenance data only to a limited extent.
+Existing intrusion detection systems (IDS) use provenance data only to a limited extent.
 ??? person analyzes
@@
 graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
-intrusion. For example, ??? Tariq use provenance graphs to build causality graphs of intrusions: once they have
-determined an intrusion has occurred, they follow the causality graph backwards determine the original source of the attack, and 
-they follow the causality graph forwards to determine all the objects tainted by the attack (CITE WORK). However, they also
+intrusion. For example, Backtracker (CITE WORK) uses provenance graphs to build causality graphs of intrusions: once the
+system has
+determined an intrusion has occurred, the system follow the causality graph backwards determine the original source of the attack, and 
+then it follow the causality graph forwards to determine all the objects tainted by the attack. However, the authors 
 do not present a way of using provenance data to automatically detect intrusions.
@@
 
-In this paper, we present an intrusion detection approach using Parzen-Windows density estimation. ???? have used
-Parzen-Windows for intrusion detection with a high degree of success in identifying various intrusion types (CITE WORK).
-They justify using Parzen-Windows for
-{\em novelty detection} when access to novel or abnormal patterns is difficult to obtain, as is the case with
-intrusions. We borrow their Parzen-Windows approach to build models of normal behavior using 
+In this paper, we present an approach to use provenance data to automatically detect intrusions. Intrusion
+detection can be framed as a {\em novelty detection} problem, in which one attempts to decide whether
+an unknown test pattern is produced by an underlying distribution corresponding to a training set
+of normal patterns (CITE WORK). However, ???authors note that in the case of intrusion detection, novel
+or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach
+for non-parametric density estimation. Using this technique, they obtain a high degree of success
+in identifying various intrusion types (CITE WORK). Because obtaining abnormal patterns is difficult, 
+we borrow their Parzen-Window approach to build models of normal behavior using 
 various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique

### 2013-12-09 writeup/paper-final.tex
skeleton for related work

@@
 
-INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
+>>> INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
 
@@
 various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
-on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
+on {\em user-to-local} (u2l) intrusions (intrusions that provide unauthorized access to local superuser (root) privileges),
+and {\em remote-to-local} (r2l) intrusions (intrusions that provide unauthorized access from a remote machine).
 
@@
 
+\subsection{Existing IDSs}
+
+inspiration
+- novelty detection problem with intrusion detection
+- somayaji use sequences of system calls
+- system calls and BSM data can be used to build provenance graphs; we believe that
+provenance graphs provide more information than somayaji and can help detect
+intrusions with better results
+
+use provenance to automatically detect intrusions
+- existing machine learning techniques on KDD data set: use features of an intrusion (number of data)
+- but not actually using graph structure
+- parzen window--allow us to build a profile of normal with lack of abnormal examples
+
+\subsection{Provenance Graph Features}
+
+how we choose features to look at
+- basic statistics for manual detection
+- graph centrality and clustering can help us identify the different kinds of tasks running
+- use these metrics for kernel estimation and
+
+\subsection{Types of Intrusions}
+
+we limit our work to these intrusions
+- explain various kinds
+- why just user 2 local
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Selecting Metrics}
+
+talk about PASS and SPADE here.
+
+put that table here
 
@@
 
-\section{Future Work}
+\section{Conclusion}
+
+\section{Limitations \& Future Work}
 

### 2013-12-10 writeup/paper-final.tex
wrote parts of related work

@@
 %\usepackage{mathabx}
+
+% my macros
+\newcommand{\paren}[1]{\left({#1}\right)}
+\newcommand{\bracket}[1]{\left[{#1}\right]}
+\newcommand{\curly}[1]{\left\{{#1}\right\}}
+\newcommand{\vecb}[1]{\mathbf{#1}}
+\newcommand{\matb}[1]{\mathbf{#1}}
+\newcommand{\V}[1]{\mathbf{#1}}
+\newcommand{\m}[1]{\mathbf{#1}}
+\newcommand{\inhomog}[1]{\widetilde{#1}}
+\newcommand{\transpose}[1]{{#1}^\top}
 
@@
 unusual digital artifacts, we hypothesize that
-intrusions also intrusions leave behind abnormalities
+intrusions also leave behind abnormalities
 in the {\em provenance} of these digital artifacts, the lineage of how these digital artifacts were created.
@@
 system has
-determined an intrusion has occurred, the system follow the causality graph backwards determine the original source of the attack, and 
-then it follow the causality graph forwards to determine all the objects tainted by the attack. However, the authors 
+determined an intrusion has occurred, the system follows the causality graph backwards determine the original source of the attack, and 
+then it follows the causality graph forwards to determine all the objects tainted by the attack. However, the authors 
 do not present a way of using provenance data to automatically detect intrusions.
@@
 \item Discuss which provenance graph statistics we chose to use, how we chose them, and why we chose them.
-\item Analyze the Parzen-Window technique on a real intrusion detection data set and evaluate its performance.
+\item Analyze the Parzen-Window technique with various provenance graph statistics on a real intrusion detection data set and evaluate its performance.
 \item Discuss the limitations of the approach and present future work to transform the technique into an online intrusion detection system.
 \end{enumerate}
+
+POSSIBLY INCLUDE THIS
+* the papers suck at finding u2r
+* because they don't look at provenance graph structure
+* we can do better?
 
@@
 \subsection{Existing IDSs}
+
+Many existing intrusion detection systems have made use of the fact that intrusion detection is an example of a
+{\em novelty detection} problem where access to novel or abnormal patterns is difficult to obtain. Somayaji and Forrest
+attempt to solve this problem by analyzing sequences of system calls (CITE WORK). By analyzing the sequences
+of system calls a process makes over time, their system builds a profile of ``normal" process behavior. When
+the process in the future makes enough patterns of unrecognized sequences of system calls, the system flags
+this process as behaving abnormally and attempts to stop the process, either through exponentially slowing down
+system calls or aborting system calls entirely. Notably, their system is trained only by analyzing ``normal" data
+without knowing explicitly what is considered ``abnormal."
+
+\subsection{Parzen-Windows}
+
+??? person attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
+density estimations of normal behavior (CITE). A density estimation is considered {\em nonparametric} if the estimation
+makes no assumptions about the forms of the PDFs, except that PDFs are smooth. By using a nonparametric
+density estimation, one can build a probability density estimate of normal behavior by only having examples of normal
+behavior and no examples of abnormal behavior.
+
+A {\em Parzen-window} estimate of a probability density function $p(\m(x))$
+based on $n$ examples in a dataset $D$ drawn from model $\mathcal{M}$ is given by:
+$$p(\m{x}) = \frac{1}{n} \sum_{i=1}^n \delta_n (\m{x} - \m{x_i})$$
+where $\delta_n(\cdot)$ is a kernel function. They chose to use radially-symmetric Gaussian kernel functions because
+Gaussian functions are smooth and therefore $p(\m{x})$ will be smooth, and because a radially-symmetric Gaussian function
+can be specified by a single parameter, the variance of the kernel. Using a common variance $\sigma^2$ for all
+the Gaussian kernels, we can rewrite $p(\m{x})$ as:
+$$p(\m{x}) = \frac{1}{n(2\pi)^{d/2} \sigma^d} \sum_{i=1}^n  \exp \left\{  - \frac{|| \m{x} - \m{x_i}||^2}{2 \sigma^2}  \right\}$$
+where $d$ is the dimensionality of the feature space $\m{x}$.
+
+Let $\omega_1$ denote the state of being normal and $\omega_0$ the state of being abnormal.
+To test if an example $\m{x} \in \omega_1$, they reframe the problem using hypothesis testing. Let $\m{y}$ be
+an arbitrary example from $D$ drawn from $\mathcal{M}$, and let 
+$$L(\m{y}) = \log p(\m{y})$$
+be the log-likelihood of $\m{y}$ with respect to $\mathcal{M}$. Then we test the hypothesis that
+$L(\m{x})$ is drawn from the distribution of the log-likelihood of the random examples in $D$ with:
+\begin{eqnarray*}
+P(L(\m{y}) \leq L(\m{x})) 
+&=&  \frac{\#\m{y} \mbox { with } L(\m{y}) \leq L(\m{x})}{n} \\
+&>& \psi
+\end{eqnarray*}
+for some threshold $0 \leq \psi \leq 1$, called the {\em false detection rate}. 
+Thus, $\m{x} \in \omega_1$ is behaving normally if and only if $P(L(\m{y}) \leq L(\m{x})) > \psi$.
+
+In this paper, we use this same model to build profiles of ``normal" behavior for each process. We use
+one-dimensional feature vectors $x$, and we use various provenance graph statistics and graph
+centrality metrics as our features. For the common variance of our Gaussian kernels, we chose
+the variance to be the variance of the data set $D$ for each unique process name. 
+We vary the features we used and values of $\psi$ and evaluate its performance on an intrusion detection data set.
+
+%%% IGNORE BELOW
 
@@
 - parzen window--allow us to build a profile of normal with lack of abnormal examples
+  explain the math of this
+
+  we limit our work to these intrusions
+- explain various kinds
+- why just user 2 local
 
@@
 - use these metrics for kernel estimation and
-
-\subsection{Types of Intrusions}
-
-we limit our work to these intrusions
-- explain various kinds
-- why just user 2 local
 

### 2013-12-10 writeup/paper-final.tex
finish first section of related work, need to do features and types of intrusions

@@
 
-\subsection{Existing IDSs}
+\subsection{Existing IDSs using Provenance}
 
-Many existing intrusion detection systems have made use of the fact that intrusion detection is an example of a
-{\em novelty detection} problem where access to novel or abnormal patterns is difficult to obtain. Somayaji and Forrest
-attempt to solve this problem by analyzing sequences of system calls (CITE WORK). By analyzing the sequences
+Many existing intrusion detection systems use provenance in some limited capacity; however, none of them
+have used provenance for a fully automated detection system. 
+Somayaji and Forrest's work in analyzing sequences of system calls for intrusion detection (CITE WORK) can
+be seen as one of the earliest works in building an intrusion detection system using provenance: sequences of
+system calls can be seen as a limited form of provenance, as many systems build provenance graphs from
+system call traces (CITE SPADE AND BURRITO). By analyzing the sequences
 of system calls a process makes over time, their system builds a profile of ``normal" process behavior. When
@@
 system calls or aborting system calls entirely. Notably, their system is trained only by analyzing ``normal" data
-without knowing explicitly what is considered ``abnormal."
+without knowing explicitly what is considered ``abnormal." Because their system achieves high
+accuracy in determining intrusions with only sequences of system calls, we believe that having real provenance
+data in a graph structure--which in theory would be a superset of system call data--would allow us
+to achieve comparable, if not better intrusion detection accuracy.
+
+More recent work make use of the structure of provenance as a directed graph for a semi-automatic form
+of intrusion detection, and to determine the scope of an intrusion. ??? developed Backtracker (CITE) a modified
+Linux kernel that tracks dependencies between operating system objects (files, processes, file names) similar
+to provenance graphs. Backtracker builds a backwards causal graph to determine the entry point of an intrusion,
+and it builds a forward causal graph to determine possibly tainted files, processes, or hosts in a distributed
+system. However, Backtracker does not provide a way of automatically determining if an intrusion occurred
+using provenance data; the provenance graph structure is only utilized after-the-fact.
+
+??? make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
+in manual intrusion-detection. Given a data set of many provenance graphs, they calculated basic statistics
+on the graphs (e.g., total nodes, total edges, max incoming edges on a node, max outgoing edges on a node,
+node/edge ratio, average total edges per node, etc.). Their approach allows a human to more easily
+notice anomalies in provenance graphs, but they do not provide a way to automate this detection.
+We borrow their ideas of using simple provenance graph statistics as features to detect intrusions.
+
+These works demonstrate that provenance data is indeed useful in collecting information about intrusions, and
+Somayaji and ???'s work suggests that it might be possible to build a fully automated intrusion detection
+system by making use of the graph structure of provenance.
 
@@
 
+To address the difficulty of obtaining abnormal patterns for novelty detection on intrusion detection, we borrow
+the Parzen-Window model approach proposed by ???.
 ??? person attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
@@
 \end{eqnarray*}
-for some threshold $0 \leq \psi \leq 1$, called the {\em false detection rate}. 
+for some threshold $0 < \psi < 1$, called the {\em false detection rate}. 
 Thus, $\m{x} \in \omega_1$ is behaving normally if and only if $P(L(\m{y}) \leq L(\m{x})) > \psi$.
@@
 one-dimensional feature vectors $x$, and we use various provenance graph statistics and graph
-centrality metrics as our features. For the common variance of our Gaussian kernels, we chose
+centrality metrics as our features. We describe the various statistics and graph centrality metrics we explored
+in the following section. For the common variance of our Gaussian kernels, we chose
 the variance to be the variance of the data set $D$ for each unique process name. 
 We vary the features we used and values of $\psi$ and evaluate its performance on an intrusion detection data set.
-
-%%% IGNORE BELOW
-
-inspiration
-- novelty detection problem with intrusion detection
-- somayaji use sequences of system calls
-- system calls and BSM data can be used to build provenance graphs; we believe that
-provenance graphs provide more information than somayaji and can help detect
-intrusions with better results
-
-use provenance to automatically detect intrusions
-- existing machine learning techniques on KDD data set: use features of an intrusion (number of data)
-- but not actually using graph structure
-- parzen window--allow us to build a profile of normal with lack of abnormal examples
-  explain the math of this
-
-  we limit our work to these intrusions
-- explain various kinds
-- why just user 2 local
 
@@
 - use these metrics for kernel estimation and 
+- cite machine learning
+
+\subsection{Intrusion Types}
+
+use provenance to automatically detect intrusions
+- existing machine learning techniques on KDD data set: use features of an intrusion (number of data)
+- but not actually using graph structure
+
+  we limit our work to these intrusions
+- explain various kinds
+- why just user 2 local
 

### 2013-12-10 writeup/paper-final.tex
change margins and add acks
related work finished

@@
 
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=0.7in]{geometry}
 
@@
 \item Discuss which provenance graph statistics we chose to use, how we chose them, and why we chose them.
-\item Analyze the Parzen-Window technique with various provenance graph statistics on a real intrusion detection data set and evaluate its performance.
+\item Analyze the Parzen-Window technique with various provenance graph statistics on a real intrusion detection data set and evaluate its accuracy.
 \item Discuss the limitations of the approach and present future work to transform the technique into an online intrusion detection system.
@@
 
-A {\em Parzen-window} estimate of a probability density function $p(\m(x))$
+A {\em Parzen-window} estimate of a probability density function $p(\m{x})$
 based on $n$ examples in a dataset $D$ drawn from model $\mathcal{M}$ is given by:
@@
 
-how we choose features to look at
-- basic statistics for manual detection
-- graph centrality and clustering can help us identify the different kinds of tasks running
-- use these metrics for kernel estimation and 
-- cite machine learning
+We were inspired by the work proposed by ??? who use simple provenance graph statistics to aide
+in manual intrusion detection by humans. We extend their idea to make use of the typed node and typed edge
+nature of provenance graphs. (CITE) For example, instead of simply counting number of edges from a node, we
+counted number of edges of type {\em input} directed towards {\em file} type nodes, or the number of {\em fork}
+edges from {\em process} type nodes. Below is the list of simple provenance graph statistics we used:
+\begin{enumerate}
+\item number of input and output files
+\item number of input and output processes
+\item number of input and output pipes
+\item number of versions stemming from this node
+\end{enumerate}
+
+Furthermore, we hypothesize that intrusions manifest as anomalies in the lineage of digital artifacts, and
+one way of observing this is through {\em graph centrality metrics}. A {\em graph centrality metric} is a function
+of a node and its containing graph, and intuitively represents how {\em central} or {\em important} the node
+is in the graph. In terms of provenance graphs, processes of the same program typically behave similarly to one another
+and this should result in similar number of inputs, outputs, number of ancestors, and number of descendants for normal
+process nodes of the same program.
+Intrusions typically change the number of expected inputs and outputs (e.g., a process forking a shell process
+when it normally does not), and we hypothesize that this can manifest as unusual changes in centrality
+for the offending node in the provenance graph. ??? note the following 3 observations unique to provenance graphs (CITE):
+\begin{enumerate}
+\item The ubiquity (importance) of a node is a function of a node's descendants, not its ancestors.
+(Note: Because edges in provenance graphs are in the direction of data dependence instead of data flow, ``descendants" in this case means ancestors in the provenance graph).
+\item There is no agreed-upon granularity in which provenance should be captured.
+\item Provenance has a temporal component.
+\end{enumerate}
+
+Using these observations, they propose several centrality metrics on provenance graphs:
+\begin{itemize}
+\item \textbf{In-degree Centrality}.
+Because a node's importance is based on its output descendants, then the number of times a node was used as input to another node indicates a simple measure of importance.
+\item \textbf{Age}. They note that large "jumps" in timestamps or age might indicate breaks between tasks, but this interpretation is complicated when scripts execute tasks in short succession.
+\item \textbf{Ancestor Centrality}. This measures the total number of descendants of a node, normalized by the total number of nodes in a graph.
+\item \textbf{Opsahl's Closeness Centrality}. This accounts for the number of edges between a node and its descendants. To compute this, we calculate
+$$CC'(v) = \sum_{x \in V -  \{v\}} (d'_{xv})^{-1}$$
+where $V$ is the set of nodes in the graph, $d'_{xv}$ is the distance from node $x$ to $v$ by only following outgoing edges from $x$.
+Unreachable nodes from $x$ will have $d'_{xv} = \infty$, and so $(d'_{xv})^{-1} = 0$.
+\item \textbf{Provenance Eigenvector Centrality}. The centrality of the nodes is given by the left dominant eigenvector of the matrix:
+$$M_{ij} =
+\begin{cases}
+1 & \text{if there is an edge } i \to j \\
+1/|V| & \text{if there is no out edge from } i \\
+0 & \text{otherwise}
+\end{cases}
+$$
+\end{itemize}
+We implemented some of these graph centrality metrics and used Parzen-Windows on centrality measurements to compute density estimates for each process.
 
@@
 
-use provenance to automatically detect intrusions
-- existing machine learning techniques on KDD data set: use features of an intrusion (number of data)
-- but not actually using graph structure
-
-  we limit our work to these intrusions
-- explain various kinds
-- why just user 2 local
+The DARPA Intrusion Detection Data Sets (CITE) describe various kinds of intrusions in a network, including:
+\begin{enumerate}
+\item {\bf Denial of Service (dos)}. Intrusion that results in denial of service, e.g., overwhelming a server.
+\item {\bf Remote to Local (r2l)}. Unauthorized access from a machine, e.g., guessing passwords.
+\item {\bf User to Root (u2l)}. Unauthorized access to local superuser (root) privileges, e.g., various buffer overflow attacks.
+\item {\bf Probing (probe)}. Surveillance and scanning, e.g. port scanning.
+\end{enumerate}
+Because we only collect provenance graphs local to the machine, we chose to limit our intrusion detection system only to u2l intrusions within this paper, as the other types of attacks involve exploiting the network and would not appear in provenance data.
 
@@
 \section{Limitations \& Future Work}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ACKNOWLEDGEMENTS
+%
+
+\section{Acknowledgements}
 

### 2013-12-10 writeup/paper-final.tex
amost done with Design & Imp

@@
 counted number of edges of type {\em input} directed towards {\em file} type nodes, or the number of {\em fork}
-edges from {\em process} type nodes. Below is the list of simple provenance graph statistics we used:
-\begin{enumerate}
-\item number of input and output files
-\item number of input and output processes
-\item number of input and output pipes
-\item number of versions stemming from this node
-\end{enumerate}
+edges from {\em process} type nodes.
 
@@
 
-\subsection{Selecting Metrics}
+\subsection{Preliminary Provenance Exploration}
 
-talk about PASS and SPADE here.
+To determine which provenance statistics seem useful, we first created simple simulated intrusions
+and ran them in a Provenance Aware Storage System (PASS) (CITE), a file system with
+a mounted volume that automatically collects provenance of all files within that volume.
+To make data collection easier, we used PASSv2, which provides an Ubuntu virtual machine with
+the PASS volume already installed.
 
-put that table here
+For each of our intrusions, we ran it normally multiple times and collected the provenance. We then ran the
+intrusion with the exploit multiple times and collected the new provenance.
+
+We used the following procedure to implement our Parzen-Window density estimates and
+intrusion/normal decision.
+\begin{enumerate}
+\item For each statistic that we implemented, we computed the counts for that statistic for all the
+nodes in the provenance graph containing the normally-behaving workload.
+\item We aggregated the counts by the name of the process (the path to the executable). For a given process $P$, denote the
+set of counts for that process name as $D_P$.
+\item For each $y \in D_P$, calculate the log-likelihood $L_P(y)$.
+\item For a test node with process name $P$ in the exploited-workload provenance graph, we calculate the value $x$ of the statistic in its graph, and then compute $L_P(x)$. 
+If the ratio of the number of nodes $y$ with $L_P(y) \leq L_P(x)$,
+to the total number of nodes in $D_P$ is greater than the false detection rate $\psi$, we label the node as normal. Otherwise we label the node as
+an intrusion.
+\end{enumerate}
+
+In the exploited workloads, we used Orbiter (CITE) to find the abnormally behaving node by hand, and we used these nodes as our test nodes.
+
+\subsection{Simulated Intrusions}
+
+Below is a list of the intrusions we simulated.
+
+\subsubsection{hello}
+We found that many of the popular exploits involved buffer overflows and arbitrary code execution and possibly obtain root access.
+We created a simple \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
+
+We ran \texttt{hello} normally 40 times on various inputs and collected the provenance times.
+We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
+
+\subsubsection{mcrypt}
+The mcrypt encryption package had a long standing buffer overflow, which would allow a malicious user to run their own code when mcrypt tried to decrypt the crafted exploit file.
+Due to limitations in our Virtual Machine set up, we were unable to exploit the vulnerable version of mcrypt, much like the authors of Backtracker \cite{backtracker} found.
+To emulate this exploit, we designed a Python program, which would decrypt (via ROT13) an input file.
+When sent a particular exploit file which contained malicious lines, our program would stop decrypting and start running the code specified in the file.
+
+We ran our program on over 200 good inputs, and stored the provenance data.
+We then ran our program on two bad inputs, first to call \texttt{ls} as a Proof of Concept, and again to call \texttt{/bin/sh} to demonstrate that more severe attacks can be performed.
+We again exported the provenance generated from this exploited workload.
+
+\subsubsection{gcc}
+Another type of exploit is more reminiscent of a system administration failure.
+By unexpectedly aliasing an important command, a user can unknowingly grant root access to a malicious program.
+To emulate this sort of attack, we replaced \texttt{/usr/bin/gcc} with our own version of gcc, which, in addition to compiling the user's code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
+
+For our system administration attack, we compiled the PASS toolset using GCC, and exported the provenance from these runs.
+We then replaced GCC with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
+
+\subsection{Selecting Provenance Statistics}
+
+Below is the list of simple provenance graph statistics we implemented:
+\begin{enumerate}
+\item number of input and output files
+\item number of input and output processes
+\item number of input and output pipes
+\item number of versions stemming from this node
+\end{enumerate}
+
+We also implemented the indegree, Opsahl, and ancestor centrality metrics.
+We chose not to implement the Provenance Eigenvector Centrality because computing
+the eigenvectors of the large matrix took too much computation time on our machines.
+
+USE FIRST PAPER's RESULTS HERE to describe which metrics we should use.
 
@@
 \subsection{Experimental Setup}
+
+talk about SPADE
 
@@
 \section{Acknowledgements}
+We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project and for
+their help in working with PASS and analyzing provenance graphs. We also want to thank the
+CS261 Program Committee for their feedback on the initial draft of this paper.
 

### 2013-12-10 writeup/paper-final.tex
finish design and imp, wrote future work

@@
 
-\usepackage[margin=0.7in]{geometry}
+\usepackage[margin=1in]{geometry}
 
@@
 
-Below is a list of the intrusions we simulated.
-
 \subsubsection{hello}
@@
 the eigenvectors of the large matrix took too much computation time on our machines.
+We also chose not to implement age because age would make our data too dependent
+on how quickly we executed our workloads.
 
-USE FIRST PAPER's RESULTS HERE to describe which metrics we should use.
+\begin{table*}[ht]
+\label{results}
+{\small
+  \begin{center}
+  \begin{tabular}{| l | l | l | l |}
+    \hline
+    statistic & hello & mcrypt & gcc \\ \hline
+     \# forked processes & 0.0 | 1 | 0.108 & 0.0 | 1 | 0.107 & 0.9 | 0 | 0.408 \\ \hline
+     \# input files & 0.5 | 0 | 0.453 & 0.5 | 0 | 0.453 & 0.5 | 1 | 0.107 \\ \hline
+     \# output files & 0.5 | 1 | 0.453 & 0.5 | 1 | 0.453 & 0.0 | 3 | 0.000 \\ \hline
+    Opsahl centrality & 1.791 | 2.83 | 0.252 & 1.250 | 4.867 | 0.000 & 2.793 | 7.415 | 0.024 \\ \hline
+    Ancestor centrality & 0.0009 | 0.0011 | 0.799 & 0.0002 | 0.0014 | 0.798 & 0.0007 | 0.0002 | 0.798 \\
+
+    \hline
+  \end{tabular}
+  \end{center}
+}
+\hfill{}
+\caption{average over good | test node value | density estimate of test node being ``normal"
+}
+\label{tb:tablename}
+\end{table*}
+
+In Table 1, we show a summary of the statistics that had any statistical difference between normal and abnormally behaving nodes.
+The buffer overflows showed the most abnormal behavior in the number of outgoing processes and the centrality measures.
+This makes sense, as arbitrary code execution will generate additional processes.
+Of the centrality metrics we implemented, the Opsahl Centrality metric did the best in distinguishing an exploited node from normally behaving nodes.
+Because the Opsahl Centrality metric accounts for the number of edges between a node and all of its descendants, this centrality metric is very sensitive to the addition of nodes.
+Exploits that have additional outputs (e.g.
+new processes and output files through a remote shell), thus generate a higher Opsahl centrality metric for the offending node.
+The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that the exploit will be caught by our model.
+
+If the exploit does directly not generate additional files, however, statistics like "number of output files" will fail to be statistically significant.
+Indeed, the two programs (hello and mcrypt) that did not generate additional files were not flagged by that measure.
+On the other hand, the node representing the exploited version of \texttt{gcc} had abnormal density estimates, as it generated many more files than a normal \texttt{gcc} process.
+
+Because of this exploratory analysis, we chose to implement Opsahl centralty and Ancestory centrality and analyze its accuracy on a real intrusion detection data set.
 
@@
 
-talk about SPADE
+talk about SPADE and building graphs from BSM
 
@@
 
+\section{Future Work}
+
+\subsection{N-distance Statistics}
+Some processes fork helper processes that do the real work (such as \texttt{gcc}).
+We found that \texttt{gcc} forks processes for the linker/assembler that handles generating output, so \texttt{gcc} does not directly have output files.
+Similarly, when running a python script, the code file itself is not directly executed, it is read in by a python process, which then goes off and collects input, generates output, and forks additional processes.
+Thus, when running a python file, the file itself only has one node in the provenance graph, with many python process nodes accessing it.
+In order to account for these kinds of relationships, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
+By utilizing information about neighboring nodes, we hope to detect intrusions that involve more complicated relationships
+between nodes.
+\subsection{ENV and ARGV}
+Currently we ignore the environment and argument information stored in the provenance graph.
+By generating histograms based on this information, we could catch subtler intrusions that rely on modified environment information.
+Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
+\subsection{Online Intrusion Detection}
+Currently, this system runs after-the-fact, but truly useful intrusion detection systems need to run while the system is online.
+To do this, we would need to modify our statistics to use centrality metrics that are {\em progressive}, i.e., they can be recomputed online without needing to recompute the metric on the entire graph.
+
+Given progressive forms of these statistics, we can modify PASS to compute Parzen-Window density estimates as it generates provenance, or we can audit system calls like SPADE (CITE) to build provenance graphs in user space and apply our Parzen-Window technique.
+\subsection{Local vs. Global}
+In our experiments, we looked at provenance graphs in their entirety.
+However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph.
+To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
+Another approach would be to discard provenance data in computing Parzen-Window density estimates after the data have reached a certain age.
+
 \section{Conclusion}
-
-\section{Limitations \& Future Work}
 
@@
 We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project and for
-their help in working with PASS and analyzing provenance graphs. We also want to thank the
+their help in working with PASS and analyzing provenance graphs. We also want to thank the members of the
 CS261 Program Committee for their feedback on the initial draft of this paper.

### 2013-12-10 writeup/paper-final.tex
wrote experimental methodology

@@
 
-Below is the list of simple provenance graph statistics we implemented:
-\begin{enumerate}
-\item number of input and output files
-\item number of input and output processes
-\item number of input and output pipes
-\item number of versions stemming from this node
-\end{enumerate}
-
-We also implemented the indegree, Opsahl, and ancestor centrality metrics.
-We chose not to implement the Provenance Eigenvector Centrality because computing
-the eigenvectors of the large matrix took too much computation time on our machines.
-We also chose not to implement age because age would make our data too dependent
-on how quickly we executed our workloads.
-
 \begin{table*}[ht]
@@
 
+Below is the list of simple provenance graph statistics we implemented:
+\begin{enumerate}
+\item number of input and output files
+\item number of input and output processes
+\item number of input and output pipes
+\item number of versions stemming from this node
+\end{enumerate}
+We also implemented the indegree, Opsahl, and ancestor centrality metrics.
+We chose not to implement the Provenance Eigenvector Centrality because computing
+the eigenvectors of the large matrix took too much computation time on our machines.
+We also chose not to implement age because age would make our data too dependent
+on how quickly we executed our workloads.
+
 In Table 1, we show a summary of the statistics that had any statistical difference between normal and abnormally behaving nodes.
@@
 
-Because of this exploratory analysis, we chose to implement Opsahl centralty and Ancestory centrality and analyze its accuracy on a real intrusion detection data set.
+Because of this exploratory analysis, we chose to implement Opsahl and ancestor centrality and analyze its accuracy on a real intrusion detection data set.
 
@@
 
-talk about SPADE and building graphs from BSM
+We evaluated our approach by analyzing its accuracy on the 1998 DARPA Intrusion Detection Challenge data set (CITE).
+The data set consists of users of several weeks of network-based attacks in the midst of normal
+background data. The data includes BSM (Basic Security Model) traces (CITE), which provide an
+audit of all the system calls made within a certain time period. The data also includes a list
+of specific moments in time $T$ and whether an intrusion occurred and if so, what kind of intrusion.
+
+For our experiments, we limited the scope of our intrusion detection system to only u2r attacks, as
+the other kinds of attacks were network-based and would not appear in the provenance data. 
+We identify each accuracy measurement with a tuple:
+$$(\mbox{centrality metric}, \psi, W)$$
+where the centrality metric is one of Opsahl, indegree, or ancestor, $\psi$ is our false detection rate for our Parzen-Window
+estimates, and $W$ is the time window that we use to form time intervals to look for anomalous nodes. We used
+the following procedure to evaluate the accuracy of our system: 
+\begin{enumerate}
+\item Use BSM traces to build provenance graphs, in a similar fashion to SPADE (CITE). 
+\item Compute the centralities of the nodes in the graph using the given centrality metric, and create a mapping from each process
+name $P$ to a list of log-likelihood values $L(y)$, where $y$ is the computer centrality measurement for each node in the graph having process name $P$.
+\item For a specific moment in time $T$, find all edges in the provenance graph we built with timestamps in the interval $[T - W, T + W]$. 
+Use the nodes attached to these edges as our {\em candidate anomaly set}. We used the specific times listed in data sets for our values of $T$.
+\item For all nodes in the candidate anomaly set, find the process name $P$ of the node, and use our Parzen-Window approach on process
+name $P$ to determine if the node is anomalous. We evaluate whether a node is normal or anomalous using the Parzen-Windows built from the graph
+containing the node.
+\item For times $T$ in the data set listed as having an intrusion, our intrusion detection system considers that time $T$ as having an intrusion
+if there are {\em sufficient} number of nodes in the $[T - W, T + W]$ window considered anomalous. A candidate anomaly set has a 
+{\em sufficient} number of anomalous nodes if TODO...
+\end{enumerate}
+
+For each $(\mbox{centrality metric}, \psi, W)$, we compute:
+\begin{itemize}
+\item {\em False Positive Rate (FPR).} Ratio of number of time intervals that do not actually contain an intrusion
+but were misidentified as having an intrusion, to the total number of time intervals.
+\item {\em True Positive Rate (TPR).} Ratio of the number of time intervals that actually do contain an intrusion
+and were correctly identified as having an intrusion, to the total number of time intervals.
+\end{itemize}
+
+We evaluated the accuracy of our system on various values of $\psi$ and $W$.
 
@@
 
+TODO
+
+Need these graphs:
+[1]:
+- fix psi
+- x-axis is W (probably log-scale, as W = 10ms, 1000ms, 100000ms, 10 ** N ms)
+- y-axis is [0,1]
+- overlay true positive rate for each of the centrality metrics in different color
+
+[2]: repeat [1] but plot false positive rate instead
+
+[3-4] repeat [1-2] with a different value of psi
+
+[5-6] repeat [1-2] with another value of psi
+
 \subsection{Discussion}
+
+TODO
 
@@
 \section{Conclusion}
+TODO
 

### 2013-12-10 writeup/paper-final.tex
cite references

@@
 We present an intrusion detection approach to find these anomalies by analyzing centrality metrics
-on provenance graphs. We use a Parzen-Window approach (TODO CITE) on various provenance graph centrality metrics (TODO CITE)
+on provenance graphs. We use a Parzen-Window approach \cite{parzen} on various provenance graph centrality metrics \cite{clustering}
 to determine probability density estimates of normal behavior, and we use these density estimates to 
-determine if an intrusion occurred. We used this approach to analyze {\em user-to-remote} (u2r) intrusions and 
-{\em remote-to-local} (r2l) intrusions (TODO: include r2l?) from the 1998 DARPA Intrusion Detection data sets (TODO CITE) and 
+determine if an intrusion occurred. We used this approach to analyze {\em user-to-remote} (u2r) intrusions 
+from the 1998 DARPA Intrusion Detection data sets \cite{darpa} and 
 achieved up to *TODO true positive rate for intrusions* accuracy in detecting
@@
 basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
-easier manual intrusion detection by a human (CITE WORK). Their work, however, does not present a way of using provenance
+easier manual intrusion detection by a human \cite{provstat}. Their work, however, does not present a way of using provenance
 graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
-intrusion. For example, Backtracker (CITE WORK) uses provenance graphs to build causality graphs of intrusions: once the
+intrusion. For example, Backtracker \cite{backtracker} uses provenance graphs to build causality graphs of intrusions: once the
 system has
@@
 
->>> INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
+>>> TODO INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
 
@@
 an unknown test pattern is produced by an underlying distribution corresponding to a training set
-of normal patterns (CITE WORK). However, ???authors note that in the case of intrusion detection, novel
+of normal patterns \cite{parzen}. However, ???authors note that in the case of intrusion detection, novel
 or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach
 for non-parametric density estimation. Using this technique, they obtain a high degree of success
-in identifying various intrusion types (CITE WORK). Because obtaining abnormal patterns is difficult, 
+in identifying various intrusion types \cite{parzen}. Because obtaining abnormal patterns is difficult, 
 we borrow their Parzen-Window approach to build models of normal behavior using 
@@
 have used provenance for a fully automated detection system. 
-Somayaji and Forrest's work in analyzing sequences of system calls for intrusion detection (CITE WORK) can
+Somayaji and Forrest's work in analyzing sequences of system calls for intrusion detection \cite{somayaji, somayaji-recent} can
 be seen as one of the earliest works in building an intrusion detection system using provenance: sequences of
 system calls can be seen as a limited form of provenance, as many systems build provenance graphs from
-system call traces (CITE SPADE AND BURRITO). By analyzing the sequences
+system call traces \cite{spade}. By analyzing the sequences
 of system calls a process makes over time, their system builds a profile of ``normal" process behavior. When
@@
 More recent work make use of the structure of provenance as a directed graph for a semi-automatic form
-of intrusion detection, and to determine the scope of an intrusion. ??? developed Backtracker (CITE) a modified
+of intrusion detection, and to determine the scope of an intrusion. ??? developed Backtracker \cite{backtracker}
+ a modified
 Linux kernel that tracks dependencies between operating system objects (files, processes, file names) similar
@@
 ??? make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
-in manual intrusion-detection. Given a data set of many provenance graphs, they calculated basic statistics
+in manual intrusion-detection \cite{provstat}. Given a data set of many provenance graphs, they calculated basic statistics
 on the graphs (e.g., total nodes, total edges, max incoming edges on a node, max outgoing edges on a node,
@@
 ??? person attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
-density estimations of normal behavior (CITE). A density estimation is considered {\em nonparametric} if the estimation
+density estimations of normal behavior \cite{parzen}. A density estimation is considered {\em nonparametric} if the estimation
 makes no assumptions about the forms of the PDFs, except that PDFs are smooth. By using a nonparametric
@@
 in manual intrusion detection by humans. We extend their idea to make use of the typed node and typed edge
-nature of provenance graphs. (CITE) For example, instead of simply counting number of edges from a node, we
+nature of provenance graphs \cite{provstat}. For example, instead of simply counting number of edges from a node, we
 counted number of edges of type {\em input} directed towards {\em file} type nodes, or the number of {\em fork}
@@
 when it normally does not), and we hypothesize that this can manifest as unusual changes in centrality
-for the offending node in the provenance graph. ??? note the following 3 observations unique to provenance graphs (CITE):
+for the offending node in the provenance graph. ??? note the following 3 observations unique to provenance graphs \cite{clustering}:
 \begin{enumerate}
@@
 
-The DARPA Intrusion Detection Data Sets (CITE) describe various kinds of intrusions in a network, including:
+The DARPA Intrusion Detection Data Sets \cite{darpa} describe various kinds of intrusions in a network, including:
 \begin{enumerate}
@@
 To determine which provenance statistics seem useful, we first created simple simulated intrusions
-and ran them in a Provenance Aware Storage System (PASS) (CITE), a file system with
+and ran them in a Provenance Aware Storage System (PASS) \cite{pass}, a file system with
 a mounted volume that automatically collects provenance of all files within that volume.
-To make data collection easier, we used PASSv2, which provides an Ubuntu virtual machine with
+To make data collection easier, we used PASSv2 \cite{passv2}, which provides an Ubuntu virtual machine with
 the PASS volume already installed.
@@
 
-In the exploited workloads, we used Orbiter (CITE) to find the abnormally behaving node by hand, and we used these nodes as our test nodes.
+In the exploited workloads, we used Orbiter \cite{orbiter} to find the abnormally behaving node by hand, and we used these nodes as our test nodes.
 
@@
 \subsubsection{hello}
-We found that many of the popular exploits involved buffer overflows and arbitrary code execution and possibly obtain root access.
+We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution and possibly obtain root access.
 We created a simple \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
@@
 
-For our system administration attack, we compiled the PASS toolset using GCC, and exported the provenance from these runs.
-We then replaced GCC with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
+For our system administration attack, we compiled the PASS toolset using \texttt{gcc}, and exported the provenance from these runs.
+We then replaced \texttt{gcc} with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
 
@@
 
-We evaluated our approach by analyzing its accuracy on the 1998 DARPA Intrusion Detection Challenge data set (CITE).
+We evaluated our approach by analyzing its accuracy on the 1998 DARPA Intrusion Detection Challenge data set \cite{darpa}.
 The data set consists of users of several weeks of network-based attacks in the midst of normal
-background data. The data includes BSM (Basic Security Model) traces (CITE), which provide an
+background data. The data includes BSM (Basic Security Model) traces \cite{bsm}, which provide an
 audit of all the system calls made within a certain time period. The data also includes a list
@@
 \begin{enumerate}
-\item Use BSM traces to build provenance graphs, in a similar fashion to SPADE (CITE). 
+\item Use BSM traces to build provenance graphs, in a similar fashion to SPADE \cite{spade}. 
 \item Compute the centralities of the nodes in the graph using the given centrality metric, and create a mapping from each process
@@
 
-Given progressive forms of these statistics, we can modify PASS to compute Parzen-Window density estimates as it generates provenance, or we can audit system calls like SPADE (CITE) to build provenance graphs in user space and apply our Parzen-Window technique.
+Given progressive forms of these statistics, we can modify PASS to compute Parzen-Window density estimates as it generates provenance, or we can audit system calls like SPADE \cite{spade} to build provenance graphs in user space and apply our Parzen-Window technique.
 \subsection{Local vs. Global}
@@
 
-\bibitem{fuzzy}
-\textsc{Cao, D., Qiu, M., Chen, Z., Hu, F., Zhu, Y., and Wang, B.} Intelligent Fuzzy Anomaly Detection of Malicious Software.
-In {\em Internal Journal of Advanced Intelligence}, vol.
-4, no.
-1, pp 69-86 (December 2012).
+\bibitem{provstat}
+\textsc{Allen, D.
+M., Chapman, A., Seligman, L., and Blaustein, B.} In {\em Proceedings of the International Conference on Collaborative Computing: Networking, Applications and Worksharing} (October 2011).
+
+\bibitem{darpa}
+\textsc{DARPA Intrusion Detection Evaluation Data Set, 1998.} {\tt http://www.ll.mit.edu/mission/\\communications/cyber/CSTcorpora/ideval/\\data/1998data.html}.
+
+\bibitem{spade}
+\textsc{Gehani, A.
+and Tariq, D.} SPADE: Support for Provenance Auditing in Distributed Environments.
+In {\em Proceedings of the 2012 ACM/IFIP/USENIX International Middleware Conference} (December 2012).
 
@@
 \textsc{King, S. T. and Chen, P. M.} Backtracking Intrusions. In {\em SOSP'03 Proceedings of the nineteenth ACM symposium on Operating systems principles} (December 2003).
-
-\bibitem{multihost}
-\textsc{King, S.
-T., Mao Z.
-M., Lucchetti, D.
-G., and Chen, P.
-M.} Enriching intrusion alerts through multi-host causality.
-In {\em Proceedings of the 2005 Network and Distributed System Security Symposium} (February 2005).
-
-\bibitem{fileprefetch}
-\textsc{Lei, H. and Duchamp, D.} An Analytical Approach to File Prefetching. In {\em Proceedings of the USENIX 1997 Annual Technical Conference} (January 1997).
 
@@
 In {\em TaPP'11 Proceedings of the 2nd conference on Theory and practice of provenance} (June 2011).
-
-\bibitem{fileattributes}
-\textsc{Margo, D., and Smogor, R.} Using Provenance to Extract Semantic File Attributes. In {\em TaPP'10 Proceedings of the 2nd conference on Theory and practice of provenance} (February 2010).
 
@@
 
-\bibitem{correlated-anomalies}
-\textsc{Tariq, D., Baig, B., Gehani, A., Mahmood, S., Tahir, R., Aqil, A., and Zaffar, F.} Identifying the provenance of correlated anomalies.
-In {\em SAC'11 Proceedings of the 2011 ACM Symposium on Applied Computing} (March 2011).
+\bibitem{bsm}
+\textsc{SunSHIELD Basic Security Module Guide.} {\tt http://docs.oracle.com/cd/\\E19455-01/806-1789/6jb25l4bc/index.html}.
+
+\bibitem{parzen}
+\textsc{Yeung, D. and Chow, C.} Parzen-window network intrusion detectors. In {\em Proceedings of 2002 International Conference on Pattern Recognition} (August 2002).
 

### 2013-12-10 writeup/paper-final.tex
replace question marks with authors' names

@@
 Existing intrusion detection systems (IDS) use provenance data only to a limited extent.
-??? person analyzes
+Allen et. al. analyze
 basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
@@
 an unknown test pattern is produced by an underlying distribution corresponding to a training set
-of normal patterns \cite{parzen}. However, ???authors note that in the case of intrusion detection, novel
+of normal patterns \cite{parzen}. However, Yeung et. al. note that in the case of intrusion detection, novel
 or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach
@@
 More recent work make use of the structure of provenance as a directed graph for a semi-automatic form
-of intrusion detection, and to determine the scope of an intrusion. ??? developed Backtracker \cite{backtracker}
+of intrusion detection, and to determine the scope of an intrusion. King et. al. developed Backtracker \cite{backtracker}
  a modified
@@
 
-??? make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
+Allen et. al. make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
 in manual intrusion-detection \cite{provstat}. Given a data set of many provenance graphs, they calculated basic statistics
@@
 These works demonstrate that provenance data is indeed useful in collecting information about intrusions, and
-Somayaji and ???'s work suggests that it might be possible to build a fully automated intrusion detection
+Somayaji and King's work suggests that it might be possible to build a fully automated intrusion detection
 system by making use of the graph structure of provenance.
@@
 To address the difficulty of obtaining abnormal patterns for novelty detection on intrusion detection, we borrow
-the Parzen-Window model approach proposed by ???.
-??? person attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
+the Parzen-Window model approach proposed by Yeung et. al.
+They attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
 density estimations of normal behavior \cite{parzen}. A density estimation is considered {\em nonparametric} if the estimation
@@
 
-We were inspired by the work proposed by ??? who use simple provenance graph statistics to aide
+We were inspired by the work proposed by Allen et. al. who use simple provenance graph statistics to aide
 in manual intrusion detection by humans. We extend their idea to make use of the typed node and typed edge
@@
 when it normally does not), and we hypothesize that this can manifest as unusual changes in centrality
-for the offending node in the provenance graph. ??? note the following 3 observations unique to provenance graphs \cite{clustering}:
+for the offending node in the provenance graph. Macko et. al. note the following 3 observations unique to provenance graphs \cite{clustering}:
 \begin{enumerate}
@@
 \textsc{Allen, D.
-M., Chapman, A., Seligman, L., and Blaustein, B.} In {\em Proceedings of the International Conference on Collaborative Computing: Networking, Applications and Worksharing} (October 2011).
+M., Chapman, A., Seligman, L., and Blaustein, B.} Provenance for Collaboration: Detecting Suspicious Behaviors and Assessing Trust in Information.
+In {\em Proceedings of the International Conference on Collaborative Computing: Networking, Applications and Worksharing} (October 2011).
 
@@
 \bibitem{parzen}
-\textsc{Yeung, D. and Chow, C.} Parzen-window network intrusion detectors. In {\em Proceedings of 2002 International Conference on Pattern Recognition} (August 2002).
+\textsc{Yeung, D. and Chow, C.} Parzen-Window Network Intrusion Detectors. In {\em Proceedings of 2002 International Conference on Pattern Recognition} (August 2002).
 

### 2013-12-10 writeup/paper-final.tex
little issues

@@
 \begin{abstract}
-Provenance is data that describes how a digital artifact came to be in its current state. We hypothesize
-that intrusions on a system leave behind anomalies in the lineage of digital artifacts.
+Security rests on being able to identify potentially malicious behavior.
+One way to identify malicious behavior is to compare system behavior to known normal behavior.
+Intrusions on a system leave typically leave behind digital artifacts, and we hypothesize
+that we can detect intrusions on a system by identifying anomalies the artifacts' provenance, the lineage of how
+the artifact came to be in its present state.
 We present an intrusion detection approach to find these anomalies by analyzing centrality metrics
@@
 file nodes have edges directed towards previous versions of the file and to the nodes of the
-processes that modified the file.
+processes that modified the file. See Figure \ref{graph-simple} for an illustration of these provenance
+graph node and edge types.
 
@@
 
->>> TODO INSERT PROVENANCE GRAPH SMALL EXAMPLE HERE SHOWING DATA FLOW
+\begin{figure}
+  \label{graph-simple}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/graph-simple.png}
+    \caption{Example of a provenance graph, showing the types of nodes and edges common in provenance graphs} 
+\end{figure}
 
@@
 
-POSSIBLY INCLUDE THIS
+POSSIBLY INCLUDE THIS TODO
 * the papers suck at finding u2r
@@
 \section{Design and Implementation}
+
+We envision a provenance aware system which constantly builds normal profiles of processes. Simultaneously,
+the system looks for instances of deviation from normal behavior and handles them appropriately
+(e.g., terminating the process, slowing down the process). In our work, we derive normal profiles of process behavior
+from provenance graphs, and we use a Parzen-Window model on various provenance graph statistics
+to determine if anomalous nodes appear in the graph.
 
@@
 
+\subsection{Limitations}
+
+TODO
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Another approach would be to discard provenance data in computing Parzen-Window density estimates after the data have reached a certain age.
+By allowing the model to evolve over time, the system
+can combat an {\em autoimmune disorders} in which attackers purposely invoke programs in weird ways to fool the system into thinking the programs are
+behaving abnormally.
 

### 2013-12-10 writeup/paper-final.tex
edits

@@
 One way to identify malicious behavior is to compare system behavior to known normal behavior.
-Intrusions on a system leave typically leave behind digital artifacts, and we hypothesize
-that we can detect intrusions on a system by identifying anomalies the artifacts' provenance, the lineage of how
+Intrusions on a system typically leave behind digital artifacts, and we hypothesize
+that we can detect intrusions on a system by identifying anomalies in the artifacts' {\em provenance}, or the lineage of how
 the artifact came to be in its present state.
-We present an intrusion detection approach to find these anomalies by analyzing centrality metrics
-on provenance graphs. We use a Parzen-Window approach \cite{parzen} on various provenance graph centrality metrics \cite{clustering}
+We present an intrusion detection approach to find these anomalies by analyzing provenance graphs statistics.
+We use a Parzen-Window approach \cite{parzen} on various provenance graph statistics \cite{clustering}
 to determine probability density estimates of normal behavior, and we use these density estimates to 

### 2013-12-10 writeup/paper-final.tex
finished writing, add more diagrams

@@
 to determine probability density estimates of normal behavior, and we use these density estimates to 
-determine if an intrusion occurred. We used this approach to analyze {\em user-to-remote} (u2r) intrusions 
-from the 1998 DARPA Intrusion Detection data sets \cite{darpa} and 
-achieved up to *TODO true positive rate for intrusions* accuracy in detecting
-intrusions with only *TODO false positive rate for intrusions* accuracy. We also present future work
-to extend our intrusion model to an online intrusion detection system.
+determine if an intrusion occurred.
+We used this approach to analyze simple intrusions on small provenance graphs and achieved promising results, however applying this technique to {\em user-to-remote} (u2r) intrusions
+from the much larger 1998 DARPA Intrusion Detection data sets  \cite{darpa} resulted in normal and abnormal behavior being indistinguishable.
+We discuss why we received the results we obtained for each of these workloads, and we propose requirements on provenance graph models and statistics that are necessary for the model to scale to large provenance graphs.
 
@@
 file nodes have edges directed towards previous versions of the file and to the nodes of the
-processes that modified the file. See Figure \ref{graph-simple} for an illustration of these provenance
+processes that modified the file. See Figure 1 for an illustration of these provenance
 graph node and edge types.
@@
 Allen et. al. analyze
-basic statistics on provenance graphs (e.g., number of nodes, number of edges, ... TODO) in order to facilitate
+basic statistics on provenance graphs (e.g., number of nodes, number of edges, max number of edges per node, etc.) in order to facilitate
 easier manual intrusion detection by a human \cite{provstat}. Their work, however, does not present a way of using provenance
-graph to automatically determine intrusions. Other systems use provenance data to determine the scope of an
+graphs to automatically determine intrusions. Other systems use provenance data to determine the scope of an
 intrusion. For example, Backtracker \cite{backtracker} uses provenance graphs to build causality graphs of intrusions: once the
@@
 various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
-on {\em user-to-local} (u2l) intrusions (intrusions that provide unauthorized access to local superuser (root) privileges),
-and {\em remote-to-local} (r2l) intrusions (intrusions that provide unauthorized access from a remote machine).
+on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
 
@@
 \item Discuss which provenance graph statistics we chose to use, how we chose them, and why we chose them.
-\item Analyze the Parzen-Window technique with various provenance graph statistics on a real intrusion detection data set and evaluate its accuracy.
-\item Discuss the limitations of the approach and present future work to transform the technique into an online intrusion detection system.
+\item Analyze the Parzen-Window technique with various provenance graph statistics on a small data set and a real intrusion detection data set and evaluate its accuracy on both.
+\item Discuss the limitations of the approach and propose requirements for provenance graph intrusion detection models to scale to larger-sized graphs.
 \end{enumerate}
-
-POSSIBLY INCLUDE THIS TODO
-* the papers suck at finding u2r
-* because they don't look at provenance graph structure
-* we can do better?
 
@@
 in the following section. For the common variance of our Gaussian kernels, we chose
-the variance to be the variance of the data set $D$ for each unique process name. 
+the variance to be the variance of the data set $D_P$ for each unique process name $P$. 
 We vary the features we used and values of $\psi$ and evaluate its performance on an intrusion detection data set.
@@
 
-We used the following procedure to implement our Parzen-Window density estimates and
-intrusion/normal decision.
+We used the following procedure to implement our Parzen-Window density estimates and to decide whether a node is behaving normally or abnormally.
 \begin{enumerate}
@@
 nodes in the provenance graph containing the normally-behaving workload.
-\item We aggregated the counts by the name of the process (the path to the executable). For a given process $P$, denote the
+\item We aggregated the counts by the name of the process (the path to the executable). For a given process name $P$, denote the
 set of counts for that process name as $D_P$.
@@
 \subsubsection{hello}
-We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution and possibly obtain root access.
-We created a simple \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
+We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution to possibly obtain root access.
+We used this as an guide to create a simple exploite \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
 
@@
 By unexpectedly aliasing an important command, a user can unknowingly grant root access to a malicious program.
-To emulate this sort of attack, we replaced \texttt{/usr/bin/gcc} with our own version of gcc, which, in addition to compiling the user's code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
+To emulate this system administration attack, we replaced \texttt{/usr/bin/gcc} with our own version of gcc, which, in addition to compiling the user's code (via a call to gcc), logs that the user has made a request and copies the files the user had tried to compile.
 
@@
 
-If the exploit does directly not generate additional files, however, statistics like "number of output files" will fail to be statistically significant.
+If the exploit does not directly generate additional files, however, statistics like ``number of output files" will fail to be statistically significant.
 Indeed, the two programs (hello and mcrypt) that did not generate additional files were not flagged by that measure.
@@
 
-Because of this exploratory analysis, we chose to implement Opsahl and ancestor centrality and analyze its accuracy on a real intrusion detection data set.
+Because of this exploratory analysis, we chose to implement Opsahl centrality and analyze its accuracy on a real intrusion detection data set.
 
@@
 audit of all the system calls made within a certain time period. The data also includes a list
-of specific moments in time $T$ and whether an intrusion occurred and if so, what kind of intrusion.
+of specific moments in time $T$ and the type of intrusion that occurred, if any.
 
@@
 the other kinds of attacks were network-based and would not appear in the provenance data. 
-We identify each accuracy measurement with a tuple:
-$$(\mbox{centrality metric}, \psi, W)$$
-where the centrality metric is one of Opsahl, indegree, or ancestor, $\psi$ is our false detection rate for our Parzen-Window
+We identify each workload with a tuple:
+$$(\psi, W)$$
+where $\psi$ is our false detection rate for our Parzen-Window
 estimates, and $W$ is the time window that we use to form time intervals to look for anomalous nodes. We used
@@
 \item Use BSM traces to build provenance graphs, in a similar fashion to SPADE \cite{spade}. 
-\item Compute the centralities of the nodes in the graph using the given centrality metric, and create a mapping from each process
+\item Compute the Opsahl centrality of the nodes in the graph and create a mapping from each process
 name $P$ to a list of log-likelihood values $L(y)$, where $y$ is the computer centrality measurement for each node in the graph having process name $P$.
@@
 \item For times $T$ in the data set listed as having an intrusion, our intrusion detection system considers that time $T$ as having an intrusion
-if there are {\em sufficient} number of nodes in the $[T - W, T + W]$ window considered anomalous. A candidate anomaly set has a 
-{\em sufficient} number of anomalous nodes if TODO...
+if a {\em majority} of the nodes in the candidate anomaly set are considered anomalous.
 \end{enumerate}
 
-For each $(\mbox{centrality metric}, \psi, W)$, we compute:
+For each $(\psi, W)$, we compute:
 \begin{itemize}
@@
 
-We evaluated the accuracy of our system on various values of $\psi$ and $W$.
+We evaluated the accuracy of our system on various values of $\psi$ and $W = 10s, 100s, 1000s$.
 
@@
 
-TODO
+\noindent
+\begin{figure}
+  \label{tpr}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/tpr.png}
+    \caption{True Positive Rate. The window size 10s, 100s, 1000s did not impact the true positive rate at all for various values of $\psi$.} 
+\end{figure}
 
-Need these graphs:
-[1]:
-- fix psi
-- x-axis is W (probably log-scale, as W = 10ms, 1000ms, 100000ms, 10 ** N ms)
-- y-axis is [0,1]
-- overlay true positive rate for each of the centrality metrics in different color
+\noindent
+\begin{figure}
+  \label{fpr}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/fpr.png}
+    \caption{False Positive Rate. The window size 10s, 100s, 1000s did not impact the false positive rate very
+    much for various values of $\psi$.} 
+\end{figure}
 
-[2]: repeat [1] but plot false positive rate instead
+We plot the true positive rates for various values of $\psi$ and window sizes $W$ in Figure 2, and similarly for the false positive rates in Figure 3.
 
-[3-4] repeat [1-2] with a different value of psi
-
-[5-6] repeat [1-2] with another value of psi
+We note that the true positive rate and false positive rates are nearly identical as we vary $\psi$.
+Furthermore, the true positive rate stayed exactly the same as we varied our window size $W$ from $10s, 100s, 1000s$.
+Thus, the window size had little to no impact on the TPR and FPR.
+As we increase $\psi$, the false detection rate, we expect to see more positives, and this is exactly what we see in these graphs.
 
@@
 
-TODO
+We note that the true positive rates and false positive rates were not what we expected.
+We were surprised to see that the true positive rates and false positive rates were nearly identical to each other.
+The fact that the TPR and FPR were identical to each other suggests that our Parzen-Window model using Opsahl centrality could not reliably distinguish between normal and abnormal behavior in the large graphs generated from the DARPA data set.
 
-\subsection{Limitations}
+We believe that the reason we had more promising results on our simulated intrusion data sets from PASS was because the provenance graphs from these workloads were much smaller (on the order of 5000 nodes).
+As a result of the much smaller size, small changes in provenance graph structure from an intrusion has a much profounder impact on the smaller provenance graphs.
+Thus, intrusions on our PASS workload had a much higher impact on the centrality calculations and resulted in more anomalous behavior in our Parzen-Window model.
+In the DARPA data set, we faced much larger provenance graphs (on the other of 20,000 nodes), and because of the much larger size, the effect of small changes in provenance graph structure from an intrusion had a much smaller impact on the overall structure of the graph.
+Intrusions did not change the centrality metrics very much in large graphs; therefore anomalous behavior went undetected, regardless of the window size we chose.
+Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical.
+From these results, we propose the following requirements for provenance graph statistics and models in order to be useful for large provenance graphs:
 
-TODO
+\begin{itemize}
+\item The effects of the provenance graph statistic should grow as the size of the provenance graph grows.
+Since provenance graphs are meant to be long-lived, a small anomaly in the provenance graph will be drowned out by the overall large structure of the graph.
+Thus, if a statistic only captures a small local property of the graph, that property will become negligible and the anomaly will go undetected.
+As a result, a provenance statistic should become {\em more} anomalous as more provenance is collected.
+
+\item The statistic must be adaptive and computable online.
+In order to make a truly useful online intrusion detection system, the statistic must be recomputed quickly over all the nodes in the graph as the provenance graph grows.
+Possible examples of such a statistic could be progressive forms of PageRank where the rank is computed relative to a starting node.
+
+\item The model must be able to predict anomalous behavior by being trained with only examples of normal behavior.
+As stated by Yeung et.
+al., intrusion detection is a form of novelty detection where access to abnormal patterns is very difficult \cite{parzen}.
+As a result, successful intrusion detection models like Somayaji and Forrest's model must be able to characterize abnormal behavior without knowing precisely what that means.
+\end{itemize}
 
@@
 Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
+\subsection{Different values for the density variance}
+In our system, we used the variance $D$ of each data set as the variance of our kernel.
+To possibly get better results, one might try using different values for the variance, possible values that capture more granularity of students.
 \subsection{Online Intrusion Detection}
@@
 \section{Conclusion}
-TODO
+
+In this paper, we frame the intrusion detection problem as a novelty detection problem, and we propose a model to use Parzen-Window estimates on provenance graph statistics to detect intrusions.
+Our model showed some success on small provenance graphs, but did not perform well on much larger provenance graphs.
+We discuss why we obtained these results, and we use these lessons to propose requirements for intrusion detection models that use provenance graphs.
 

### 2013-12-10 writeup/paper-final.tex
beginning final proofread

@@
 We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
+See Figure 2 for examples of provenance graphs for the normally behaving and exploited versions of the program.
 
@@
 We again exported the provenance generated from this exploited workload.
+See Figure 3 for examples of provenance graphs for the normally behaving and exploited versions of the program.
+
+\begin{figure*}
+  \label{hello-orbiter}
+  \caption{Descendants of a \texttt{hello} program.
+The top image shows multiple invocations of a \texttt{hello} behaving normally when invoked with \texttt{cat foo.txt | ./hello > c.txt}.
+The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.}
+  \centering
+    \includegraphics[width=\textwidth]{img/hello.png}
+\end{figure*}
+\begin{figure*}
+  \label{mcrypt-orbiter}
+  \caption{Descendants of an unexploited mcrypt-style program.
+Simply an output file.
+Below, descendants of an exploited mcrypt-style program.
+In addition to an (offscreen) output file, there is also an outgoing \texttt{/bin/sh} process, with its own children.}
+  \centering
+    \includegraphics[width=\textwidth]{img/mcrypt.png}
+\end{figure*}
 
@@
 \subsection{Selecting Provenance Statistics}
+
+\begin{figure*}
+  \label{kde-example}
+  \centering
+    \includegraphics[width=\textwidth]{img/hist.png}
+      \caption{Example Parzen-Window density estimates of a centrality measure and process statistic for the \texttt{gcc} exploit.
+The green bars represent the raw histogram counts, the blue curve represents the density estimate, and the red vertical line represents the estimation of a node of interest.}
+\end{figure*}
 
@@
 
-In Table 1, we show a summary of the statistics that had any statistical difference between normal and abnormally behaving nodes.
+In Table 1, we show a summary of the statistics that had any statistical difference between normal and abnormally behaving nodes, and Figure 4 highlights the most salient statistics for the \texttt{gcc} exploit.
 The buffer overflows showed the most abnormal behavior in the number of outgoing processes and the centrality measures.
@@
 
-We plot the true positive rates for various values of $\psi$ and window sizes $W$ in Figure 2, and similarly for the false positive rates in Figure 3.
+We plot the true positive rates for various values of $\psi$ and window sizes $W$ in Figure 5, and similarly for the false positive rates in Figure 6.
 

### 2013-12-10 writeup/paper-final.tex
done

@@
 from the much larger 1998 DARPA Intrusion Detection data sets  \cite{darpa} resulted in normal and abnormal behavior being indistinguishable.
-We discuss why we received the results we obtained for each of these workloads, and we propose requirements on provenance graph models and statistics that are necessary for the model to scale to large provenance graphs.
+We discuss why we received the results we obtained for each of these workloads, and we propose requirements on intrusion detection models using provenance graph models that are necessary for the model to scale to large provenance graphs.
 
@@
 of normal patterns \cite{parzen}. However, Yeung et. al. note that in the case of intrusion detection, novel
-or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach
-for non-parametric density estimation. Using this technique, they obtain a high degree of success
+or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach that allows them to characterize abnormal behavior by only knowing normal behavior.
+Using this technique, they obtain a high degree of success
 in identifying various intrusion types \cite{parzen}. Because obtaining abnormal patterns is difficult, 
@@
 We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution to possibly obtain root access.
-We used this as an guide to create a simple exploite \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
+We used this as an guide to create a simple exploited \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
 
@@
 Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical.
-From these results, we propose the following requirements for provenance graph statistics and models in order to be useful for large provenance graphs:
+From these results, we propose the following requirements for intrusion detection models that use provenance graph statistics in order to be useful for large provenance graphs:
 
@@
 By allowing the model to evolve over time, the system
-can combat an {\em autoimmune disorders} in which attackers purposely invoke programs in weird ways to fool the system into thinking the programs are
+can combat {\em autoimmune disorders} in which attackers purposely invoke programs in weird ways to fool the system into thinking the programs are
 behaving abnormally.
@@
 \section{Acknowledgements}
-We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project and for
-their help in working with PASS and analyzing provenance graphs. We also want to thank the members of the
+We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project, for
+their help in working with PASS, and for their insight in analyzing provenance graphs. We also want to thank the members of the
 CS261 Program Committee for their feedback on the initial draft of this paper.

### 2013-12-10 writeup/paper-final.tex
Spelling, 'et.' fixes

@@
   \and
-    R. J. Aquino\\
+    R.J. Aquino\\
     Harvard University\\
@@
 Existing intrusion detection systems (IDS) use provenance data only to a limited extent.
-Allen et. al. analyze
+Allen et al. analyze
 basic statistics on provenance graphs (e.g., number of nodes, number of edges, max number of edges per node, etc.) in order to facilitate
@@
 an unknown test pattern is produced by an underlying distribution corresponding to a training set
-of normal patterns \cite{parzen}. However, Yeung et. al. note that in the case of intrusion detection, novel
+of normal patterns \cite{parzen}. However, Yeung et al. note that in the case of intrusion detection, novel
 or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach that allows them to characterize abnormal behavior by only knowing normal behavior.
@@
 More recent work make use of the structure of provenance as a directed graph for a semi-automatic form
-of intrusion detection, and to determine the scope of an intrusion. King et. al. developed Backtracker \cite{backtracker}
+of intrusion detection, and to determine the scope of an intrusion. King et al. developed Backtracker \cite{backtracker}
  a modified
@@
 
-Allen et. al. make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
+Allen et al. make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
 in manual intrusion-detection \cite{provstat}. Given a data set of many provenance graphs, they calculated basic statistics
@@
 To address the difficulty of obtaining abnormal patterns for novelty detection on intrusion detection, we borrow
-the Parzen-Window model approach proposed by Yeung et. al.
+the Parzen-Window model approach proposed by Yeung et al.
 They attempt to solve the lack of abnormal patterns problem by using Parzen-Windows to build {\em nonparametric}
@@
 
-We were inspired by the work proposed by Allen et. al. who use simple provenance graph statistics to aide
+We were inspired by the work proposed by Allen et al. who use simple provenance graph statistics to aide
 in manual intrusion detection by humans. We extend their idea to make use of the typed node and typed edge
@@
 when it normally does not), and we hypothesize that this can manifest as unusual changes in centrality
-for the offending node in the provenance graph. Macko et. al. note the following 3 observations unique to provenance graphs \cite{clustering}:
+for the offending node in the provenance graph. Macko et al. note the following 3 observations unique to provenance graphs \cite{clustering}:
 \begin{enumerate}
@@
 a mounted volume that automatically collects provenance of all files within that volume.
-To make data collection easier, we used PASSv2 \cite{passv2}, which provides an Ubuntu virtual machine with
-the PASS volume already installed.
+To make data collection easier, we used PASSv2 \cite{passv2}, which provides an Ubuntu virtual machine with the PASS volume already installed.
 
@@
 We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution to possibly obtain root access.
-We used this as an guide to create a simple exploited \texttt{hello} program that echos standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
+We used this as an guide to create a simple exploited \texttt{hello} program that echoes standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
 
@@
 \item The model must be able to predict anomalous behavior by being trained with only examples of normal behavior.
-As stated by Yeung et.
-al., intrusion detection is a form of novelty detection where access to abnormal patterns is very difficult \cite{parzen}.
+As stated by Yeung et al., intrusion detection is a form of novelty detection where access to abnormal patterns is very difficult \cite{parzen}.
 As a result, successful intrusion detection models like Somayaji and Forrest's model must be able to characterize abnormal behavior without knowing precisely what that means.
@@
 \section{Acknowledgements}
-We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project, for
-their help in working with PASS, and for their insight in analyzing provenance graphs. We also want to thank the members of the
-CS261 Program Committee for their feedback on the initial draft of this paper.
+We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project, for their help in working with PASS, and for their insight in analyzing provenance graphs.
+We also want to thank the members of the CS261 Program Committee for their feedback on the initial draft of this paper.
 

### 2013-12-10 writeup/paper-final.tex
Outline

@@
 \end{enumerate}
+
+\subsection{Outline}
+In Section 2 we discuss work related to both provenance collection and intrusion detection.
+In Section 3 we detail our initial exploratory experiments on small data sets, determining which provenance graph features we wanted to explore further in designing our approach, as well as the implementation details of our initial experimental setup.
+In Section 4, we discuss the setup, results, and findings of our main experiment, in which we tested our approach on existing large intrusion detection data sets.
+In Section 5 we discuss future work and directions that this project has motivated, and in Section 6 we conclude.
 

### 2013-12-10 writeup/paper-final.tex
u2l -> u2r

@@
 various provenance graph statistics and graph centrality metrics, and we evaluate the success of this technique
-on {\em user-to-local} (u2l) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
+on {\em user-to-root} (u2r) intrusions, intrusions that provide unauthorized access to local superuser (root) privileges.
 
@@
 In this paper, we use this same model to build profiles of ``normal" behavior for each process. We use
-one-dimensional feature vectors $x$, and we use various provenance graph statistics and graph
+one-dimensional feature vectors $\m{x}$, and we use various provenance graph statistics and graph
 centrality metrics as our features. We describe the various statistics and graph centrality metrics we explored
@@
 \item {\bf Remote to Local (r2l)}. Unauthorized access from a machine, e.g., guessing passwords.
-\item {\bf User to Root (u2l)}. Unauthorized access to local superuser (root) privileges, e.g., various buffer overflow attacks.
+\item {\bf User to Root (u2r)}. Unauthorized access to local superuser (root) privileges, e.g., various buffer overflow attacks.
 \item {\bf Probing (probe)}. Surveillance and scanning, e.g. port scanning.
 \end{enumerate}
-Because we only collect provenance graphs local to the machine, we chose to limit our intrusion detection system only to u2l intrusions within this paper, as the other types of attacks involve exploiting the network and would not appear in provenance data.
+Because we only collect provenance graphs local to the machine, we chose to limit our intrusion detection system only to u2r intrusions within this paper, as the other types of attacks involve exploiting the network and would not appear in provenance data.
 

### 2013-12-10 writeup/paper-final.tex
More typos, added to discussin

@@
 We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution to possibly obtain root access.
-We used this as an guide to create a simple exploited \texttt{hello} program that echoes standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell.
+We used this as an guide to create a simple exploited \texttt{hello} program that echoes standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell or other process.
 
@@
 The top image shows multiple invocations of a \texttt{hello} behaving normally when invoked with \texttt{cat foo.txt | ./hello > c.txt}.
-The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.}
+The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.
+The gray circles are a provenance representation of a ``pipe''}
   \centering
@@
 Thus, intrusions on our PASS workload had a much higher impact on the centrality calculations and resulted in more anomalous behavior in our Parzen-Window model.
-In the DARPA data set, we faced much larger provenance graphs (on the other of 20,000 nodes), and because of the much larger size, the effect of small changes in provenance graph structure from an intrusion had a much smaller impact on the overall structure of the graph.
+In the DARPA data set, we faced much larger provenance graphs (on the order of 20,000 nodes), and because of the much larger size, the effect of small changes in provenance graph structure from an intrusion had a much smaller impact on the overall structure of the graph.
 Intrusions did not change the centrality metrics very much in large graphs; therefore anomalous behavior went undetected, regardless of the window size we chose.
 Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical.
+In addition, we found that many more nodes were being flagged as intrusions (indeed, a positive result meant that a majority of nodes in a given time range were considered anomalous).
+This result differed greatly from our preliminary exploration, where normal nodes were consistently identified as normal.
+We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so complete outliers were less present.
 From these results, we propose the following requirements for intrusion detection models that use provenance graph statistics in order to be useful for large provenance graphs:
@@
 We found that \texttt{gcc} forks processes for the linker/assembler that handles generating output, so \texttt{gcc} does not directly have output files.
-Similarly, when running a python script, the code file itself is not directly executed, it is read in by a python process, which then goes off and collects input, generates output, and forks additional processes.
-Thus, when running a python file, the file itself only has one node in the provenance graph, with many python process nodes accessing it.
+Similarly, when running a Python script, the code file itself is not directly executed, it is read in by a \texttt{python} process, which then goes off and collects input, generates output, and forks additional processes.
+Thus, when running a Python file, the file itself only has one node in the provenance graph, with many \texttt{python} process nodes accessing it.
 In order to account for these kinds of relationships, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
@@
 By generating histograms based on this information, we could catch subtler intrusions that rely on modified environment information.
-Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
+Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information (consider a program which does not limit the number or size of command line arguments, and uses this vulnerability to execute malicious code).
 \subsection{Different values for the density variance}
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 We want to thank Margo Seltzer and Daniel Margo for their guidance on the direction of the project, for their help in working with PASS, and for their insight in analyzing provenance graphs.

### 2013-12-10 writeup/paper-final.tex
% of intrusions

@@
     \includegraphics[width=0.5\textwidth]{img/graph-simple.png}
-    \caption{Example of a provenance graph, showing the types of nodes and edges common in provenance graphs} 
+    \caption{Example of a simple provenance graph, showing the types of nodes and edges common in provenance graphs} 
 \end{figure}
@@
 In addition, we found that many more nodes were being flagged as intrusions (indeed, a positive result meant that a majority of nodes in a given time range were considered anomalous).
+In Figure 7, we see that the percentage of nodes during exploit windows were identified similarly to those during windows containing no exploit.
+As expected, as PSI increases, so too does the number of events marked as exploits.
 This result differed greatly from our preliminary exploration, where normal nodes were consistently identified as normal.
 We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so complete outliers were less present.
+\begin{figure}
+  \label{intrusion-v-psi}
+  \centering
+    \includegraphics[width=0.5\textwidth]{img/intrusion_v_psi.png}
+    \caption{The percentage of nodes flagged as an intrusion for various values of PSI.
+This figure was generated from the Week 5 Monday data, using a window size of 10 seconds.
+Other window sizes showed similar results.
+Recall that our system flags events with a percentage of 0.5 or greater as ``anomalous''}
+\end{figure}
 From these results, we propose the following requirements for intrusion detection models that use provenance graph statistics in order to be useful for large provenance graphs:

### 2013-12-10 writeup/paper-final.tex
ran text through word to catch grammar mistakes. found a handful.

@@
 
-More recent work make use of the structure of provenance as a directed graph for a semi-automatic form
+More recent works make use of the structure of provenance as a directed graph for a semi-automatic form
 of intrusion detection, and to determine the scope of an intrusion. King et al. developed Backtracker \cite{backtracker}
@@
 We found that many of the popular exploits \cite{metasploit, exploitdb} involved buffer overflows and arbitrary code execution to possibly obtain root access.
-We used this as an guide to create a simple exploited \texttt{hello} program that echoes standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell or other process.
+We used this as a guide to create a simple exploited \texttt{hello} program that echoes standard in to standard out, and we gave it a buffer overflow vulnerability that allows a malicious user to fork a shell or other process.
 
@@
 The bottom image shows \texttt{hello} being exploited to fork an \texttt{ls} process.
-The gray circles are a provenance representation of a ``pipe''}
+The gray circles are a provenance representation of a ``pipe''.}
   \centering
@@
   \label{mcrypt-orbiter}
-  \caption{Descendants of an unexploited mcrypt-style program.
-Simply an output file.
+  \caption{Descendants of an unexploited mcrypt-style program - simply an output file.
 Below, descendants of an exploited mcrypt-style program.
@@
 new processes and output files through a remote shell), thus generate a higher Opsahl centrality metric for the offending node.
-The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that the exploit will be caught by our model.
+The more active the exploit is (say, if it forks a user shell), the more the centrality measure will be impacted, and the more likely it is that our model will catch the exploit.
 
@@
 Other window sizes showed similar results.
-Recall that our system flags events with a percentage of 0.5 or greater as ``anomalous''}
+Recall that our system flags events with a percentage of 0.5 or greater as ``anomalous''.}
 \end{figure}
@@
 By generating histograms based on this information, we could catch subtler intrusions that rely on modified environment information.
-Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information (consider a program which does not limit the number or size of command line arguments, and uses this vulnerability to execute malicious code).
+Additionally, if an exploit relies on a particular argument or argument patterns, this information might be reflected in the ARGV information (consider a program which does not limit the number or size of command line arguments, and uses this vulnerability to execute malicious code).
 \subsection{Different values for the density variance}

### 2013-12-10 writeup/paper-final.tex
Added BSM to provenance section

@@
 We evaluated the accuracy of our system on various values of $\psi$ and $W = 10s, 100s, 1000s$.
+\subsection{BSM to Provenance Conversion}
+For our experiment, we needed to convert the BSM traces into a provenance graph.
+The designers of SPADE \cite{spade} used this same approach, so we ported their conversion code to our system.
+The BSM data only contains PIDs, however, not process names.
+While SPADE used \texttt{ps} to determine the process names for a particular PID, we needed to find another way, because we did not have access to the \texttt{ps} output for all timestamps.
+We initially tried to use minutely \texttt{ps} output provided by DARPA, but too many processes started and completed within this window, leaving them unnamed in our graph.
+We then parsed the BSM traces directly, looking for \texttt{execve} calls, which we could use to name our process nodes.
 
+The SPADE code we used as a model parses the BSM file, adding nodes when processes and files are introduced (e.g.
+via \texttt{fork} or \texttt{creat}) and adding edges when processes and files interact (e.g.
+via \texttt{write} and \texttt{fork}).
+The types of edges were very similar to those found in the PASS data - namely ``triggered by'' edges between process nodes and ``used'' edges between process nodes and file nodes.
 \subsection{Results}

### 2013-12-10 writeup/paper-final.tex
Lots of copy editing

@@
 \begin{abstract}
-Security rests on being able to identify potentially malicious behavior.
+Security depends on being able to identify potentially malicious behavior.
 One way to identify malicious behavior is to compare system behavior to known normal behavior.
@@
 the artifact came to be in its present state.
-We present an intrusion detection approach to find these anomalies by analyzing provenance graphs statistics.
+We present an intrusion detection approach to find these anomalies by analyzing provenance graph statistics.
 We use a Parzen-Window approach \cite{parzen} on various provenance graph statistics \cite{clustering}
@@
 determine if an intrusion occurred.
-We used this approach to analyze simple intrusions on small provenance graphs and achieved promising results, however applying this technique to {\em user-to-remote} (u2r) intrusions
+We used this approach to analyze simple intrusions on small provenance graphs and achieved promising results.
+However, applying this technique to {\em user-to-remote} (u2r) intrusions
 from the much larger 1998 DARPA Intrusion Detection data sets  \cite{darpa} resulted in normal and abnormal behavior being indistinguishable.
@@
 in systems to gain unintended privileged access. As a result, system designers and 
-administrators place a large effort in securing systems and preventing intrusions. However, intrusions
+administrators make a large effort to secure systems and prevent intrusions. However, intrusions
 inevitably occur because of bugs or flaws within the system, and as a result, system administrators want to
@@
 Intrusions often leave behind digital artifacts, e.g., unusual output or log files, a process executed
-with an unusual arguments or environment, unusual system call activity by a process. In addition to
+with unusual arguments or environment, or unusual system call activity by a process. In addition to
 unusual digital artifacts, we hypothesize that
@@
 easier manual intrusion detection by a human \cite{provstat}. Their work, however, does not present a way of using provenance
-graphs to automatically determine intrusions. Other systems use provenance data to determine the scope of an
+graphs to automatically determine whether intrusions have occurred. Other systems use provenance data to determine the scope of an
 intrusion. For example, Backtracker \cite{backtracker} uses provenance graphs to build causality graphs of intrusions: once the
 system has
-determined an intrusion has occurred, the system follows the causality graph backwards determine the original source of the attack, and 
-then it follows the causality graph forwards to determine all the objects tainted by the attack. However, the authors 
+determined an intrusion has occurred, the system follows the causality graph backwards to determine the original source of the attack, and 
+then it follows the causality graph forward to determine all the objects tainted by the attack. However, the authors 
 do not present a way of using provenance data to automatically detect intrusions.
@@
     \includegraphics[width=0.5\textwidth]{img/graph-simple.png}
-    \caption{Example of a simple provenance graph, showing the types of nodes and edges common in provenance graphs} 
+    \caption{Example of a simple provenance graph, showing the types of nodes and edges common in provenance graphs.} 
 \end{figure}
@@
 of normal patterns \cite{parzen}. However, Yeung et al. note that in the case of intrusion detection, novel
-or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach that allows them to characterize abnormal behavior by only knowing normal behavior.
-Using this technique, they obtain a high degree of success
+or abnormal patterns are typically difficult to obtain, and as a result, they present a {\em Parzen-Window} approach that allows them to characterize abnormal behavior using only normal behavior.
+With this technique, they obtain a high degree of success
 in identifying various intrusion types \cite{parzen}. Because obtaining abnormal patterns is difficult, 
@@
 \item Discuss which provenance graph statistics we chose to use, how we chose them, and why we chose them.
-\item Analyze the Parzen-Window technique with various provenance graph statistics on a small data set and a real intrusion detection data set and evaluate its accuracy on both.
+\item Analyze the Parzen-Window technique with various provenance graph statistics on a small data set and a real intrusion detection data set and evaluate the accuracy we achieve for both.
 \item Discuss the limitations of the approach and propose requirements for provenance graph intrusion detection models to scale to larger-sized graphs.
@@
 \subsection{Outline}
-In Section 2 we discuss work related to both provenance collection and intrusion detection.
-In Section 3 we detail our initial exploratory experiments on small data sets, determining which provenance graph features we wanted to explore further in designing our approach, as well as the implementation details of our initial experimental setup.
+In Section 2, we discuss work related to both provenance collection and intrusion detection.
+In Section 3, we detail our initial exploratory experiments on small data sets, determining which provenance graph features we wanted to explore further in designing our approach, as well as the implementation details of our initial experimental setup.
 In Section 4, we discuss the setup, results, and findings of our main experiment, in which we tested our approach on existing large intrusion detection data sets.
-In Section 5 we discuss future work and directions that this project has motivated, and in Section 6 we conclude.
+In Section 5, we discuss future work and directions that this project has motivated, and in Section 6 we conclude.
 
@@
 system calls can be seen as a limited form of provenance, as many systems build provenance graphs from
-system call traces \cite{spade}. By analyzing the sequences
+system call traces (as done in SPADE \cite{spade}). By analyzing the sequences
 of system calls a process makes over time, their system builds a profile of ``normal" process behavior. When
@@
 Allen et al. make use of simple provenance graph statistics to categorize provenance graphs and to assist a human
-in manual intrusion-detection \cite{provstat}. Given a data set of many provenance graphs, they calculated basic statistics
-on the graphs (e.g., total nodes, total edges, max incoming edges on a node, max outgoing edges on a node,
+in manual intrusion detection \cite{provstat}. Given a data set of many provenance graphs, they calculated basic statistics
+of the graphs (e.g., total nodes, total edges, max incoming edges on a node, max outgoing edges on a node,
 node/edge ratio, average total edges per node, etc.). Their approach allows a human to more easily
@@
 
-To address the difficulty of obtaining abnormal patterns for novelty detection on intrusion detection, we borrow
+To address the difficulty of obtaining abnormal patterns for intrusion detection, we borrow
 the Parzen-Window model approach proposed by Yeung et al.
@@
 one-dimensional feature vectors $\m{x}$, and we use various provenance graph statistics and graph
-centrality metrics as our features. We describe the various statistics and graph centrality metrics we explored
-in the following section. For the common variance of our Gaussian kernels, we chose
+centrality metrics as our features. In the following section, we describe the various statistics and graph centrality metrics we explored. For the common variance of our Gaussian kernels, we chose
 the variance to be the variance of the data set $D_P$ for each unique process name $P$. 
@@
 one way of observing this is through {\em graph centrality metrics}. A {\em graph centrality metric} is a function
-of a node and its containing graph, and intuitively represents how {\em central} or {\em important} the node
+of a node and the graph that contains it, and the value of the function intuitively represents how {\em central} or {\em important} the node
 is in the graph. In terms of provenance graphs, processes of the same program typically behave similarly to one another
@@
 Because a node's importance is based on its output descendants, then the number of times a node was used as input to another node indicates a simple measure of importance.
-\item \textbf{Age}. They note that large "jumps" in timestamps or age might indicate breaks between tasks, but this interpretation is complicated when scripts execute tasks in short succession.
-\item \textbf{Ancestor Centrality}. This measures the total number of descendants of a node, normalized by the total number of nodes in a graph.
-\item \textbf{Opsahl's Closeness Centrality}. This accounts for the number of edges between a node and its descendants. To compute this, we calculate
+\item \textbf{Age}. They note that large ``jumps'' in timestamps or age might indicate breaks between tasks, but this interpretation is complicated when scripts execute tasks in short succession.
+\item \textbf{Ancestor Centrality}. This metric measures the total number of descendants of a node, normalized by the total number of nodes in a graph.
+\item \textbf{Opsahl's Closeness Centrality}. This metric accounts for the number of edges between a node and its descendants. To compute this, we calculate
 $$CC'(v) = \sum_{x \in V -  \{v\}} (d'_{xv})^{-1}$$
@@
 the system looks for instances of deviation from normal behavior and handles them appropriately
-(e.g., terminating the process, slowing down the process). In our work, we derive normal profiles of process behavior
+(e.g., terminating the process or slowing down the process). In our work, we derive normal profiles of process behavior
 from provenance graphs, and we use a Parzen-Window model on various provenance graph statistics
@@
 a mounted volume that automatically collects provenance of all files within that volume.
-To make data collection easier, we used PASSv2 \cite{passv2}, which provides an Ubuntu virtual machine with the PASS volume already installed.
+To make data collection easier, we used PASSv2 \cite{passv2}, which provides an Ubuntu virtual machine with a PASS volume already installed.
 
-For each of our intrusions, we ran it normally multiple times and collected the provenance. We then ran the
+We ran each of our exploited programs normally (i.e. without taking advantage of the exploit) multiple times and collected the provenance. We then ran the
 intrusion with the exploit multiple times and collected the new provenance.
@@
 \begin{enumerate}
-\item For each statistic that we implemented, we computed the counts for that statistic for all the
+\item For each statistic that we implemented, compute the counts for that statistic for all the
 nodes in the provenance graph containing the normally-behaving workload.
-\item We aggregated the counts by the name of the process (the path to the executable). For a given process name $P$, denote the
+\item Aggregated the counts by the name of the process (the path to the executable). For a given process name $P$, denote the
 set of counts for that process name as $D_P$.
 \item For each $y \in D_P$, calculate the log-likelihood $L_P(y)$.
-\item For a test node with process name $P$ in the exploited-workload provenance graph, we calculate the value $x$ of the statistic in its graph, and then compute $L_P(x)$. 
+\item For a test node with process name $P$ in the exploited-workload provenance graph, calculate the value $x$ of the statistic in its graph, and then compute $L_P(x)$. 
 If the ratio of the number of nodes $y$ with $L_P(y) \leq L_P(x)$,
-to the total number of nodes in $D_P$ is greater than the false detection rate $\psi$, we label the node as normal. Otherwise we label the node as
+to the total number of nodes in $D_P$ is greater than the false detection rate $\psi$, label the node as normal. Otherwise label the node as
 an intrusion.
@@
 
-We ran \texttt{hello} normally 40 times on various inputs and collected the provenance times.
-We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process, and saved the standard out of that process to a file.
+We ran \texttt{hello} normally 40 times on various inputs and collected the provenance.
+We then exploited the buffer overflow vulnerability in \texttt{hello} to spawn an \texttt{ls} process and saved the standard out of that process to a file.
 See Figure 2 for examples of provenance graphs for the normally behaving and exploited versions of the program.
@@
 
-We ran our program on over 200 good inputs, and stored the provenance data.
+We ran our program on over 200 good inputs and stored the provenance data.
 We then ran our program on two bad inputs, first to call \texttt{ls} as a Proof of Concept, and again to call \texttt{/bin/sh} to demonstrate that more severe attacks can be performed.
@@
 
-For our system administration attack, we compiled the PASS toolset using \texttt{gcc}, and exported the provenance from these runs.
+For our system administration attack, we compiled the PASS toolset using \texttt{gcc} and exported the provenance from these runs.
 We then replaced \texttt{gcc} with our exploited version, compiled the toolset several times with different configurations, and exported the provenance again.
@@
 
-In Table 1, we show a summary of the statistics that had any statistical difference between normal and abnormally behaving nodes, and Figure 4 highlights the most salient statistics for the \texttt{gcc} exploit.
+In Table 1, we show a summary of the statistics that had any statistical difference between normally and abnormally behaving nodes.
+Figure 4 highlights the most salient statistics for the \texttt{gcc} exploit.
 The buffer overflows showed the most abnormal behavior in the number of outgoing processes and the centrality measures.
 This makes sense, as arbitrary code execution will generate additional processes.
-Of the centrality metrics we implemented, the Opsahl Centrality metric did the best in distinguishing an exploited node from normally behaving nodes.
+Of the centrality metrics we implemented, the Opsahl Centrality metric performed best in distinguishing an exploited node from normally behaving nodes.
 Because the Opsahl Centrality metric accounts for the number of edges between a node and all of its descendants, this centrality metric is very sensitive to the addition of nodes.
@@
 \item For all nodes in the candidate anomaly set, find the process name $P$ of the node, and use our Parzen-Window approach on process
-name $P$ to determine if the node is anomalous. We evaluate whether a node is normal or anomalous using the Parzen-Windows built from the graph
+name $P$ to determine if the node is anomalous. Evaluate whether a node is normal or anomalous using the Parzen-Windows built from the graph
 containing the node.
@@
 We evaluated the accuracy of our system on various values of $\psi$ and $W = 10s, 100s, 1000s$.
+
 \subsection{BSM to Provenance Conversion}
@@
 While SPADE used \texttt{ps} to determine the process names for a particular PID, we needed to find another way, because we did not have access to the \texttt{ps} output for all timestamps.
-We initially tried to use minutely \texttt{ps} output provided by DARPA, but too many processes started and completed within this window, leaving them unnamed in our graph.
+We initially tried to use \texttt{ps} output provided by DARPA, but they only provided data at 60 second intervals, and too many processes started and completed within this window, leaving them unnamed in our graph.
 We then parsed the BSM traces directly, looking for \texttt{execve} calls, which we could use to name our process nodes.
@@
 via \texttt{write} and \texttt{fork}).
-The types of edges were very similar to those found in the PASS data - namely ``triggered by'' edges between process nodes and ``used'' edges between process nodes and file nodes.
+The types of edges were very similar to those found in the PASS data - namely, ``triggered by'' edges between process nodes and ``used'' edges between process nodes and file nodes.
 \subsection{Results}
@@
     \includegraphics[width=0.5\textwidth]{img/tpr.png}
-    \caption{True Positive Rate. The window size 10s, 100s, 1000s did not impact the true positive rate at all for various values of $\psi$.} 
+    \caption{True Positive Rate. The window size 10s, 100s, 1000s did not impact the true positive rate in any way for various values of $\psi$.} 
 \end{figure}
@@
     \includegraphics[width=0.5\textwidth]{img/fpr.png}
-    \caption{False Positive Rate. The window size 10s, 100s, 1000s did not impact the false positive rate very
-    much for various values of $\psi$.} 
+    \caption{False Positive Rate. The window size 10s, 100s, 1000s did not impact the false positive rate significantly for various values of $\psi$.} 
 \end{figure}
 
-We plot the true positive rates for various values of $\psi$ and window sizes $W$ in Figure 5, and similarly for the false positive rates in Figure 6.
+We plot the true positive rates for various values of $\psi$ and window sizes $W$ in Figure 5 and similarly for the false positive rates in Figure 6.
 
@@
 
-We believe that the reason we had more promising results on our simulated intrusion data sets from PASS was because the provenance graphs from these workloads were much smaller (on the order of 5000 nodes).
+We believe that the reason we had more promising results on our simulated intrusion data sets from PASS was because the provenance graphs from these workloads were much smaller (on the order of 5,000 nodes).
 As a result of the much smaller size, small changes in provenance graph structure from an intrusion has a much profounder impact on the smaller provenance graphs.
@@
 Intrusions did not change the centrality metrics very much in large graphs; therefore anomalous behavior went undetected, regardless of the window size we chose.
-Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical.
+Since our system could not distinguish between anomalous and normal behavior, the true positive rates and false positive rates are essentially identical.
 In addition, we found that many more nodes were being flagged as intrusions (indeed, a positive result meant that a majority of nodes in a given time range were considered anomalous).
-In Figure 7, we see that the percentage of nodes during exploit windows were identified similarly to those during windows containing no exploit.
-As expected, as PSI increases, so too does the number of events marked as exploits.
+In Figure 7, we see that the percentage of nodes during exploit windows that were identified as intrusions is very similar to the percentage identified during windows containing no exploit.
+As expected, as $\psi$ increases, so too does the number of events marked as exploits.
 This result differed greatly from our preliminary exploration, where normal nodes were consistently identified as normal.
-We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so complete outliers were less present.
+We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so the statistics contained fewer complete outliers.
 \begin{figure}
@@
     \includegraphics[width=0.5\textwidth]{img/intrusion_v_psi.png}
-    \caption{The percentage of nodes flagged as an intrusion for various values of PSI.
-This figure was generated from the Week 5 Monday data, using a window size of 10 seconds.
+    \caption{The percentage of nodes flagged as intrusions for various values of $\psi$.
+This figure was generated from the Week 5 Monday DARPA data, using a window size of 10 seconds.
 Other window sizes showed similar results.
@@
 Since provenance graphs are meant to be long-lived, a small anomaly in the provenance graph will be drowned out by the overall large structure of the graph.
-Thus, if a statistic only captures a small local property of the graph, that property will become negligible and the anomaly will go undetected.
+Thus, if a statistic only captures a small local property of the graph, that property will become negligible, and the anomaly will go undetected.
 As a result, a provenance statistic should become {\em more} anomalous as more provenance is collected.
@@
 Some processes fork helper processes that do the real work (such as \texttt{gcc}).
-We found that \texttt{gcc} forks processes for the linker/assembler that handles generating output, so \texttt{gcc} does not directly have output files.
-Similarly, when running a Python script, the code file itself is not directly executed, it is read in by a \texttt{python} process, which then goes off and collects input, generates output, and forks additional processes.
-Thus, when running a Python file, the file itself only has one node in the provenance graph, with many \texttt{python} process nodes accessing it.
-In order to account for these kinds of relationships, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node, and then aggregate them.
+We found that \texttt{gcc} forks processes for the linker and assembler that handle generating output, so \texttt{gcc} does not directly have output files.
+Similarly, when running a Python script, the code file itself is not directly executed; it is read in by a \texttt{python} process, which then goes off and collects input, generates output, and forks additional processes.
+Thus, when running a Python file, the file itself only has one node in the provenance graph with many \texttt{python} process nodes accessing it.
+In order to account for these kinds of relationships, we could calculate the relevant statistics for the test node and all nodes at distance $N$ from the test node and then aggregate them.
 By utilizing information about neighboring nodes, we hope to detect intrusions that involve more complicated relationships
@@
 By generating histograms based on this information, we could catch subtler intrusions that rely on modified environment information.
-Additionally, if an exploit relies on a particular argument or argument patterns, this information might be reflected in the ARGV information (consider a program which does not limit the number or size of command line arguments, and uses this vulnerability to execute malicious code).
+Additionally, if an exploit relies on a particular argument or argument pattern, this information might be reflected in the ARGV information (consider a program which does not limit the number or size of command line arguments and a user who uses this vulnerability to execute malicious code).
 \subsection{Different values for the density variance}
 In our system, we used the variance $D$ of each data set as the variance of our kernel.
-To possibly get better results, one might try using different values for the variance, possible values that capture more granularity of students.
+To possibly get better results, one might try using different values for the variance that capture more granularity of values.
 \subsection{Online Intrusion Detection}
@@
 However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph.
-To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
+To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent'' nodes when applying our histogram technique.
 Another approach would be to discard provenance data in computing Parzen-Window density estimates after the data have reached a certain age.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ACKNOWLEDGEMENTS
+% ACKNOWLEDGMENTS
 %
